---
title: "FIEBRE Adult Severity Scores analysis: draft results"
output:
  officedown::rdocx_document:
    reference_docx: reference.docx
knit: (
  function(inputFile, encoding) { 

    pSubTitle <- paste0('adult_severity_scores_', as.character(Sys.Date()))

    rmarkdown::render( 
      input       = inputFile, 
      encoding    = encoding, 
      params      = list(sub_title = pSubTitle),      
      output_file = pSubTitle) })
---

```{r libraries, include = F}

# Libraries ----

library(tidyverse)
library(kableExtra)
library(knitr)
library(forcats)
library(janitor)
library(gtsummary)
library(flextable)
library(ggforce)
library(officedown)
library(officer)
library(pROC)
library(plotROC)
library(finalfit)
# Set flextable defaults

set_flextable_defaults(
  font.family = "Tahoma",
  font.size = 8,
  font.color = "black",
  text.align = "center",
  padding = 0,
  padding.bottom = 0,
  padding.top = 0,
  padding.left = 0,
  padding.right = 0,
)

# officedown::rdocx_document:

# theme_gtsummary_compact()

```

```{r global-options, include=FALSE}

knitr::opts_chunk$set(fig.width = 7, fig.height = 5, dpi = 200, 
                      fig.path = 'Figs/', fig.keep = 'all', 
                      echo = FALSE, warning = FALSE, message = FALSE
                      )

fp <- fp_par(
  text.align = "justify", 
  padding.bottom = 20, padding.top = 120, 
  border.bottom = fp_border())

ft <- fp_text(shading.color='#EFEFEF', bold = TRUE)

```

```{r import_data, include = F, message = F, warning = F, paged.print = F, eval = T}

data_file <- "/Users/sham/Filr/My Files/sync/My Files/odk/fiebre_server/analysis_data/fiebre_adult_child_sites_combined_2021-08-26.csv"

fiebre_data <- read_csv(data_file, col_types = cols(.default = "c"))

```

```{r clean_data, include = F, message = F, warning = F, paged.print = F, eval = T}

# Select data

severity_data <- fiebre_data %>%
  select(consent_participate, patient_id_calculate, country, sex, px_group, modGCS, qsofa, uva, crb_65, mews, cf_9_px_status_final, cf_4_discharge_outcome, discharge_outcome, cf_5_inpatient_outcome, hiv_comb, age, tx_any_antimicro, hiv, hiv_test_final, hiv_comb, temp, respiratory_rate, respiratory_rate_explain, bp_systolic, bp_diastolic, heart_rate, o2_sats, px_orient) %>%
  mutate(across(c(temp, heart_rate, bp_systolic, bp_diastolic, respiratory_rate, modGCS, qsofa, uva, crb_65, mews, o2_sats, age), as.numeric)) %>%
  filter(age >= 16) %>%
  # select(country, px_group, cf_4_discharge_outcome, discharge_outcome, cf_5_inpatient_outcome, cf_9_px_status_final) %>%
  mutate(cf_9_px_status_final = str_replace_all(cf_9_px_status_final, "[^[:alnum:]]", " "),
         cf_5_inpatient_outcome = str_replace_all(cf_5_inpatient_outcome, "[^[:alnum:]]", " "),
         cf_4_discharge_outcome = str_replace_all(cf_4_discharge_outcome, "[^[:alnum:]]", " "),
         discharge_outcome = str_replace_all(discharge_outcome, "[^[:alnum:]]", " ")
         ) %>%
  mutate(outcome_discharge = case_when(
           !is.na(cf_4_discharge_outcome) & px_group == "outpatient-adult" ~ cf_4_discharge_outcome,
           !is.na(discharge_outcome) & px_group == "outpatient-adult" ~ discharge_outcome,
           !is.na(cf_5_inpatient_outcome) & px_group == "inpatient-adult" ~ cf_5_inpatient_outcome,
           TRUE ~ NA_character_)
         ) %>%
  # select(cf_5_inpatient_outcome) %>%
  # distinct()
  mutate(outcome_d28 = case_when(cf_9_px_status_final == "alive comp recov" ~ "Alive", 
                               cf_9_px_status_final == "alive improved" ~ "Alive",
                               cf_9_px_status_final == "alive no info" ~ "Alive",
                               cf_9_px_status_final == "alive same day0" ~ "Alive",
                               cf_9_px_status_final == "alive worse day0" ~ "Alive", 
                               cf_9_px_status_final == "dead" ~ "Dead", 
                               TRUE ~ NA_character_),
        outcome_inpatient = case_when(cf_5_inpatient_outcome == "discharged" ~ "Alive", 
                               cf_5_inpatient_outcome == "discharged pallative" ~ "Alive",
                               cf_5_inpatient_outcome == "other" ~ "Alive",
                               cf_5_inpatient_outcome == "died" ~ "Dead",
                               cf_5_inpatient_outcome == "dead" ~ "Dead", 
                               TRUE ~ NA_character_),
        outcome_all = case_when(
          outcome_d28 == "Alive" ~ "Alive",
          outcome_d28 == "Dead" ~ "Dead",
          outcome_inpatient == "Alive" ~ "Alive",
          outcome_inpatient == "Dead" ~ "Dead",
          TRUE  ~ NA_character_
        )
        ) %>%
  mutate(altered_mental_status = case_when(
           modGCS < 15 ~ "Yes",
           modGCS >= 15 ~ "No",
           TRUE ~ NA_character_
         ),
         age_cat = case_when(
           age >= 15 & age < 35 ~ 1,
           age >= 35 & age < 55 ~ 2,
           age >= 55 & age < 70 ~ 3,
           age >= 70 ~ 4,
         )
         ) %>%
    mutate(sex = str_to_sentence(sex), 
           px_group = str_to_sentence(str_remove(px_group, "-adult")),
           hiv_comb = str_to_sentence(hiv_comb)
           ) %>%
  mutate(across(c(country, age_cat, sex, hiv_comb, cf_5_inpatient_outcome, cf_9_px_status_final, px_group, px_orient, outcome_d28, outcome_inpatient, outcome_discharge, outcome_all, altered_mental_status), as.factor)) %>%
  mutate(across(c(respiratory_rate, bp_systolic, bp_diastolic, heart_rate, o2_sats, temp), ~ifelse(.x >= 999, NA, .x)))


# relevel factors

relev_yndk <- function(f) forcats::fct_relevel(factor(f), "Yes", "No", "Don't know")
relev_yn <- function(f) forcats::fct_relevel(factor(f), "Yes", "No")
relev_pos_neg <- function(f) forcats::fct_relevel(factor(f), "Positive", "Negative", "Indet", "Don't know")

severity_data$altered_mental_status <- relev_yn(severity_data$altered_mental_status)

severity_data$hiv_comb <- relev_pos_neg(severity_data$hiv_comb)


# Label dataset variables

severity_data <- severity_data %>%
  sjlabelled::var_labels(
                age = "Age (years)",
                age_cat = "Age group",
                sex = "Sex",
                px_group = "Participant group",
                temp = "Temperature (°C)",
                hiv_comb = "HIV status",
                respiratory_rate = "Respiratory rate (breaths/min)",
                bp_systolic = "Systolic blood pressure, mm Hg",
                bp_diastolic = "Diastolic blood pressure, mm Hg",
                heart_rate = "Heart rate, beats/min",
                o2_sats = "Oxygen saturation (%)",
                mews = "Modified early warning score (MEWS)",
                crb_65 = "CRB-65",
                modGCS = "Modified GCS",
                qsofa = "qSOFA",
                uva = "UVA",
                altered_mental_status = "Altered mental status \n (Glasgow Coma Scale <15)",
                outcome_inpatient = "Inpatient outcome",
                outcome_d28 = "Day 28 outcome",
                outcome_all = "All outcomes",
                px_group = "Patient group",
                cf_9_px_status_final = "All day 28 outcomes",
                cf_5_inpatient_outcome = "All inpatient discharge outcomes",
                outcome_discharge = "All day 0 discharge outcomes"
) 

# x <- fiebre_data %>%
#   select(patient_id_calculate, inpatient_outpatient, contains("temp")) %>%
#   mutate(temp = as.numeric(temp)) %>%
#   filter(temp >= 999)
  

summary(as.factor(severity_data$outcome_d28))
summary(as.factor(severity_data$outcome_inpatient))

tabyl(severity_data, outcome_inpatient, outcome_d28) %>%
  adorn_totals(where = "row")

summary(as.factor(severity_data$outcome_all))


glimpse(severity_data$crb_65)
median(severity_data$crb_65, na.rm = T)

glimpse(severity_data$mews)
glimpse(severity_data$altered_mental_status)
glimpse(severity_data$sex)


# adorn_totals(where = "row") %>%             # add a total row
#   adorn_percentages(denominator = "col") %>%  # convert to proportions
#   adorn_pct_formatting() %>%                  # convert to percents
#   adorn_ns(position = "front") %>%            # display as: "count (percent)"
#   adorn_title(                                # adjust titles
#     row_name = "Age Category",
#     col_name = "Gender")

summary(as.factor(severity_data$cf_5_inpatient_outcome))
summary(as.factor(severity_data$cf_9_px_status_final))
summary(as.factor(severity_data$px_group))


severity_data <- severity_data %>% 
  mutate(uva4 = ifelse(uva > 4,1,0),
                  mews4=ifelse(mews > 4,1,0),
                  qsofa2=ifelse(qsofa > 2,1,0))



summary(as.factor(severity_data$px_group))
summary(as.factor(severity_data$country))
summary(as.factor(severity_data$outcome_d28))



# Three populations
#  1 - Day28 outcomes
#  2 - Inpatient outcomes
#  3 - Any death



# labels(mz_demo)  <- c(consent_participate = 'Number consented', sex = "Sex", age = "Age (years)", temp = "Temperature (celsius)", px_group = "Participant group")

#  1 - Day28 outcomes

total_px <- nrow(severity_data)

severity_data_no_outcome_d28 <- severity_data %>%
  filter(is.na(outcome_d28)) %>%
  select(patient_id_calculate) %>%
  mutate(no_outcome_d28 = 1) 

px_no_outcome <- severity_data_no_outcome_d28 %>%
  nrow()

no_sev_score <- left_join(
  severity_data,
  severity_data_no_outcome_d28,
  by = "patient_id_calculate"
)


px_no_sev_score <- no_sev_score %>%
  filter(is.na(no_outcome_d28)) %>%
  # select(modGCS, qsofa, uva, mews) %>%
  filter(if_any(c(modGCS, qsofa, uva, mews), ~ is.na(.))) %>%
  nrow()

severity_data_cc_d28 <- no_sev_score %>%
  filter(is.na(no_outcome_d28)) %>%
  # select(modGCS, qsofa, uva, mews, no_outcome) %>%
  filter(if_all(c(modGCS, qsofa, uva, mews), ~ !is.na(.))) 

total_px - px_no_outcome - px_no_sev_score

# LTFU

ltfu <- severity_data %>%
  select(cf_9_px_status_final) %>%
  filter(cf_9_px_status_final == "lost to fup") %>%
  nrow()

missing_day28 <- severity_data %>%
  select(cf_9_px_status_final) %>%
  filter(is.na(cf_9_px_status_final)) %>%
  nrow()


```

<!---BLOCK_TOC--->

<!---BLOCK_TOC{seq_id: 'tab'}--->

<!---BLOCK_TOC{seq_id: 'fig'}--->



\newpage

<!---BLOCK_LANDSCAPE_START--->


## Number of patients recruited

* These are patient outcomes reported at day 28 follow-up. 
* Patient group (whether a patient is an outpatient or an inpatient) is defined at recruitment.   

```{r, tab.cap = "Day 28 follow-up outcomes",  include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T }

tbl_outcome_d28 <-
  severity_data  %>%
    select(px_group, country, cf_9_px_status_final, outcome_d28) %>%
    tbl_summary(
    by = country,
    type = list(all_dichotomous() ~ "categorical",
                all_continuous() ~ "continuous2"
                ),
    statistic = list(c("px_group",  "cf_9_px_status_final") ~ " {n} ({p}%)"),
    digits = all_continuous() ~ 1,
    missing_text = "(Missing)"
      ) %>%
      add_stat_label() %>%
      add_overall() %>%
     modify_header(list(label ~"Variable")) %>%
     as_flex_table() %>%
      flextable::fontsize(part = "header", size = 8) %>%
      flextable::autofit()

tbl_outcome_d28

```

```{r, tab.cap = "Day 28 outcome by patient group",  include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T }

tbl_outcome_d28_px_group <-
  severity_data_cc_d28 %>%
  tbl_cross(row = px_group, col = outcome_d28, percent = "row") %>%
   modify_header(list(label ~ "")) %>%
  as_flex_table() %>%
  flextable::fontsize(part = "header", size = 8)

tbl_outcome_d28_px_group

```

\newpage

* These are patient outcomes reported at day 0. 

```{r, tab.cap = "Day 0 outcomes",  include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T }

tbl_outcome_d0 <-
  severity_data  %>%
    select(px_group, country, outcome_discharge) %>%
    tbl_summary(
    by = country,
    type = list(all_dichotomous() ~ "categorical",
                all_continuous() ~ "continuous2"
                ),
    statistic = list(c("px_group",  "outcome_discharge") ~ " {n} ({p}%)"),
    digits = all_continuous() ~ 1,
    missing_text = "(Missing)"
      ) %>%
      add_stat_label() %>%
      add_overall() %>%
     modify_header(list(label ~"Variable")) %>%
     as_flex_table() %>%
      flextable::fontsize(part = "header", size = 8) %>%
      flextable::autofit()

tbl_outcome_d0

```

```{r, tab.cap = "Day 0 outcomes by patient group",  include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T }

tbl_outcome_d0_px_group <-
  severity_data %>%
  tbl_cross(row = px_group, col = outcome_discharge, percent = "row") %>%
   modify_header(list(label ~ "")) %>%
  as_flex_table() %>%
  flextable::fontsize(part = "header", size = 8)

tbl_outcome_d0_px_group %>%
  flextable::width(j = 1, width = 1) %>%
  flextable::width(j=c(2,3,4,5,6,7,8,9,10,11,12), width = 0.5)


  # width(j=1, width = 2.7) %>% 
  # width(j=2, width = 1.5) %>% 
  # width(j=c(4,5,7,8), width = 1)

```

Due to small number of deaths at day 0 (only 87 inpatient deaths), this analysis is limited to outcomes at day 28.    
* Total number of recruited adult patients: __`r total_px`__  
* Number of patients _missing_ outcome (dead or alive) data at day 28 follow-up: __`r px_no_outcome`__.   
    -   These were patients loss to follow-up at day 28 (__`r ltfu`__) and where no outcome data were reported at day 28 (__`r missing_day28`__).   
* Number _missing_ severity score data: __`r px_no_sev_score`__.      
* Patients with _complete_ outcome data at day 28 and _complete_ clinical data (complete cases): __`r total_px - px_no_outcome - px_no_sev_score`__   

<!---BLOCK_LANDSCAPE_STOP--->

\newpage


## Descriptive results tables


```{r tab.cap = "Demographic and clinical characteristics of patients with complete data at day 28", tab.id = "tbl_1a", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T}


tbl_1a_d28 <- severity_data_cc_d28 %>%
  select(age, age_cat, sex, px_group, temp, hiv_comb, respiratory_rate, bp_systolic, bp_diastolic, heart_rate, o2_sats, mews, crb_65, modGCS, qsofa, uva, altered_mental_status) %>%
  tbl_summary(
    # by = country,
    type = list(all_dichotomous() ~ "categorical",
                all_continuous() ~ "continuous2"
                ),
    statistic = list(c("age", "temp", "respiratory_rate", "bp_systolic", "bp_diastolic", "heart_rate", "mews" ,"o2_sats", "modGCS", "uva") ~ 
                     c("{median} ({p25}, {p75})", "{min}, {max}"),
                     c("px_group",  "hiv_comb", "altered_mental_status") ~ " {n} ({p}%)"),
    digits = all_continuous() ~ 1,
    missing_text = "(Missing)"
      ) %>%
    add_stat_label() %>%
     modify_header(list(label ~"Variable")) %>%
    as_flex_table() %>%
      flextable::fontsize(part = "header", size = 8) %>%
      flextable::autofit()

tbl_1a_d28


```

\newpage

<!---BLOCK_LANDSCAPE_START--->


```{r, tab.cap = "Demographic and clinical characteristics of patients with complete data at day 28, by site", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T }


tbl_1a_d28 <- severity_data_cc_d28 %>%
  select(country, age, age_cat, sex, px_group, temp, hiv_comb, respiratory_rate, bp_systolic, bp_diastolic, heart_rate, o2_sats, mews, crb_65, modGCS, qsofa, uva, altered_mental_status) %>%
  tbl_summary(
    by = country,
    type = list(all_dichotomous() ~ "categorical",
                all_continuous() ~ "continuous2"
                ),
    statistic = list(c("age", "temp", "respiratory_rate", "bp_systolic", "bp_diastolic", "heart_rate", "mews" ,"o2_sats", "modGCS", "uva") ~ 
                     c("{median} ({p25}, {p75})", "{min}, {max}"),
                     c("px_group", "age_cat", "hiv_comb", "altered_mental_status") ~ "{n} ({p}%)"),
    digits = all_continuous() ~ 1,
    missing_text = "(Missing)"
      ) %>%
    add_stat_label() %>%
     modify_header(list(label ~"Variable")) %>%
    as_flex_table() %>%
      flextable::fontsize(part = "header", size = 8) %>%
      flextable::autofit()

tbl_1a_d28

```


```{r fig1, fig.cap = "Age distribution by sex",  include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T }

fig1 <- severity_data_cc_d28 %>%
  ggplot(mapping = aes(x = age, color = sex)) +
  geom_freqpoly(binwidth = 2, size = 1.5) 

fig1

```



```{r fig2, fig.cap = "Age distribution by sex and site",  include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T }

fig2 <- severity_data_cc_d28 %>%
  ggplot(mapping = aes(x = age, color = sex)) +
  geom_freqpoly(binwidth = 2, size = 1.5) +
  facet_wrap(vars(country))

fig2

```


```{r fig3, fig.cap = "Distribution of patient outcomes by age", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T}

fig3 <- ggplot(
  data = severity_data_cc_d28, 
  aes(y = age,           # numeric variable
      x = outcome_d28)) +    # group variable
  geom_violin(
    aes(fill = outcome_d28), # fill (color of violin background)
    color = "white",     # white outline
    alpha = 0.2)+        # transparency
  geom_sina(
    size=1,                # Change the size of the jitter
    aes(color = outcome_d28))+ # color (color of dots)
  scale_fill_manual(       # Define fill for violin background by death/recover
    values = c("Death" = "#bf5300", 
              "Recover" = "#11118c")) + 
  scale_color_manual(      # Define colours for points by death/recover
    values = c("Death" = "#bf5300", 
              "Recover" = "#11118c")) + 
  theme_minimal() +                                # Remove the gray background
  theme(legend.position = "none") +                # Remove unnecessary legend
  labs(title = "") 

fig3


```

<!---BLOCK_LANDSCAPE_STOP--->


```{r fig4a, fig.cap = "Distribution of MEWs by Day 28 outcome", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T}

mews_tbl <- severity_data_cc_d28 %>%
  count(mews, outcome_d28) %>%  
  group_by(outcome_d28) %>%        
  mutate(proportion = n/sum(n)*100,
         mews = as.factor(mews)) 

fig4a <- ggplot(data = mews_tbl, 
                aes(x = mews, y = proportion, fill = outcome_d28)) +
    scale_y_continuous(limits = c(0, 60), name = "Percent (%)") +
    # geom_bar()
    scale_x_discrete(name = "Modified Early Warning Score (MEWS)") +
    geom_bar(stat = "identity", position = position_dodge2(preserve = "single", padding = 0.1), width = 1.0) +
    theme_minimal() + 
    labs(fill = "Day 28 patient outcome")

fig4a

```


```{r fig4b, fig.cap = "Scatter plot of MEWs by Day 28 outcome", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T}

mews_tbl2 <- severity_data_cc_d28 %>%
  # count(mews, outcome_d28) %>%
  # group_by(outcome_d28) %>%
  mutate(mews = as.factor(mews))

fig4b <- ggplot(data = mews_tbl2, 
                aes(x = outcome_d28, y = mews)) +
    scale_y_discrete(name = "Modified Early Warning Score (MEWS)") +
    # geom_bar()
    scale_x_discrete(name = "Day 28 outcome") +
    geom_jitter() +
    theme_minimal()


fig4b

```


```{r fig5a, fig.cap = "Scatter plot of qSOFA by Day 28 outcome", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T}


qsofa_tbl2 <- severity_data_cc_d28 %>%
  count(qsofa, outcome_d28) %>%  
  group_by(outcome_d28) %>%        
  mutate(proportion = n/sum(n)*100,
         qsofa = as.factor(qsofa)) 

fig5a <- ggplot(data = qsofa_tbl2, 
                aes(x = qsofa, y = proportion, fill = outcome_d28)) +
    scale_y_continuous(limits = c(0, 60), name = "Percent (%)") +
    # geom_bar()
    scale_x_discrete(name = "Quick sequential organ failure assessment score (qSOFA)") +
    geom_bar(stat = "identity", position = position_dodge2(preserve = "single", padding = 0.1), width = 1.0) +
    theme_minimal() + 
    labs(fill = "Day 28 patient outcome") +
    theme_minimal()

fig5a

```

```{r fig5b, fig.cap = "Distribution of qSOFA by Day 28 outcome", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T}

qsofa_tbl <- severity_data_cc_d28 %>%
  # count(mews, outcome_d28) %>%
  # group_by(outcome_d28) %>%
  mutate(qsofa = as.factor(qsofa))

fig5b <- ggplot(data = qsofa_tbl, 
                aes(x = outcome_d28, y = qsofa)) +
    scale_y_discrete(name = "Quick sequential organ failure assessment score (qSOFA)") +
    # geom_bar()
    scale_x_discrete(name = "Day 28 outcome") +
    geom_jitter() +
    theme_minimal()

fig5b

```


```{r fig6a, fig.cap = "Distribution of UVA score by Day 28 outcome", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T}

uva_tbl2 <- severity_data_cc_d28 %>%
  count(uva, outcome_d28) %>%  
  group_by(outcome_d28) %>%        
  mutate(proportion = n/sum(n)*100,
         uva = as.factor(uva)) 

fig6a <- ggplot(data = uva_tbl2, 
                aes(x = uva, y = proportion, fill = outcome_d28)) +
    scale_y_continuous(limits = c(0, 70), name = "Percent (%)") +
    # geom_bar()
    scale_x_discrete(name = "Universal Vital Assessment (UVA) score") +
    geom_bar(stat = "identity", position = position_dodge2(preserve = "single", padding = 0.1), width = 1.0) +
    theme_minimal() + 
    labs(fill = "Day 28 patient outcome") 

fig6a


```

```{r fig6b, fig.cap = "Scatter plot of UVA by Day 28 outcome", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T}


uva_tbl <- severity_data_cc_d28 %>%
  # count(mews, outcome_d28) %>%
  # group_by(outcome_d28) %>%
  mutate(uva = as.factor(uva))

fig6b <- ggplot(data = uva_tbl, 
                aes(x = outcome_d28, y = uva)) +
    scale_y_discrete(name = "Universal Vital Assessment (UVA) score") +
    # geom_bar()
    scale_x_discrete(name = "Day 28 outcome") +
    geom_jitter() +
    theme_minimal()

fig6b


```

\newpage

## UVA results tables and figures

```{r, fig7, fig.cap = "UVA Receiver operating characteristic curve", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T }


uva_auc <- roc(severity_data_cc_d28$outcome_d28, 
           severity_data_cc_d28$uva, percent = T)
# 
# severity_data_cc_d28 %>%
#   roc(outcome_d28, uva) %>%
#   coords(ret = "all", transpose = F)


# unique(severity_data_cc_d28$uva)

# plot_uva <- plot.roc(severity_data_cc_d28$outcome_d28, 
#         severity_data_cc_d28$uva,
#                    percent=TRUE,
#                    # ci = TRUE, of="thresholds",
#                    print.auc = F,
#         thresholds = T, # select the (best) threshold
#         print.thres = "all", 
#         legacy.axes = TRUE
# )
# coords(plot_uva)

# x <- plot(uva_auc, print.thres="all", 
#       print.auc=TRUE, 
#      auc.polygon=TRUE, 
#      print.thres.pattern=ifelse(uva_auc$percent, "%.1f (%1.1f%%, %1.1f%%)", "%.1f (%1.1f, %1.1f)"), 
#      legacy.axes = TRUE)
# calc_auc(calc_auc)$AUC

fig7 <- ggplot(severity_data_cc_d28, aes(d = outcome_d28, m = uva)) +
  geom_roc(n.cuts = 20, labelsize = 5, labelround = 2)

fig7 +
  annotate("text", x = .5, y = .5,
  label = paste("UVA score AUC =", round(calc_auc(fig7)$AUC, 2))) +
  scale_x_continuous("1 - Specificity",
                     breaks = seq(0, 1, by = 0.1),
                     labels = scales::label_percent()
                     ) +
  scale_y_continuous("Sensitivity",
                     breaks = seq(0, 1, by = 0.1),
                     labels = scales::label_percent()
                     )

# fig7

uva_auc_value <- round(as.numeric(auc(uva_auc)[1])/100, 2)

uva_auc_ci <- paste0("95% CI: ", round(ci.auc(uva_auc)[1]/100, 2), "-", round(ci.auc(uva_auc)[3]/100, 2))
 
# paste0(round(uva_auc_value, 2)), "",  )


```

UVA area under the curve = `r uva_auc_value`,  `r uva_auc_ci`   


```{r, tab.cap = "UVA Sensitivity and specificity values", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = F }

uva_values <- coords(roc = uva_auc, x = "all", transpose = FALSE, ret = "all")

uva_values <- uva_values %>%
  select(threshold, specificity, sensitivity, "1-specificity", accuracy, npv, ppv) %>%
  mutate(across(c(sensitivity, specificity, "1-specificity", accuracy, npv, ppv), ~ round(.*100,1) )) 

uva_values_plot <- uva_values %>%
  select(threshold, specificity, specificity_1 = "1-specificity", sensitivity, specificity) %>%
  mutate(across(c(threshold, specificity, specificity_1, sensitivity), ~ .*100)) %>%
  mutate(label = paste0("(", round(specificity, 1), ", ", round(sensitivity, 1), ")"))


uva_values_plot %>%
  ggplot(mapping = aes(x = specificity, y = sensitivity)) +
  geom_line() +
  geom_point() +
  geom_label(aes(specificity +.5 , sensitivity + 0.5, label = label), 
             hjust = 0, nudge_x = 3,
             label.size = NA
             ) +
  scale_x_reverse("Specificity", limits = c(100, 0),  breaks = scales::breaks_extended(10)) + 
  scale_y_continuous("Sensitivity", limits = c(0, 100), breaks = scales::breaks_extended(10)) 

```
\newpage

### UVA regression results


```{r, tab.cap = "UVA Univariate and multivariate regression", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T }


severity_data_cc_d28_reg <- severity_data_cc_d28 %>%
  # filter(uva != 10) %>%
  mutate(uva_fact = as.factor(uva)) 

# Univariate regression

# uva_tbl_reg <- severity_data_cc_d28_reg %>%
#   finalfit(dependent, explanatory) %>%
#   ff_remove_p()

# uva_uv_reg <- glm(outcome_d28 ~ uva_fact + age_cat + sex, family = "binomial", data = severity_data_cc_d28_reg)


uva_uv_reg_tbl <- severity_data_cc_d28_reg %>%
  select("uva_fact", "age_cat", "sex", "outcome_d28") %>%
  tbl_uvregression(
    method = glm,
    y = outcome_d28,
    method.args = list(family = binomial),
    exponentiate = TRUE,
    pvalue_fun = ~style_pvalue(.x, digits = 2), 
    hide_n = T
  ) %>%
  add_global_p() %>% # add global p-value 
  modify_header(update = list(
    label ~ "Variable",
    estimate ~ "Unadjusted OR",
    ci ~ "95% CI",
    p.value ~ "p-value"
  )
  )
  # add_nevent() %>%    # add number of events of the outcome
  # add_q() %>%         # adjusts global p-values for multiple testing
  # bold_p() %>%        # bold p-values under a given threshold (default 0.05)
  # bold_p(t = 0.10, q = TRUE) %>% # now bold q-values under the threshold of 0.10
  # bold_labels()

#  Multivariate regression

uva_mv_reg <- glm(outcome_d28 ~ uva_fact + age_cat + sex, family = "binomial", data = severity_data_cc_d28_reg)

# summary(uva_mv_reg)



uva_mv_reg_tbl <- tbl_regression(uva_mv_reg, 
               exponentiate = TRUE, 
               pvalue_fun = ~style_pvalue(.x, digits = 2)
  ) %>%
  add_global_p() %>%
    modify_header(update = list(
    label ~ "Variable",
    estimate ~ "Adjusted OR",
    ci ~ "95% CI",
    p.value ~ "p-value"
  )
  )

# counts

uva_tbl_count <- severity_data_cc_d28_reg %>% 
  select(uva_fact, age_cat, sex, outcome_d28) %>% # keep only columns of interest
  tbl_summary(     
    by = outcome_d28,                                               # stratify entire table by outcome
    percent = "row",
    statistic = list(c(age_cat) ~ "{n} ({p}%)"),   # stats and format for categorical columns
    digits = all_continuous() ~ 1,                              # rounding for continuous columns
    type   = all_categorical() ~ "categorical"                 # force all categorical levels to display
) %>%
  add_overall() %>%
      modify_header(update = list(
    label ~ "Variable"
  )
  )


uva_reg_tbl <- tbl_merge(
  tbls = list(uva_tbl_count, uva_uv_reg_tbl, uva_mv_reg_tbl)
) %>%
  # tab_spanner = c("****","**Univariate**", "**Multivariable**")) %>%
  as_tibble() 

col_Names <- names(uva_reg_tbl)

col_Names <- str_remove_all(col_Names, "\\*")

for(i in 1 : length(uva_reg_tbl))
{
  col_Names[i] <- paste0(col_Names[i], paste0(rep("\r", i), collapse = ""), collapse = "")
}

colnames(uva_reg_tbl) <- col_Names

uva_reg_tbl <- uva_reg_tbl %>%
  mutate("Overall, N = 3,481\r\r" = str_remove_all(uva_reg_tbl[[2]], " \\(100%\\)"))

uva_reg_tbl <- uva_reg_tbl %>%
  flextable() %>%
  flextable::fontsize(part = "header", size = 8) %>%
  flextable::add_footer_lines("OR = Odds ratio, CI = Confidence intervals") %>%
  align_text_col(align = "center")


uva_reg_tbl

# str(severity_data_cc_d28)
# 
# explanatory <- c("uva_fact", "age_cat", "sex") 
# 
# dependent <- 'outcome_d28'

# uva_tbl_reg <- severity_data_cc_d28_reg %>%
#   finalfit(dependent, explanatory) %>%
#   ff_remove_p()


```


```{r, fig8, fig.cap = "UVA Adjusted odds ratio plot", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T }

uva_mv_reg_plot <- uva_mv_reg %>%
  broom::tidy(exponentiate = TRUE, conf.int = TRUE) %>%
  mutate(across(where(is.numeric), round, digits = 2))

## remove the intercept term from your multivariable results

fig8 <- uva_mv_reg_plot %>% 
    filter(!term %in%  c("(Intercept)", "age_cat2", "age_cat3", "age_cat4", "sexMale", "uva_fact10")) %>% 
  #set order of levels to appear along y-axis
  # mutate(term = fct_relevel(term,
  #   "uva_fact9", "uva_fact8", "uva_fact7","uva_fact6", "uva_fact5", "uva_fact4", "uva_fact3", "uva_fact2", "uva_fact1")
  #   ) %>%

  ggplot(aes(x = estimate, y = term)) +
  geom_point(size = 3, shape=19) + 
  geom_errorbar(aes(xmin = conf.low, xmax = conf.high), width = .3) + 
  geom_vline(xintercept = 1, linetype = "dashed") +
  scale_x_log10(name = "Adjusted odds ratio with 95% CI", 
                breaks = scales::log_breaks(8)) +
  scale_y_discrete(name = "Universal Vital Assessment (UVA) score", 
                   labels = c("1", "2", "3", "4", "5", "6", "7", "8", "9")
                   ) +
  coord_flip() +
  theme_minimal()

fig8

```
