---
title: "FIEBRE Adult Severity Scores analysis: Figure and tables for manuscript"
date: "`r format(Sys.Date(), '%d %B %Y')`   \n"
output:
  officedown::rdocx_document:
    reference_docx: reference.docx
    tables:
      style: Table
      layout: autofit
      width: 1.0
      caption:
        style: Table Caption
        pre: 'Table '
        sep: ': '
      conditional:
        first_row: true
        first_column: false
        last_row: false
        last_column: false
        no_hband: false
        no_vband: true
    plots:
      style: Normal
      align: left
      caption:
        style: Image Caption
        pre: 'Figure '
        sep: ': '
    lists:
      ol.style: null
      ul.style: null
    mapstyles:
      Normal: ['First Paragraph', 'Author', 'Date']
    page_size:
      width: 8.3
      height: 11.7
      orient: "portrait"
    page_margins:
      bottom: 0.5
      top: 0.5
      right: 1
      left: 1
      header: 0.2
      footer: 0.2
      gutter: 0.2
    reference_num: true
knit: (
  function(inputFile, encoding) { 

    pSubTitle <- paste0('analysis_reports/adult_severity_scores_db2024-06-16_', as.character(Sys.Date()))

    rmarkdown::render( 
      input       = inputFile, 
      encoding    = encoding, 
      params      = list(sub_title = pSubTitle),      
      output_file = pSubTitle) })
---

```{r libraries, include = F}

# Libraries ----

library(tidyverse)
library(kableExtra)
library(knitr)
library(forcats)
library(janitor)
library(gtsummary)
library(flextable)
library(ggforce)
library(officedown)
library(officer)
library(Gmisc)
library(pROC)
library(plotROC)
library(finalfit)
library(glue)
library(htmlTable)
library(grid)
library(ragg)
# Set flextable defaults

set_flextable_defaults(
  font.family = "Tahoma",
  font.size = 8,
  font.color = "black",
  text.align = "center",
  padding = 0,
  padding.bottom = 0,
  padding.top = 0,
  padding.left = 0,
  padding.right = 0,
)

# officedown::rdocx_document:

# theme_gtsummary_compact()

```

```{r global-options, include=FALSE}

knitr::opts_chunk$set(fig.width = 7, fig.height = 5, dpi = 350, 
                      fig.path = 'Figs/', fig.keep = 'all', 
                      echo = FALSE, warning = FALSE, message = FALSE,
                      dev = "ragg_png"
                      )

fp <- fp_par(
  text.align = "justify", 
  padding.bottom = 20, padding.top = 120, 
  border.bottom = fp_border())

ft <- fp_text(shading.color='#EFEFEF', bold = TRUE)

```


```{r import_data, eval=T, message=FALSE, warning=FALSE, include=F, paged.print=FALSE}

data_file <- paste0(here::here(), "/data/", "fiebre_cases_adult_child_sites_combined_2024-05-26.csv")


fiebre_data <- read_csv(data_file, col_types = cols(.default = "c"))

```



```{r select_data, include = F, message = F, warning = F, paged.print = F, eval = T}

# Select variables and adults (aged > 15yrs)

severity_data <- fiebre_data %>%
  select(consent_participate, patient_id_calculate, date_day0_visit, country, sex, px_group, cf_9_px_status_final, cf_9_date_death, cf_4_discharge_outcome, cf_4_other_specify, discharge_outcome, cf_5_inpatient_outcome, cf_5_inpatient_outcome_other, cf_5_discharge_date, hiv_comb, age, tx_any_antimicro, hiv, hiv_test_final, hiv_comb, cf_hiv_hiv_test_done, cf_hiv_no_hiv_other, ear_temperature, axil_temperature, temp, respiratory_rate, respiratory_rate_explain, bp_systolic, bp_diastolic, heart_rate, o2_sats, px_orient, px_coop, px_coop_no, px_assess_speech, px_eyes, px_eyes_na, px_eyes_no, outp_switched) %>%
  mutate(across(c(age, temp, ear_temperature, axil_temperature, heart_rate, bp_systolic, bp_diastolic, o2_sats, respiratory_rate), as.numeric)) %>%
  filter(age >= 15)

```



```{r define_severity_scores, include = F, message = F, warning = F, paged.print = F, eval = T}

# Define adult severity severity scores functions

# qSOFA Score ----

qsofa<-function(severity_data){
  severity_data$gcs_u15<-NA
  severity_data$gcs_u15[which((severity_data$px_eyes=="yes")&(severity_data$px_orient=="yes")&(severity_data$px_coop=="yes"))]<-0
  severity_data$gcs_u15[which((severity_data$px_eyes!="yes")|(severity_data$px_orient!="yes")|(severity_data$px_coop!="yes"))]<-1
    severity_data$bp_syst_lt100<-NA
  severity_data$bp_syst_lt100[which(severity_data$bp_systolic<=100)]<-1
  severity_data$bp_syst_lt100[which(severity_data$bp_systolic>100)]<-0
  severity_data$high_resp<-NA
  severity_data$high_resp[which((severity_data$respiratory_rate<98)&(severity_data$respiratory_rate>=22))]<-1
  severity_data$high_resp[which((severity_data$respiratory_rate<22)&(severity_data$respiratory_rate>0))]<-0
  severity_data$qsofa<-severity_data$bp_syst_lt100+severity_data$high_resp+severity_data$gcs_u15
  return((severity_data$qsofa))
}

# UVA Score ----


uva<-function(severity_data){
  severity_data$gcs_u15<-NA
  severity_data$gcs_u15[which((severity_data$px_eyes=="yes")&(severity_data$px_orient=="yes")&(severity_data$px_coop=="yes"))]<-0
  severity_data$gcs_u15[which((severity_data$px_eyes!="yes")|(severity_data$px_orient!="yes")|(severity_data$px_coop!="yes"))]<-1
  
  severity_data$uva<-0
  severity_data$uva[which(severity_data$ear_temperature<36)]<-severity_data$uva[which(severity_data$ear_temperature<36)]+2
  severity_data$uva[which(severity_data$axil_temperature<36&is.na(severity_data$ear_temperature))]<-severity_data$uva[which(severity_data$axil_temperature<36&is.na(severity_data$ear_temperature))]+2
  severity_data$uva[which(is.na(severity_data$axil_temperature)&is.na(severity_data$ear_temperature))]<-NA

  severity_data$uva[which(is.na(severity_data$heart_rate)|is.na(severity_data$respiratory_rate)|is.na(severity_data$bp_systolic)|is.na(severity_data$o2_sats)|is.na(severity_data$gcs_u15)|is.na(severity_data$hiv_comb))]<-NA

  severity_data$uva[which(severity_data$heart_rate>=120)]<-severity_data$uva[which(severity_data$heart_rate>=120)]+1
  severity_data$uva[which(severity_data$respiratory_rate>=30)]<-severity_data$uva[which(severity_data$respiratory_rate>=30)]+1
  severity_data$uva[which(severity_data$bp_systolic<90)]<-severity_data$uva[which(severity_data$bp_systolic<90)]+1
  severity_data$uva[which(severity_data$o2_sats<92)]<-severity_data$uva[which(severity_data$o2_sats<92)]+2 
  severity_data$uva[which(severity_data$gcs_u15==1)]<-severity_data$uva[which(severity_data$gcs_u15==1)]+4 
  severity_data$uva[which(severity_data$hiv_comb=="Positive")]<-severity_data$uva[which(severity_data$hiv_comb=="Positive")]+2  
  severity_data$uva[which(is.na(severity_data$uva))]<-NA
  return((severity_data$uva))
}



# UVA age ---
# Adds +1 for >55 and +2 for >65

uva_age <- function(severity_data){
  severity_data$gcs_u15<-NA
  severity_data$gcs_u15[which((severity_data$px_eyes=="yes")&(severity_data$px_orient=="yes")&(severity_data$px_coop=="yes"))]<-0
  severity_data$gcs_u15[which((severity_data$px_eyes!="yes")|(severity_data$px_orient!="yes")|(severity_data$px_coop!="yes"))]<-1
  

  
  severity_data$uva<-0
  severity_data$uva[which(severity_data$ear_temperature<36)]<-severity_data$uva[which(severity_data$ear_temperature<36)]+2
  severity_data$uva[which(severity_data$axil_temperature<36&is.na(severity_data$ear_temperature))]<-severity_data$uva[which(severity_data$axil_temperature<36&is.na(severity_data$ear_temperature))]+2
  severity_data$uva[which(is.na(severity_data$axil_temperature)&is.na(severity_data$ear_temperature))]<-NA
  severity_data$uva[which(is.na(severity_data$heart_rate)|is.na(severity_data$respiratory_rate)|is.na(severity_data$bp_systolic)|is.na(severity_data$o2_sats)|is.na(severity_data$gcs_u15)|is.na(severity_data$hiv_comb))]<-NA
  severity_data$uva[which(severity_data$heart_rate>=120)]<-severity_data$uva[which(severity_data$heart_rate>=120)]+1
  severity_data$uva[which(severity_data$respiratory_rate>=30)]<-severity_data$uva[which(severity_data$respiratory_rate>=30)]+1
  severity_data$uva[which(severity_data$bp_systolic<90)]<-severity_data$uva[which(severity_data$bp_systolic<90)]+1
  severity_data$uva[which(severity_data$o2_sats<92)]<-severity_data$uva[which(severity_data$o2_sats<92)]+2 
  severity_data$uva[which(severity_data$gcs_u15==1)]<-severity_data$uva[which(severity_data$gcs_u15==1)]+4 
  severity_data$uva[which(severity_data$hiv_comb=="Positive")]<-severity_data$uva[which(severity_data$hiv_comb=="Positive")]+2  
  severity_data$uva[which(severity_data$age >= 55 & severity_data$age < 65)]<-severity_data$uva[which(severity_data$age >= 55 & severity_data$age < 65)]+1  
  severity_data$uva[which(severity_data$age >= 65)]<-severity_data$uva[which(severity_data$age >= 65)]+2
  severity_data$uva[which(is.na(severity_data$uva))]<-NA
  return((severity_data$uva))
}


# ---CRB 65 Score 

crb_65<-function(severity_data){
  severity_data$crb<-0
  severity_data$crb[is.na(severity_data$px_orient)|is.na(severity_data$respiratory_rate)|is.na(severity_data$heart_rate)|is.na(severity_data$bp_systolic)|is.na(severity_data$bp_diastolic)|is.na(severity_data$age)]<-NA
  severity_data$temp<-0
  severity_data$temp[which(severity_data$px_orient=="no")]<-1
  severity_data$crb<-severity_data$crb+severity_data$temp
  
  severity_data$temp<-0
  severity_data$temp[which(severity_data$respiratory_rate>=30)]<-1
  severity_data$crb<-severity_data$crb+severity_data$temp
  
  severity_data$temp<-0
  severity_data$temp[which(severity_data$heart_rate>=120)]<-1
  severity_data$crb<-severity_data$crb+severity_data$temp
  
  severity_data$temp<-0
  severity_data$temp[which(severity_data$bp_systolic<=90|severity_data$bp_diastolic<=60)]<-1
  severity_data$crb<-severity_data$crb+severity_data$temp
  
  severity_data$temp<-0
  severity_data$temp[which(severity_data$age>=65)]<-1
  severity_data$crb<-severity_data$crb+severity_data$temp
  return((severity_data$crb))
}

# MEWS Score ----

mews<-function(severity_data){
  severity_data$px_eyes_no <- fct_collapse(severity_data$px_eyes_no,
                                             verbal=c("pain verbal_request","verbal_request"))
  severity_data$avpu<-NA
  severity_data$avpu[which(severity_data$px_eyes=="yes")]<-0
  severity_data$avpu[which(severity_data$px_eyes=="no"&severity_data$px_eyes_no=="verbal")]<-1
  severity_data$avpu[which(severity_data$px_eyes=="no"&severity_data$px_eyes_no=="pain")]<-2
  severity_data$avpu[which(severity_data$px_eyes=="no"&severity_data$px_eyes_no=="not_at_all")]<-3
  severity_data$temperature<-ifelse(!is.na(severity_data$ear_temperature),severity_data$ear_temperature,severity_data$axil_temperature)
  severity_data$temperature[which(severity_data$temperature==999)]<-NA
  
  severity_data$mews<-0
  severity_data$mews[which(is.na(severity_data$avpu)|is.na(severity_data$bp_systolic)|is.na(severity_data$heart_rate)|is.na(severity_data$respiratory_rate)|is.na(severity_data$temperature))]<-NA
  severity_data$mews[which(severity_data$bp_systolic<=100)]<-severity_data$mews[which(severity_data$bp_systolic<=100)]+1
  severity_data$mews[which(severity_data$bp_systolic<=80)]<-severity_data$mews[which(severity_data$bp_systolic<=80)]+1
  severity_data$mews[which(severity_data$bp_systolic<=70)]<-severity_data$mews[which(severity_data$bp_systolic<=70)]+1
  severity_data$mews[which(severity_data$bp_systolic>=200)]<-severity_data$mews[which(severity_data$bp_systolic>=200)]+2
  severity_data$mews[which(severity_data$heart_rate<=50)]<-severity_data$mews[which(severity_data$heart_rate<=50)]+1
  severity_data$mews[which(severity_data$heart_rate<=40)]<-severity_data$mews[which(severity_data$heart_rate<=40)]+1
  severity_data$mews[which(severity_data$heart_rate>100)]<-severity_data$mews[which(severity_data$heart_rate>100)]+1
  severity_data$mews[which(severity_data$heart_rate>110)]<-severity_data$mews[which(severity_data$heart_rate>110)]+1
  severity_data$mews[which(severity_data$heart_rate>=130)]<-severity_data$mews[which(severity_data$heart_rate>=130)]+1
  severity_data$mews[which(severity_data$respiratory_rate<9)]<-severity_data$mews[which(severity_data$respiratory_rate<9)]+2
  severity_data$mews[which(severity_data$respiratory_rate>14)]<-severity_data$mews[which(severity_data$respiratory_rate>14)]+1
  severity_data$mews[which(severity_data$respiratory_rate>20)]<-severity_data$mews[which(severity_data$respiratory_rate>20)]+1
  severity_data$mews[which(severity_data$respiratory_rate>29)]<-severity_data$mews[which(severity_data$respiratory_rate>29)]+1
  severity_data$mews[which(severity_data$temperature<35.0)]<-severity_data$mews[which(severity_data$temperature<35.0)]+2
  severity_data$mews[which(severity_data$temperature>=38.5)]<-severity_data$mews[which(severity_data$temperature>=38.5)]+2
  severity_data$mews<-severity_data$mews+severity_data$avpu
  
  return((severity_data$mews))
}


# Modified GCS Score ----


modGCS<-function(severity_data){
  severity_data$px_eyes_no<-fct_collapse(severity_data$px_eyes_no,verbal=c("pain verbal_request","verbal_request"))
  severity_data$px_assess_speech<-fct_collapse(severity_data$px_assess_speech,confused=c("confused","confused inap_words non_sense_sounds","confused inap_words not_responding_verbal","confused not_responding_verbal inap_words","confused not_responding_verbal non_sense_sounds","not_responding_verbal confused","inap_words confused"))
  severity_data$oriented<-ifelse(severity_data$px_orient=="yes",5,
                                   ifelse(severity_data$px_orient=="no",ifelse(severity_data$px_assess_speech=="confused",4,
                                                                                 ifelse(severity_data$px_assess_speech=="inap_words",3,
                                                                                        ifelse(severity_data$px_assess_speech=="non_sense_sounds",2,
                                                                                               ifelse(severity_data$px_assess_speech=="not_responding_verbal",1,NA)))),
                                          NA))
  
  severity_data$eyes<-ifelse(severity_data$px_eyes=="yes",4,
                               ifelse(severity_data$px_eyes=="no",ifelse(severity_data$px_eyes_no=="verbal",3,
                                                                           ifelse(severity_data$px_eyes_no=="pain",2,
                                                                                  ifelse(severity_data$px_eyes_no=="not_at_all",1,
                                                                                         NA))),
                                      NA))
  
  severity_data$move<-ifelse(severity_data$px_coop=="yes",6,
                               ifelse(severity_data$px_coop=="no",ifelse(severity_data$px_coop_no=="painful_stim",3,
                                                                           ifelse(severity_data$px_coop_no=="not_at_all",1,
                                                                                  NA)),
                                      NA))
  
  severity_data<-severity_data%>%mutate(modGCS=oriented+eyes+move)
  return((severity_data$modGCS))
}




```



```{r assign_scores , include = F, message = F, warning = F, paged.print = F, eval = T}

# Apply severity score functions to dataset


severity_data$crb_65 <- crb_65(severity_data)

severity_data$qsofa <- qsofa(severity_data)

severity_data <- severity_data %>%
  mutate(qsofa = case_when(
    bp_systolic == 999 ~ NA_real_,
    respiratory_rate == 999 ~ NA_real_ ,
    TRUE ~ qsofa
  )
  )

severity_data$modGCS <- modGCS(severity_data)

severity_data$mews <- mews(severity_data)


severity_data <- severity_data %>%
  mutate(mews = case_when(
    heart_rate == 999 ~ NA_real_,
    respiratory_rate == 999 ~ NA_real_ ,
    bp_systolic == 999 ~ NA_real_,
    o2_sats == 999 ~ NA_real_,
    TRUE ~ mews
  )
  )

severity_data$uva <- uva(severity_data)


severity_data <- severity_data %>%
  mutate(uva = case_when(
    heart_rate == 999 ~ NA_real_,
    respiratory_rate == 999 ~ NA_real_ ,
    bp_systolic == 999 ~ NA_real_,
    o2_sats == 999 ~ NA_real_,
    hiv_comb == "Don't know" ~ NA_real_,
    hiv_comb == "indet" ~ NA_real_,
    is.na(as.character(hiv_comb)) ~ NA_real_,
        TRUE ~ uva
  )
  )

severity_data$uva_age <- uva_age(severity_data)


severity_data <- severity_data %>%
  mutate(uva_age = case_when(
    heart_rate == 999 ~ NA_real_,
    respiratory_rate == 999 ~ NA_real_,
    bp_systolic == 999 ~ NA_real_,
    o2_sats == 999 ~ NA_real_,
    hiv_comb == "Don't know" ~ NA_real_,
    hiv_comb == "indet" ~ NA_real_,
    is.na(as.character(hiv_comb)) ~ NA_real_,
    TRUE ~ uva_age
  ))


severity_data$uva_age <- uva_age(severity_data)


# MEWs & qSOFA HIV

severity_data <- severity_data %>%
  mutate(qsofa = as.numeric(as.character(qsofa)),
         hiv_comb = as.character(hiv_comb),
         mews = as.numeric(as.character(mews))
         ) %>%
  mutate(qsofa_hiv = ifelse(as.character(hiv_comb) == "Positive", qsofa+1,  qsofa),
         mews_hiv = ifelse(as.character(hiv_comb) == "Positive", mews+2,  mews)
         ) 


```

```{r clean_data , include = F, message = F, warning = F, paged.print = F, eval = T}

# Clean data and define health outcomes

# alive comp recov
# lost to fup
# dead
# alive improved
# alive worse day0
# alive same day0
# alive - unknown
# alive no info

severity_data <- severity_data %>%
  mutate(across(c(temp, heart_rate, bp_systolic, bp_diastolic, respiratory_rate, modGCS, qsofa, uva, uva_age, crb_65, mews, o2_sats, age, qsofa_hiv, mews_hiv), as.numeric)) %>%
   mutate(cf_9_px_status_final = str_replace_all(cf_9_px_status_final, "[^[:alnum:]]", " "),
          cf_5_inpatient_outcome = str_replace_all(cf_5_inpatient_outcome, "[^[:alnum:]]", " "),
          cf_4_discharge_outcome = str_replace_all(cf_4_discharge_outcome, "[^[:alnum:]]", " "),
          discharge_outcome = str_replace_all(discharge_outcome, "[^[:alnum:]]", " ")
          )  %>%
  mutate(outcome_discharge = case_when(
           !is.na(cf_4_discharge_outcome) & px_group == "outpatient - adult" ~ cf_4_discharge_outcome,
           !is.na(discharge_outcome) & px_group == "outpatient - adult" ~ discharge_outcome,
           !is.na(cf_5_inpatient_outcome) & px_group == "inpatient - adult" ~ cf_5_inpatient_outcome,
           TRUE ~ NA_character_),
         cf_9_date_death_fmt = as.Date((lubridate::parse_date_time(cf_9_date_death, c("dmY","dmy", "Ymd")))),
         date_day0_visit = lubridate::dmy(date_day0_visit)
         ) %>%
  mutate(cf_9_date_death_fmt = if_else(cf_9_date_death_fmt < date_day0_visit, lubridate::ymd(NA), cf_9_date_death_fmt),
         cf_5_discharge_date = lubridate::dmy(cf_5_discharge_date),
         time_to_discharge = as.numeric(cf_5_discharge_date - date_day0_visit)
         ) %>%
  mutate(time_to_discharge = ifelse(time_to_discharge >90 | time_to_discharge <0, NA, time_to_discharge)) %>%
    mutate(time_to_death = as.numeric(cf_9_date_death_fmt - date_day0_visit)) %>%
  mutate(time_to_death = ifelse(time_to_death >99, NA, time_to_death )) %>%
  mutate(outcome_discharge = case_when(
    outcome_discharge == "discharge home" ~ "Discharge home",
    outcome_discharge == "discharged" ~ "Discharge home",
    outcome_discharge == "discharged pallative" ~ "Discharged to palliative care",
    outcome_discharge == "other health facility" ~ "Referred other",
    outcome_discharge == "referred other" ~ "Referred other",
    outcome_discharge == "referred other" ~ "Referred other",
    outcome_discharge == "discharge home referred other" ~ "Referred other",
    outcome_discharge == "referred other discharge home" ~ "Referred other",
    outcome_discharge == "other health facility" ~ "Referred other",
    outcome_discharge == "other" ~ "Other",
    outcome_discharge == "other" ~ "Other",
    outcome_discharge == "died" ~ "Died",
    TRUE ~ outcome_discharge)
    ) %>%
  mutate(hiv_test_final = ifelse(hiv_test_final == "Indet", "Indeterminant", hiv_test_final)) %>%
  mutate(outcome_d28 = case_when(cf_9_px_status_final == "alive comp recov" ~ "Alive", 
                               cf_9_px_status_final == "alive improved" ~ "Alive",
                               cf_9_px_status_final == "alive no info" ~ "Alive",
                               cf_9_px_status_final == "alive same day0" ~ "Alive",
                               cf_9_px_status_final == "alive worse day0" ~ "Alive",
                               cf_9_px_status_final == "alive - unknown" ~ "Alive",
                               cf_9_px_status_final == "alive   unknown" ~ "Alive",
                               cf_9_px_status_final == "dead" ~ "Dead", 
                               is.na(cf_9_px_status_final) ~ "lost to fup",
                               TRUE ~ cf_9_px_status_final
                               ), 
        outcome_inpatient = case_when(cf_5_inpatient_outcome == "discharged" ~ "Alive", 
                               cf_5_inpatient_outcome == "discharged pallative" ~ "Alive",
                               cf_5_inpatient_outcome == "other" ~ "Alive",
                               cf_5_inpatient_outcome == "died" ~ "Dead",
                               cf_5_inpatient_outcome == "dead" ~ "Dead", 
                               TRUE ~ NA_character_),
        outcome_all = case_when(
          outcome_d28 == "Alive" ~ "Alive",
          outcome_d28 == "Dead" ~ "Dead",
          outcome_inpatient == "Alive" ~ "Alive",
          outcome_inpatient == "Dead" ~ "Dead",
          TRUE  ~ NA_character_
        )
        ) %>%
  mutate(outcome_d28_label = case_when(cf_9_px_status_final == "alive comp recov" ~ "Alive and completely recovered", 
                               cf_9_px_status_final == "alive improved" ~ "Alive and improved but incompletely recovered",
                               cf_9_px_status_final == "alive no info" ~ "Alive and no other information",
                               cf_9_px_status_final == "alive - unknown" ~ "Alive and no other information",
                               cf_9_px_status_final == "alive   unknown" ~ "Alive and no other information",
                               cf_9_px_status_final == "alive same day0" ~ "Alive and same as enrolment",
                               cf_9_px_status_final == "alive worse day0" ~ "Alive and worse than enrolment", 
                               cf_9_px_status_final == "dead" ~ "Dead", 
                               is.na(cf_9_px_status_final)  ~ "lost to fup",
                               TRUE ~ cf_9_px_status_final)
  )

# Altered mental status and age groups

severity_data <- severity_data %>%
  mutate(altered_mental_status = case_when(
           modGCS < 15 ~ "Yes",
           modGCS >= 15 ~ "No",
           TRUE ~ NA_character_
         ),
         age_cat = case_when(
           age >= 15 & age < 25 ~ "15 - <25",
           age >= 25 & age < 35 ~ "25 - <35",
           age >= 35 & age < 45 ~ "35 - <40",
           age >= 45 & age < 55 ~ "45 - <55",
           age >= 55 & age < 65 ~ "55 - <65",
           age >= 65 ~ "65+"
         ),
         hiv_comb = str_replace(hiv_comb, "Don't know", "Unknown"),
         hiv_comb = str_replace(hiv_comb, "indet", "Indeterminant")
         ) %>%
    mutate(sex = str_to_sentence(sex), 
           px_group = str_to_sentence(str_remove(px_group, "-adult")),
           hiv_comb = str_to_sentence(hiv_comb)
           ) 



severity_data <- severity_data %>%
  mutate(across(c(country, age_cat, sex, hiv_comb, cf_5_inpatient_outcome, cf_9_px_status_final, px_group, px_orient, outcome_d28, outcome_discharge, outcome_all, outcome_d28_label, altered_mental_status), as.factor)) %>%
  mutate(across(c(respiratory_rate, bp_systolic, bp_diastolic, heart_rate, o2_sats, temp), ~ifelse(.x >= 999, NA, .x)),
         mews = if_else(mews == 0, 1, mews)) # set those with a MEWS of 0 to 1, currently only 2 patients have  MEWs of 0. 



relev_yndk <- function(f) forcats::fct_relevel(factor(f), "Yes", "No", "Don't know")
relev_yn <- function(f) forcats::fct_relevel(factor(f), "Yes", "No")
relev_pos_neg <- function(f) forcats::fct_relevel(factor(f), "Positive", "Negative", "Indeterminant", "Unknown")

severity_data$altered_mental_status <- relev_yn(severity_data$altered_mental_status)

severity_data$hiv_comb <- relev_pos_neg(severity_data$hiv_comb)


# Label dataset variables

severity_data <- severity_data %>%
  sjlabelled::var_labels(
                age = "Age (years)",
                age_cat = "Age group",
                sex = "Gender",
                px_group = "Participant group",
                temp = "Temperature (°C)",
                hiv_comb = "HIV status",
                respiratory_rate = "Respiratory rate (breaths/min)",
                bp_systolic = "Systolic blood pressure, mm Hg",
                bp_diastolic = "Diastolic blood pressure, mm Hg",
                heart_rate = "Heart rate, beats/min",
                o2_sats = "Oxygen saturation (%)",
                mews = "Modified early warning score (MEWS)",
                crb_65 = "CRB-65",
                modGCS = "Modified GCS",
                qsofa = "qSOFA",
                uva = "UVA",
                altered_mental_status = "Altered mental status \n (Glasgow Coma Scale <15)",
                outcome_inpatient = "Inpatient outcome",
                outcome_d28 = "Day 28 outcome",
                outcome_all = "Outcomes",
                px_group = "Patient group",
                cf_9_px_status_final = "Day 28 outcome",
                cf_5_inpatient_outcome = "Inpatient discharge outcome",
                outcome_d28_label = "Day 28 outcome",
                outcome_discharge = "In-hospital outcomes",
                px_eyes = "While the participant is awake, do they open their eyes spontaneously?",
                time_to_death = "Time to death (days)",
                time_to_discharge = "Length of inpatient stay (days)"
) 




```

```{r data_summary, include = F, message = F, warning = F, paged.print = F, eval = T}

summary((severity_data$mews))

summary(as.factor(severity_data$outcome_d28))

summary(as.factor(severity_data$outcome_inpatient))

tabyl(severity_data, outcome_inpatient, outcome_d28) %>%
  adorn_totals(where = "row")

tabyl(severity_data, px_group, outcome_d28) %>%
  adorn_totals(where = "row")

summary(as.factor(severity_data$outcome_all))


summary(as.factor(severity_data$cf_5_inpatient_outcome))
summary(as.factor(severity_data$cf_9_px_status_final))
summary(as.factor(severity_data$px_group))

summary(as.factor(severity_data$outp_switched))


summary(as.factor(severity_data$px_group))
summary(as.factor(severity_data$country))
summary(as.factor(severity_data$outcome_d28))
summary(as.factor(severity_data$uva))


#  1 - Day28 outcomes ----

total_px <- nrow(severity_data)

severity_data_v1 <- severity_data %>%
  mutate(outcome_d28 = as.character(outcome_d28)) %>%
  filter(!is.na(outcome_d28)) %>%
  filter(outcome_d28 != "lost to fup")


severity_data_cc_d28 <- severity_data_v1 %>%
    filter(if_all(c(modGCS, qsofa, uva, mews), ~ !is.na(.))) 

severity_data_cc_d28 %>%
  select(uva, hiv_comb) %>%
  filter(!is.na(hiv_comb))

severity_data_v1 %>%
  select(qsofa, uva, mews, hiv_comb) %>%
  filter(is.na(hiv_comb))


tabyl(severity_data, outcome_inpatient, outcome_d28) %>%
  adorn_totals(where = "row")


severity_data_no_outcome_d28 <- severity_data %>%
  filter(is.na(outcome_d28)) %>%
  select(patient_id_calculate) %>%
  mutate(no_outcome_d28 = 1) 

px_no_outcome <- severity_data_no_outcome_d28 %>%
  nrow()

no_sev_score <- left_join(
  severity_data,
  severity_data_no_outcome_d28,
  by = "patient_id_calculate"
)


px_no_sev_score <- no_sev_score %>%
  filter(is.na(no_outcome_d28)) %>%
  filter(as.character(outcome_d28) != "lost to fup") %>%
  # select(modGCS, qsofa, uva, mews) %>%
  filter(if_any(c(modGCS, qsofa, uva, mews), ~ is.na(.))) %>%
  nrow()

severity_data_cc_d28 <- no_sev_score %>%
  filter(is.na(no_outcome_d28)) %>%
    filter(as.character(outcome_d28) != "lost to fup") %>%
  # select(modGCS, qsofa, uva, mews, no_outcome) %>%
  filter(if_all(c(modGCS, qsofa, uva, mews), ~ !is.na(.))) %>%
  droplevels()


total_px - px_no_outcome - px_no_sev_score

# LTFU

ltfu <- severity_data %>%
  select(cf_9_px_status_final) %>%
  filter(cf_9_px_status_final == "lost to fup") %>%
  nrow()

missing_day28 <- severity_data %>%
  select(cf_9_px_status_final) %>%
  filter(is.na(cf_9_px_status_final)) %>%
  nrow()

miss_outcome <- severity_data %>%
  filter(is.na(cf_9_px_status_final))

miss_outcome_ltfu <- severity_data %>%
  filter(cf_9_px_status_final == "lost to fup")


miss_scores <- severity_data %>%
  filter(!is.na(as.character(outcome_d28))) %>%
  filter(if_any(c(modGCS, qsofa, uva, mews), ~ is.na(.))) %>%
  select(patient_id_calculate, modGCS, qsofa, uva, mews)



(total_px - nrow(miss_outcome)) - nrow(miss_outcome_ltfu) - nrow(miss_scores)

(max(severity_data$date_day0_visit)- min(severity_data$date_day0_visit))/30.437


sev_scores <- severity_data %>%
  select(patient_id_calculate, uva, mews, qsofa, modGCS, ends_with("_hiv")) 

# rio::export(sev_scores, "/Users/sham/Filr/My Files/sync/My Files/fiebre/data/analysis_scripts/adult_severity_scores/data/fiebre_adult_severity_scores_2023_07_27.csv")

summary(as.factor(severity_data_cc_d28$outp_switched))
# severity_data$outp_switched


```
   
   
   
Table of Contents   
   
<!---BLOCK_TOC--->   

<!---BLOCK_TOC{seq_id: 'tab'}--->   

<!---BLOCK_TOC{seq_id: 'fig'}--->   
   


\newpage

<!---BLOCK_LANDSCAPE_START--->


## Participants recruited

* These are patient outcomes reported at day 28 follow-up. 
* Patient group (whether a patient is an outpatient or an inpatient) is defined at recruitment.   

```{r eval=T, echo=F, message=FALSE, warning=FALSE, include=T, paged.print=FALSE, results='asis', tab.cap="Day 28 follow-up outcomes"}

tbl_outcome_d28 <-
  severity_data  %>%
    select(px_group, country, outcome_d28_label, outcome_d28) %>%
    tbl_summary(
    by = country,
    type = list(all_dichotomous() ~ "categorical",
                all_continuous() ~ "continuous2"
                ),
    statistic = list(c("px_group",  "outcome_d28_label") ~ "{n} ({p}%)"),
    digits = all_continuous() ~ 1,
    missing_text = "(Missing)"
      ) %>%
      add_stat_label() %>%
      add_overall() %>%
     modify_header(list(label ~"Variable")) %>%
     as_flex_table() %>%
      flextable::fontsize(part = "header", size = 8) %>%
      flextable::autofit()

tbl_outcome_d28

```



\newpage

* These are in-hospital patient outcomes. 

```{r, tab.cap = "In-hospital outcomes",  include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T }

tbl_outcome_d0 <-
  severity_data  %>%
    select(px_group, country, outcome_discharge) %>%
    tbl_summary(
    by = country,
    type = list(all_dichotomous() ~ "categorical",
                all_continuous() ~ "continuous2"
                ),
    statistic = list(c("px_group",  "outcome_discharge") ~ "{n} ({p}%)"),
    digits = all_continuous() ~ 1,
    missing_text = "(Missing)"
      ) %>%
      add_stat_label() %>%
      add_overall() %>%
     modify_header(list(label ~"Variable")) %>%
     as_flex_table() %>%
      flextable::fontsize(part = "header", size = 8) %>%
      flextable::autofit()

tbl_outcome_d0

```

```{r, tab.cap = "In-hospital outcomes by patient group",  include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T }

tbl_outcome_d0_px_group <- severity_data %>%
  tbl_cross(row = px_group, col = outcome_discharge, percent = "row") %>%
   modify_header(list(label ~ "")) %>%
  as_flex_table() %>%
  flextable::fontsize(part = "header", size = 8)

tbl_outcome_d0_px_group %>%
  flextable::width(j = 1, width = 1) %>%
  flextable::width(j=c(2,3,4,5,6,7,8), width = 0.7) 

```
\newpage

Due to a small number of in-hospital deaths, this analysis is limited to outcomes at day 28.    
* Total number of recruited adult patients: __`r total_px`__  
* Number of patients _missing_ self-reported outcome (dead or alive) data at day 28 follow-up: __`r px_no_outcome`__.   
    -   These were patients loss to follow-up at day 28 (__`r ltfu`__) and where no outcome data were reported at day 28 (__`r missing_day28`__).   
* Number _missing_ severity score data: __`r px_no_sev_score`__.      
* Patients with _complete_ self-reported outcome data at day 28 and _complete_ clinical data (complete cases): __`r total_px - px_no_outcome - px_no_sev_score`__   
     
     

```{r, tab.cap = "Day 28 outcome by patient group",  include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T }

tbl_outcome_d28_px_group <-
  severity_data_cc_d28 %>%
  tbl_cross(row = px_group, col = outcome_d28, percent = "row") %>%
   modify_header(list(label ~ "")) %>%
  as_flex_table() %>%
  flextable::fontsize(part = "header", size = 8)

tbl_outcome_d28_px_group

```

<!---BLOCK_LANDSCAPE_STOP--->



## Descriptive results tables

```{r tab.cap = "Demographic and clinical characteristics of patients with complete data at day 28", tab.id = "tbl_1a", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T}


tbl_1a_d28 <- severity_data_cc_d28 %>%
  select(age, age_cat, sex, px_group, time_to_discharge, temp, hiv_comb, respiratory_rate, bp_systolic, heart_rate, o2_sats, mews, modGCS, qsofa, uva, altered_mental_status, outcome_discharge) %>%
  tbl_summary(
    # by = country,
    type = list(all_dichotomous() ~ "categorical",
                all_continuous() ~ "continuous2"
                ),
    statistic = list(c("age", "temp", "time_to_discharge", "respiratory_rate", "bp_systolic", "heart_rate", "mews" ,"o2_sats", "modGCS", "uva") ~ 
                     c("{median} ({p25}, {p75})", "{min}, {max}"),
                     c("px_group",  "hiv_comb", "altered_mental_status", "outcome_discharge") ~ "{n} ({p}%)"),
    digits = all_continuous() ~ 1,
    missing_text = "(Missing)"
      ) %>%
    add_stat_label() %>%
     modify_header(list(label ~"Variable")) %>%
    as_flex_table() %>%
      flextable::fontsize(part = "header", size = 8) %>%
      flextable::autofit()

tbl_1a_d28


```



<!---BLOCK_LANDSCAPE_START--->

### Table 2 
```{r tab.cap="Demographic and clinical characteristics of patients with complete data at day 28 follow-up, by site", eval=T, echo=F, message=FALSE, warning=FALSE, include=T, paged.print=FALSE, results='asis'}


tbl_1a_d28 <- severity_data_cc_d28 %>%
  select(country, age, age_cat, sex, px_group, time_to_discharge, temp, hiv_comb, respiratory_rate, bp_systolic, heart_rate, o2_sats, mews, modGCS, qsofa, uva, altered_mental_status, outcome_discharge) %>%
  tbl_summary(
    by = country,
    type = list(all_dichotomous() ~ "categorical",
                all_continuous() ~ "continuous2"
                ),
    statistic = list(c("age", "temp", "time_to_discharge" , "respiratory_rate", "bp_systolic", "heart_rate", "mews" ,"o2_sats", "modGCS", "uva") ~ 
                     c("{median} ({p25}, {p75})", "{min}, {max}"),
                     c("px_group", "age_cat", "hiv_comb", "altered_mental_status", "outcome_discharge") ~ "{n} ({p}%)"),
    digits = all_continuous() ~ 1,
    missing_text = "(Missing)",
      ) %>%
      add_overall(last = T) %>%
    add_stat_label() %>%
     modify_header(list(label ~"Variable")) %>%
    as_flex_table() %>%
      flextable::fontsize(part = "header", size = 8) %>%
      flextable::autofit()

tbl_1a_d28

```

\newpage
### Table 3 
```{r tab.cap = "Demographic and clinical characteristics of patients by day 28 outcome", tab.id = "tbl_2a", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T}


tbl_2a_d28 <- severity_data_cc_d28 %>%
  select(outcome_d28, age, age_cat, sex, px_group, time_to_discharge, country, temp, hiv_comb, respiratory_rate, bp_systolic, heart_rate, o2_sats, mews, modGCS, qsofa, uva, altered_mental_status, time_to_death, outcome_discharge) %>%
  tbl_summary(
    by = outcome_d28,
    percent = "row",
    type = list(all_dichotomous() ~ "categorical",
                all_continuous() ~ "continuous2"
                ),
    statistic = list(c("age", "temp", "time_to_discharge", "respiratory_rate", "bp_systolic", "heart_rate", "mews" ,"o2_sats", "modGCS", "uva" ) ~ 
                     c("{N_nonmiss}", "{median} ({p25}, {p75})", "{min}, {max}"),
                     c("px_group", "age_cat", "hiv_comb", "altered_mental_status", "outcome_discharge") ~ "{n} ({p}%)"),
    digits = all_continuous() ~ 1,
    missing_text = "(Missing)"
      ) %>%
    add_overall(last = T) %>%
    add_stat_label() %>%
     modify_header(list(label ~"Variable")) %>%
    as_flex_table() %>%
      flextable::fontsize(part = "header", size = 8) %>%
      flextable::autofit()

tbl_2a_d28





# Increasing age and mortality trend

tbl3_age_death <- severity_data_cc_d28 %>% 
  janitor::tabyl(age_cat, outcome_d28) %>% 
  select(-1) %>%
  rstatix::prop_trend_test()

# by site
  # tbl3_age_death <- severity_data_cc_d28 %>% 
  #   filter(country == "Laos") %>%
  # janitor::tabyl(age_cat, outcome_all) %>% 
  # select(-1) %>%
  # rstatix::prop_trend_test()
  # 
  #   tbl3_age_death <- severity_data_cc_d28 %>% 
  #   filter(country == "Malawi") %>%
  # janitor::tabyl(age_cat, outcome_all) %>% 
  # select(-1) %>%
  # rstatix::prop_trend_test()
  #   
  #   severity_data_cc_d28 %>% 
  #   filter(country == "Mozambique") %>%
  # janitor::tabyl(age_cat, outcome_all) %>% 
  # select(-1) %>%
  # rstatix::prop_trend_test()
  # 
  #       severity_data_cc_d28 %>% 
  #   filter(country == "Zimbabwe") %>%
  # janitor::tabyl(age_cat, outcome_all) %>% 
  # select(-1) %>%
  # rstatix::prop_trend_test()

# Mortality and sex
tbl3_sex_death <- severity_data_cc_d28 %>% 
  janitor::tabyl(sex, outcome_d28) %>%
  select(sex, Dead, Alive) %>%
  select(-1) %>%
  rstatix::prop_test(detailed = TRUE)

# Mortality and inpatients

severity_data_cc_d28 %>% 
  janitor::tabyl(px_group, outcome_d28) %>%
  select(px_group, Dead, Alive) %>%
  select(-1) %>%
  rstatix::prop_test(detailed = TRUE)


# Mortality and HIV

severity_data_cc_d28 %>% 
  janitor::tabyl(hiv_comb, outcome_d28) %>%
  # filter(hiv_comb == "Positive" | hiv_comb == "Negative") %>%
  select(hiv_comb, Dead, Alive) %>%
  select(-1) %>%
  rstatix::prop_test(detailed = TRUE)


```

```{r tab.cap = "Severity scores by site", tab.id = "tbl_2b", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T}


tbl_2b <- severity_data_cc_d28 %>%
  select(country, qsofa, mews, uva) %>%
  tbl_summary(
    by = country,
    type = list(all_dichotomous() ~ "categorical",
                all_continuous() ~ "continuous2"
                ),
    statistic = list(c("uva", "mews" ) ~ 
                     c("{median} ({p25}, {p75})", "{min}, {max}"),
                     c("qsofa") ~ "{n} ({p}%)"),
    digits = all_continuous() ~ 1,
    missing_text = "(Missing)"
      ) %>%
    add_stat_label() %>%
     modify_header(list(label ~"Variable")) %>%
    as_flex_table() %>%
      flextable::fontsize(part = "header", size = 8) %>%
      flextable::autofit()

tbl_2b


```


```{r fig_time_death, fig.cap = "Time (days) to death by day 28 by patient group", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T}


time2death_px_tbl <- severity_data_cc_d28 %>%
  filter(outcome_d28 == "Dead") %>%
  select(px_group, time_to_death) %>%
  tbl_summary(
    by = px_group,
    type = list(all_dichotomous() ~ "categorical",
                all_continuous() ~ "continuous2"
                ),
    statistic = list(c("time_to_death") ~ 
                     c("{median} ({p25}, {p75})", "{min}, {max}")
    ),
    digits = all_continuous() ~ 1,
    missing_text = "(Missing)"
      ) %>%
    add_overall(last = T) %>%
    add_stat_label() %>%
  add_p(time_to_death ~ "wilcox.test")  %>%
     modify_header(list(label ~"Variable")) %>%
    as_flex_table() %>%
      flextable::fontsize(part = "header", size = 8) %>%
      flextable::autofit()


time2death_country_tbl <- severity_data_cc_d28 %>%
  filter(outcome_d28 == "Dead") %>%
  select(country, time_to_death) %>%
  tbl_summary(
    by = country,
    type = list(all_dichotomous() ~ "categorical",
                all_continuous() ~ "continuous2"
                ),
    statistic = list(c("time_to_death") ~ 
                     c("{median} ({p25}, {p75})", "{min}, {max}")
    ),
    digits = all_continuous() ~ 1,
    missing_text = "(Missing)"
      ) %>%
    add_overall() %>%
    add_stat_label() %>%
     modify_header(list(label ~"Variable")) %>%
    as_flex_table() %>%
      flextable::fontsize(part = "header", size = 8) %>%
      flextable::autofit()


time2death_px_fig <- ggplot(data = severity_data_cc_d28 %>% 
                              select(outcome_d28, px_group, time_to_death) %>%
                              filter(outcome_d28 == "Dead"), 
                        mapping = aes(y = time_to_death, x = px_group, fill = px_group)) + 
  geom_boxplot() + 
  scale_x_discrete(name = "Time to death (days)") +
  scale_y_continuous(breaks = seq(0, 90, 10), limits = c(0, 90), name = "Days") +
  theme_minimal() +
  theme(legend.position = "none") +
  labs(title = "")    

time2death_px_fig


  
```


<!---BLOCK_LANDSCAPE_START--->

## Demographic and clinical characteristics of patients with incomplete data
### Table S1
\newpage

```{r, tab.cap = "Demographic and clinical characteristics of patients with incomplete data at day 28, by site", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T }


comp_data <- severity_data_cc_d28 %>%
  mutate(has_comp_data = 1) %>%
  select(patient_id_calculate, has_comp_data)

incomp_data <- left_join(
  severity_data,
  comp_data,
  by = "patient_id_calculate"
) %>%
  filter(is.na(has_comp_data))

tbl_incomp_data <- incomp_data %>%
  select(country, age, age_cat, sex, px_group, time_to_discharge, temp, hiv_comb, respiratory_rate, bp_systolic, heart_rate, o2_sats, mews, crb_65, modGCS, qsofa, uva, altered_mental_status, outcome_discharge, outcome_d28) %>%
  mutate(mews = as.numeric(as.character(mews)), 
         qsofa = as.numeric(as.character(qsofa)),
         uva = as.numeric(as.character(uva))
         ) 
#   
# tbl_incomp_data %>%
#   group_by(mews) %>%
#   tally()

tbl_incomp_data <- tbl_incomp_data %>%  
  tbl_summary( 
    by = country,
    type = list(c("altered_mental_status") ~ "categorical",
                c("age", "temp", "time_to_discharge", "respiratory_rate", "bp_systolic", "heart_rate", "mews" ,"o2_sats", "modGCS", "uva") ~ 'continuous2'),
 statistic = list(c("age", "temp", "time_to_discharge", "respiratory_rate", "bp_systolic", "heart_rate", "mews" ,"o2_sats", "modGCS", "uva" ) ~ 
                     c("{N_nonmiss}", "{median} ({p25}, {p75})", "{min}, {max}"),
                     c("px_group", "age_cat", "hiv_comb", "altered_mental_status", "outcome_discharge") ~ "{n} ({p}%)"),
    digits = all_continuous() ~ 1,
    missing_text = "(Missing)"
      ) %>%
  add_stat_label() %>%
  modify_header(list(label ~"Variable")) %>%
  add_overall(last = T) %>%
    as_flex_table() %>%
      flextable::fontsize(part = "header", size = 8) %>%
      flextable::autofit()

tbl_incomp_data


```

<!---BLOCK_LANDSCAPE_STOP--->

\newpage
## Day 28 by outcome by site
### Table S2 
```{r tab.cap = "Demographic and clinical characteristics of patients by day 28 outcome, by site", tab.id = "tbl_2a", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T}


tbl_2a_d28_site <- severity_data_cc_d28 %>%
  select(outcome_d28, age_cat, sex, px_group, hiv_comb, outcome_d28, country) %>%
  tbl_strata(
        strata = country,
    .tbl_fun =
      ~ .x %>%
    tbl_summary(
      by = outcome_d28,
      percent = "row",
      missing_text = "(Missing)"
        ) %>%
      add_overall(last = T, statistic = ~"{n}") %>%
      add_stat_label()
  ) %>%
     modify_header(list(label ~"Variable")) %>%
    as_flex_table() %>%
      flextable::fontsize(part = "header", size = 8) %>%
      flextable::autofit()

tbl_2a_d28_site
```

### Table S3 
```{r tab.cap = "Time (days) to death by day 28, by patient group ", tab.id = "tbl_2a", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T}

time2death_px_tbl

```



\newpage
### Figure S1
```{r fig_time_death_country, fig.cap = "Time (days) to death by recruitment site", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T}


time2death_country_fig <- ggplot(data = severity_data_cc_d28 %>% 
                                   select(outcome_d28, country, time_to_death) %>%
                                   mutate(country = if_else(country == "Laos", "Lao PDR", country)),
                                   filter(outcome_d28 == "Dead"), 
                                 mapping = aes(y = time_to_death, x = country, fill = country)) + 
  geom_boxplot() + 
  scale_x_discrete(name = "FIEBRE Site") +
  scale_y_continuous(breaks = seq(0, 90, 10), limits = c(0, 90), name = "Time to death (days)") +
  theme_minimal() +
  theme(legend.position = "none") +
  labs(title = "")    

time2death_country_fig
```

```{r tab.cap = "Time (days) to death by day 28, by recruiting site", tab.id = "tbl_2a", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T}

time2death_country_tbl


```
\newpage
```{r fig1, fig.cap = "Age distribution by gender",  include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T }

# fig1 <- severity_data_cc_d28 %>%
#   ggplot(mapping = aes(x = age, color = sex)) +
#   geom_freqpoly(binwidth = 2, size = 1.5) 
# 
# fig1


age_sex <- severity_data_cc_d28 %>%
  ggplot(mapping = aes(x = age, fill = sex)) +
  geom_histogram( color='#e9ecef', alpha=0.8, bins = 9, boundary = 0, closed = "left", position = position_dodge2(preserve = "single", padding = 0.1)) +
  scale_x_continuous(breaks=seq(0, 100, 20)) +
  theme_classic()

age_sex

```



```{r age_gender_site, fig.cap = "Age distribution by gender and site",  include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T }

# fig2 <- severity_data_cc_d28 %>%
#   ggplot(mapping = aes(x = age, color = sex)) +
#   geom_freqpoly(binwidth = 2, size = 1.5) +
#   facet_wrap(vars(country))
# 
# fig2


age_sex_site <- severity_data_cc_d28 %>%
  ggplot(mapping = aes(x = age, fill = sex)) +
  geom_histogram(color='#e9ecef', alpha=0.8, bins = 9, boundary = 0, closed = "left", position = position_dodge2(preserve = "single", padding = 0.1)) +
  scale_x_continuous(breaks=seq(0, 100, 15)) +
  facet_wrap(vars(country)) 

age_sex_site

```



```{r fig_bar_mews, fig.cap = "Distribution of MEWs by Day 28 outcome", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T}

mews_tbl_dead <- severity_data_cc_d28 %>%
  count(mews, outcome_d28) %>%  
  group_by(mews) %>%        
  mutate(proportion_dead = n/sum(n)*100) %>%
  filter(outcome_d28 == "Dead")

mews_tbl_tot_p <- severity_data_cc_d28 %>%
  count(mews) %>%  
  mutate(proportion = n/sum(n)*100) 

mews_fig <- left_join(
  mews_tbl_tot_p,
  mews_tbl_dead,
  by = "mews"
) %>%
  mutate(proportion_dead = ifelse(is.na(proportion_dead), 0, proportion_dead)
  ) %>%
  select(mews, proportion, proportion_dead) %>%
  ungroup()


fig4 <- ggplot(data = mews_fig, 
       mapping = aes(x = mews)) + 
  geom_col(aes(y = proportion, fill = "Percent at MEW Score")) +
  geom_line(aes(y = proportion_dead, color = "Percent mortality")) +
  geom_point(aes(y = proportion_dead)) +
    scale_y_continuous(limits = c(0, 100), name = "Percent (%)") +
    scale_x_continuous(breaks = c(0,1,2,3,4,5,6,7,8,9,10,11),
                     labels  = c(0,1,2,3,4,5,6,7,8,9,10,11),
                     name = "Modified Early Warning Score (MEWS)") +
  scale_fill_manual(name = "", values = c("Percent at MEW Score" = "grey")) +
  scale_color_manual(name = "", values = c("Percent mortality" = "black")) +
  theme_bw()

fig4

```

```{r fig_bar_mews_a, fig.cap = "Distribution of MEWs by Day 28 outcome (grouped)", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T}

mews_tbl_dead_a <- severity_data_cc_d28 %>%
  mutate(mews_grouped = case_when(
    mews == 10 ~ 9,
    mews == 11 ~ 9,
    TRUE ~ mews
  ) ) %>%
  count(mews_grouped, outcome_d28) %>%  
  group_by(mews_grouped) %>%        
  mutate(proportion_dead = n/sum(n)*100) %>%
  filter(outcome_d28 == "Dead")

mews_tbl_tot_p_a <- severity_data_cc_d28 %>%
    mutate(mews_grouped = case_when(
    mews == 10 ~ 9,
    mews == 11 ~ 9,
    TRUE ~ mews
  ) ) %>%
  count(mews_grouped) %>%  
  mutate(proportion = n/sum(n)*100) 

mews_fig_a <- left_join(
  mews_tbl_tot_p_a,
  mews_tbl_dead_a,
  by = "mews_grouped"
) %>%
  mutate(proportion_dead = ifelse(is.na(proportion_dead), 0, proportion_dead)
  ) %>%
  select(mews_grouped, proportion, proportion_dead) %>%
  ungroup()


fig4a <- ggplot(data = mews_fig_a, 
       mapping = aes(x = mews_grouped)) + 
  geom_col(aes(y = proportion, fill = "Percent at MEWS")) +
  geom_line(aes(y = proportion_dead, color = "Percent mortality")) +
  geom_point(aes(y = proportion_dead)) +
    scale_y_continuous(limits = c(0, 100), name = "Percent (%)") +
    scale_x_continuous(breaks = c(0,1,2,3,4,5,6,7,8,9),
                     labels  = c(0,1,2,3,4,5,6,7,8,"9-11"),
                     name = "MEWS") +
  scale_fill_manual(name = "", values = c("Percent at MEWS" = "grey")) +
  scale_color_manual(name = "", values = c("Percent mortality" = "black")) +
  labs(caption = "") +
  theme_bw() +
  theme(legend.position = "bottom")

fig4a

```

```{r fig_scatter_mews, fig.cap = "Scatter plot of MEWs by Day 28 outcome", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = F}
  
  mews_tbl2 <- severity_data_cc_d28 %>%
    # count(mews, outcome_d28) %>%
    # group_by(outcome_d28) %>%
    mutate(mews = as.factor(mews))
  
  fig4b <- ggplot(data = mews_tbl2, 
                  aes(x = outcome_d28, y = mews)) +
      scale_y_discrete(name = "Modified Early Warning Score (MEWS)") +
      # geom_bar()
      scale_x_discrete(name = "Day 28 outcome") +
      geom_jitter() +
      theme_minimal()
  
  
  fig4b

```


```{r fig_bar_qsofa, fig.cap = "Distribution of qSOFA by Day 28 outcome", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T}


qsofa_tbl_dead <- severity_data_cc_d28 %>%
  count(qsofa, outcome_d28) %>%  
  group_by(qsofa) %>%        
  mutate(proportion_dead = n/sum(n)*100) %>%
    filter(outcome_d28 == "Dead")

qsofa_tbl_tot_p <- severity_data_cc_d28 %>%
  count(qsofa) %>%  
  mutate(proportion = n/sum(n)*100) 

qsofa_tbl <- left_join(
  qsofa_tbl_tot_p,
  qsofa_tbl_dead,
  by = "qsofa"
) %>%
  select(qsofa, proportion, proportion_dead)

fig5 <- ggplot(data = qsofa_tbl, 
       mapping = aes(x = qsofa)) + 
  geom_col(aes(y = proportion, fill = "Percent at qSOFA score")) +
  geom_line(aes(y = proportion_dead, color = "Percent mortality")) +
  geom_point(aes(y = proportion_dead)) +
    scale_y_continuous(limits = c(0, 100), name = "Percent (%)") +
    scale_x_continuous(breaks = c(0,1,2,3,4,5,6,7,8,9,10,11),
                     labels  = c(0,1,2,3,4,5,6,7,8,9,10,11),
                     name = "qSOFA score") +
  scale_fill_manual(name = "", values = c("Percent at qSOFA score" = "grey")) +
  scale_color_manual(name = "", values = c("Percent mortality" = "black")) +
  theme_bw() +
  theme(legend.position = "bottom")

fig5

```

```{r fig_scatter_qsofa, fig.cap = "Distribution of qSOFA by Day 28 outcome", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = F}

qsofa_tbl <- severity_data_cc_d28 %>%
  # count(mews, outcome_d28) %>%
  # group_by(outcome_d28) %>%
  mutate(qsofa = as.factor(qsofa))

fig5b <- ggplot(data = qsofa_tbl, 
                aes(x = outcome_d28, y = qsofa)) +
    scale_y_discrete(name = "qSOFA score") +
    # geom_bar()
    scale_x_discrete(name = "Day 28 outcome") +
    geom_jitter() +
    theme_minimal()

fig5b

```


```{r fig_bar_uva, fig.cap = "Distribution of UVA score by Day 28 outcome", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T}

uva_tbl_dead <- severity_data_cc_d28 %>%
  count(uva, outcome_d28) %>%  
  group_by(uva) %>%        
  mutate(proportion_dead = n/sum(n)*100) %>%
  filter(outcome_d28 == "Dead")

uva_tbl_tot_p <- severity_data_cc_d28 %>%
  count(uva) %>%  
  mutate(proportion = n/sum(n)*100) 

uva_tbl <- left_join(
  uva_tbl_tot_p,
  uva_tbl_dead,
  by = c("uva")
) %>%
  select(uva, proportion, proportion_dead)



fig6 <- ggplot(data = uva_tbl, 
       mapping = aes(x = uva)) + 
  geom_col(aes(y = proportion, fill = "Percent at UVA Score")) +
  geom_line(aes(y = proportion_dead, color = "Percent mortality")) +
  geom_point(aes(y = proportion_dead)) +
    scale_y_continuous(limits = c(0, 100), name = "Percent (%)") +
    scale_x_continuous(breaks = c(0,1,2,3,4,5,6,7,8,9,10,11, 12),
                     labels  = c(0,1,2,3,4,5,6,7,8,9,10,11,12),
                     name = "UVA score") +
  scale_fill_manual(name = "", values = c("Percent at UVA Score" = "grey")) +
  scale_color_manual(name = "", values = c("Percent mortality" = "black")) +
  # theme(legend.box.spacing = unit(0.1, "cm"),
  #       legend.key.size = unit(1, "cm"))
  theme_bw() +
  theme(legend.position = "bottom")

fig6


```



```{r fig_bar_uva_group, fig.cap = "Distribution of UVA score by Day 28 outcome (grouped)", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T}

summary(as.factor(severity_data_cc_d28$uva))

uva_tbl_dead_a <- severity_data_cc_d28 %>%
   mutate(uva_grouped = case_when(
    uva == 9 ~ 8,
    uva == 10 ~ 8,
    TRUE ~ uva
  ) ) %>%
  count(uva_grouped, outcome_d28) %>%  
  group_by(uva_grouped) %>%        
  mutate(proportion_dead = n/sum(n)*100) %>%
  filter(outcome_d28 == "Dead")

uva_tbl_tot_p_a <- severity_data_cc_d28 %>%
   mutate(uva_grouped = case_when(
    uva == 9 ~ 8,
    uva == 10 ~ 8,
    TRUE ~ uva
  ) ) %>%
  count(uva_grouped) %>%  
  mutate(proportion = n/sum(n)*100) 

uva_tbl_a <- left_join(
  uva_tbl_tot_p_a,
  uva_tbl_dead_a,
  by = c("uva_grouped")
) %>%
  select(uva_grouped, proportion, proportion_dead)



fig6a <- ggplot(data = uva_tbl_a, 
       mapping = aes(x = uva_grouped)) + 
      geom_col(aes(y = proportion, fill = "Percent at UVA score")) +
    geom_line(aes(y = proportion_dead, color = "Percent mortality")) +
  geom_point(aes(y = proportion_dead)) +

    scale_y_continuous(limits = c(0, 100), name = "Percent (%)") +
    scale_x_continuous(breaks = c(0,1,2,3,4,5,6,7,8),
                     labels  = c(0,1,2,3,4,5,6,7,"8-12"),
                     name = "UVA score") +
  scale_fill_manual(name = "", values = c("Percent at UVA score" = "grey")) +
  scale_color_manual(name = "", values = c("Percent mortality" = "black")) +

  labs(caption = "") +
  theme_bw() +
  theme(legend.position = "bottom") +
  guides(color = guide_legend(order = 1),
    fill = guide_legend(order = 0)) 

  fig6a



```


```{r fig_scatter_uva, fig.cap = "Scatter plot of UVA by Day 28 outcome", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = F}


uva_tbl <- severity_data_cc_d28 %>%
  # count(mews, outcome_d28) %>%
  # group_by(outcome_d28) %>%
  mutate(uva = as.factor(uva))

fig6b <- ggplot(data = uva_tbl, 
                aes(x = outcome_d28, y = uva)) +
    scale_y_discrete(name = "Universal Vital Assessment (UVA) score") +
    # geom_bar()
    scale_x_discrete(name = "Day 28 outcome") +
    geom_jitter() +
    theme_minimal()

fig6b


```

### Figure 2
```{r all_severity_scores_plot, fig.cap = "ROCs for 1) MEWs, 2) qSOFA, 3) UVA", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T, fig.width = 9, fig.height=6}


combined_scores <- cowplot::plot_grid(fig4a, fig5, fig6a,
                               labels = c('a', 'b', "c"), label_size = 12, align = "h"
                               )

combined_scores  




```



### Table S3
```{r tbl_bar_score_death_site, tab.cap = "Distribution of outcomes for each severity each site", tab.id = "tbl_2a", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T}


tbl_2a_d28_site_scores <- severity_data_cc_d28 %>%
  mutate(uva = as.factor(uva),
         mews = as.factor(mews), 
         qsofa = as.factor(qsofa)
         ) %>%
  select(country, outcome_d28, mews, uva, qsofa ) %>%
  tbl_strata(
        strata = country,
    .tbl_fun =
      ~ .x %>%
    tbl_summary(
      by = outcome_d28,
      type = list(all_dichotomous() ~ "categorical"),
      percent = "row",
      missing_text = "(Missing)",
      # ,
      label = list(uva ~ "UVA", mews ~ "MEWS", qsofa ~ "qSOFA")
        ) %>%
      add_overall(last = T, statistic = ~"{n}") %>%
      add_stat_label()
  ) %>%
     modify_header(list(label ~ "Severity score")) %>%
    modify_footnote(list(label ~ "modified early warning score (MEWS), quick Sequential (Sepsis-Related) Organ Failure Assessment (qSOFA) score and Universal Vital Assessment (UVA)")) %>%
    as_flex_table() %>%
      flextable::fontsize(part = "header", size = 8) %>%
      flextable::autofit()

tbl_2a_d28_site_scores


```






\newpage
## UVA ROC results

```{r eval=T, echo=F, message=FALSE, warning=FALSE, fig7, fig.cap="UVA Receiver operating characteristic curve", include=T, paged.print=FALSE, results='asis'}


uva_auc <- pROC::roc(severity_data_cc_d28$outcome_d28, 
           severity_data_cc_d28$uva, percent = T)


uva_age_auc <- pROC::roc(severity_data_cc_d28$outcome_d28, 
           severity_data_cc_d28$uva_age, percent = T)



uva_age_auc_value <- round(as.numeric(auc(uva_age_auc)[1]) / 100, 2)

uva_age_auc_ci <- paste0("95% CI: ", round(pROC::ci.auc(uva_age_auc)[1] / 100, 2), "-", round(ci.auc(uva_age_auc)[3] / 100, 2))
 
uva_age_auc_fig_text <- paste0("AUC = ", uva_age_auc_value, "\n", uva_age_auc_ci)



uva_auc_value <- round(as.numeric(auc(uva_auc)[1])/100, 2)

uva_auc_ci <- paste0("95% CI: ", round(pROC::ci.auc(uva_auc)[1]/100, 2), "-", round(ci.auc(uva_auc)[3]/100, 2))
 
uva_auc_fig_text <- paste0("AUC = ", uva_auc_value, "\n", uva_auc_ci)

```

UVA area under the curve (AUC) = `r uva_auc_value`,  `r uva_auc_ci`   
   
  
  
  


\newpage


```{r, tab.cap = "UVA Specificity (%) and specificity values", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T }

uva_values <- coords(roc = uva_auc, x = "all", transpose = FALSE, ret = "all")

uva_values <- uva_values %>%
  select(threshold, specificity, sensitivity, "1-specificity", accuracy, npv, ppv) %>%
  mutate(threshold_cat = case_when(
    threshold -Inf & threshold <0.5 ~ 0,
    threshold >=0.5 & threshold <1.5 ~ 1,
    threshold >=1.5 & threshold <2.5 ~ 2,
    threshold >=2.5 & threshold <3.5 ~ 3,
    threshold >=3.5 & threshold <4.5 ~ 4,
    threshold >=4.5 & threshold <5.5 ~ 5,
    threshold >=5.5 & threshold <6.5 ~ 6,
    threshold >=6.5 & threshold <7.5 ~ 7,
    threshold >=7.5 & threshold <8.5 ~ 8,
    threshold >=8.5 & threshold <9.5 ~ 9,
    threshold >=9.5 & threshold <10.5 ~ 10,
    threshold >=10.5 ~ 11
  )
  )



uva_age_values <- coords(roc = uva_age_auc, x = "all", transpose = FALSE, ret = "all")

uva_age_values <- uva_age_values %>%
  select(threshold, specificity, sensitivity, "1-specificity", accuracy, npv, ppv) %>%
  mutate(threshold_cat = case_when(
    threshold -Inf & threshold <0.5 ~ 0,
    threshold >=0.5 & threshold <1.5 ~ 1,
    threshold >=1.5 & threshold <2.5 ~ 2,
    threshold >=2.5 & threshold <3.5 ~ 3,
    threshold >=3.5 & threshold <4.5 ~ 4,
    threshold >=4.5 & threshold <5.5 ~ 5,
    threshold >=5.5 & threshold <6.5 ~ 6,
    threshold >=6.5 & threshold <7.5 ~ 7,
    threshold >=7.5 & threshold <8.5 ~ 8,
    threshold >=8.5 & threshold <9.5 ~ 9,
    threshold >=9.5 & threshold <10.5 ~ 10,
    threshold >=10.5 ~ 11
  )
  )


```
\newpage

```{r, fig7a, fig.cap = "UVA Receiver operating characteristic curve", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T }


uva_values_plot <- uva_values %>%
  select(threshold_cat, threshold, specificity, specificity_1 = "1-specificity", sensitivity, specificity) %>%
  # mutate(across(c(threshold, specificity, specificity_1, sensitivity)) %>%
  mutate(label = paste0(threshold_cat, " ", "(", round(specificity, 1), ", ", round(sensitivity, 1), ")"))


uva_values_plot <- uva_values_plot %>%
  ggplot(mapping = aes(x = specificity, y = sensitivity)) +
  geom_line() +
  geom_point() +
  annotate("text", x = 50, y = 50, label = uva_auc_fig_text) +
  geom_label(aes(specificity +.5 , sensitivity + 0.5, label = label), 
             hjust = 0, nudge_x = 3,
             label.size = NA
             ) +
  scale_x_reverse("Specificity (%)", limits = c(100, 0),  breaks = scales::breaks_extended(10)) + 
  scale_y_continuous("Sensitivity (%)", limits = c(0, 100), breaks = scales::breaks_extended(10)) 

uva_values_plot

uva_values <- uva_values %>%
  select(threshold_cat, specificity, sensitivity, npv, ppv) %>%
  mutate(across(c(specificity, sensitivity, npv, ppv), ~round(.x, 2))
         ) %>%
  flextable()

uva_values

```

\newpage
```{r, fig7aa, fig.cap = "UVA age Receiver operating characteristic curve", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T }


uva_age_values_plot <- uva_age_values %>%
  select(threshold_cat, threshold, specificity, specificity_1 = "1-specificity", sensitivity, specificity) %>%
  # mutate(across(c(threshold, specificity, specificity_1, sensitivity)) %>%
  mutate(label = paste0(threshold_cat, " ", "(", round(specificity, 1), ", ", round(sensitivity, 1), ")"))


uva_age_values_plot <- uva_age_values_plot %>%
  ggplot(mapping = aes(x = specificity, y = sensitivity)) +
  geom_line() +
  geom_point() +
  annotate("text", x = 50, y = 50, label = uva_age_auc_fig_text) +
  geom_label(aes(specificity +.5 , sensitivity + 0.5, label = label), 
             hjust = 0, nudge_x = 3,
             label.size = NA
             ) +
  scale_x_reverse("Specificity (%)", limits = c(100, 0),  breaks = scales::breaks_extended(10)) + 
  scale_y_continuous("Sensitivity (%)", limits = c(0, 100), breaks = scales::breaks_extended(10)) 

uva_age_values_plot

uva_age_values <- uva_age_values %>%
  select(threshold_cat, specificity, sensitivity, npv, ppv) %>%
  mutate(across(c(specificity, sensitivity, npv, ppv), ~round(.x, 2))
         ) %>%
  flextable()

uva_age_values

```
\newpage


```{r, tab.cap = "UVA unadjusted and adjusted regression results", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T }


severity_data_cc_d28_reg <- severity_data_cc_d28 %>%
  # filter(uva != 10) %>%
  mutate(uva_fact = as.factor(uva)) 


# Univariate regression

uva_uv_reg_tbl <- severity_data_cc_d28_reg %>%
  select("uva_fact", "age_cat", "sex", "outcome_d28") %>%
  tbl_uvregression(
    method = glm,
    y = outcome_d28,
    method.args = list(family = binomial),
    exponentiate = TRUE,
    pvalue_fun = ~style_pvalue(.x, digits = 2), 
    hide_n = T
  ) %>%
    modify_cols_merge(
    pattern = "{estimate} ({conf.low} - {conf.high})",
    rows = !is.na(estimate)
  ) %>%
    modify_column_hide(columns = ci) %>%
  add_global_p() %>% # add global p-value 
  modify_header(update = list(
    label ~ "Variable",
    estimate ~ "Unadjusted OR (95% CI)",
    # ci ~ "95% CI",
    p.value ~ "p-value"
  )
  )


#  Multivariate regression

uva_mv_reg <- glm(outcome_d28 ~ uva_fact + age_cat + sex, family = "binomial", data = severity_data_cc_d28_reg)

# linear and quadratic term

# uva_mv_reg <- glm(outcome_d28 ~ uva_fact + age + sex, family = "binomial", data = severity_data_cc_d28_reg)

# option to group scores above 6 or 7 

# summary(uva_mv_reg)



uva_mv_reg_tbl <- tbl_regression(uva_mv_reg, 
               exponentiate = TRUE, 
               pvalue_fun = ~style_pvalue(.x, digits = 2)
  ) %>%
  add_global_p() %>%
    modify_header(update = list(
    label ~ "Variable",
    estimate ~ "Adjusted OR",
    ci ~ "95% CI",
    p.value ~ "p-value"
  )
  )

# counts

uva_tbl_count <- severity_data_cc_d28_reg %>% 
  select(uva_fact, age_cat, sex, outcome_d28) %>% # keep only columns of interest
  tbl_summary(     
    by = outcome_d28,                                               # stratify entire table by outcome
    percent = "row",
    statistic = list(c(age_cat) ~ "{n} ({p}%)"),   # stats and format for categorical columns
    digits = all_continuous() ~ 1,                              # rounding for continuous columns
    type   = all_categorical() ~ "categorical"                 # force all categorical levels to display
) %>%
  add_overall(
    statistic = ~"{n}"
    ) %>%
      modify_header(update = list(
    label ~ "Variable"
  )
  )


uva_reg_tbl <- tbl_merge(
  tbls = list(uva_tbl_count, uva_uv_reg_tbl, uva_mv_reg_tbl)
) %>%
  # tab_spanner = c("****","**Univariate**", "**Multivariable**")) %>%
  as_tibble() 

col_Names <- names(uva_reg_tbl)

col_Names <- str_remove_all(col_Names, "\\*")

for(i in 1 : length(uva_reg_tbl))
{
  col_Names[i] <- paste0(col_Names[i], paste0(rep("\r", i), collapse = ""), collapse = "")
}

colnames(uva_reg_tbl) <- col_Names

uva_reg_tbl <- uva_reg_tbl %>%
  mutate("Overall, N = 2,779\r\r" = str_remove_all(uva_reg_tbl[[2]], " \\(100%\\)"))

uva_reg_tbl <- uva_reg_tbl %>%
  flextable() %>%
  flextable::fontsize(part = "header", size = 8) %>%
  flextable::add_footer_lines("OR = Odds ratio, CI = Confidence intervals, Adjusted for age and gender") %>%
  align_text_col(align = "center")


uva_reg_tbl





```




```{r, fig8, fig.cap = "UVA Adjusted odds ratio plot", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T }

uva_mv_reg_plot <- uva_mv_reg %>%
  broom::tidy(exponentiate = TRUE, conf.int = TRUE) %>%
  mutate(across(where(is.numeric), round, digits = 2)) %>%
  tibble()

## remove the intercept term from your multivariable results

fig8 <- uva_mv_reg_plot %>% 
    filter(str_detect(term, "uva_fact*") & 
           term != "uva_fact10") %>% 
  #set order of levels to appear along y-axis
  # mutate(term = fct_relevel(term,
  #   "uva_fact9", "uva_fact8", "uva_fact7","uva_fact6", "uva_fact5", "uva_fact4", "uva_fact3", "uva_fact2", "uva_fact1")
  #   ) %>%

  ggplot(aes(x = estimate, y = term)) +
  geom_point(size = 3, shape=19) + 
  geom_errorbar(aes(xmin = conf.low, xmax = conf.high), width = .3) + 
  geom_vline(xintercept = 1, linetype = "dashed") +
  scale_x_log10(name = "", 
                breaks = scales::log_breaks(8)) +
  scale_y_discrete(name = "Universal Vital Assessment (UVA) score", 
                   labels = c("1", "2", "3", "4", "5", "6", "7", "8", "9")
                   ) +
  labs(caption = "Adjusted for age and gender, (OR for UVA score of 10 was removed in figure)") +
  coord_flip() +
  theme_minimal()

fig8

```
\newpage

Due to the small number of outcomes in UVA scores above 8, UVA scores 9 and 10, were grouped into a UVA score of 8.  
### Table S7
```{r, tab.cap = "Grouped UVA unadjusted and adjusted regression results", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T }


severity_data_cc_d28_reg <- severity_data_cc_d28 %>%
  mutate(uva_grouped = case_when(
    uva == 9 ~ 8,
    uva == 10 ~ 8,
    TRUE ~ uva
  ) ) %>%
  # filter(uva != 10) %>%
  mutate(uva_grouped_fact = as.factor(uva_grouped)) 

severity_data_cc_d28_reg$uva_grouped_fact <- sjlabelled::set_label(severity_data_cc_d28_reg$uva_grouped_fact, "UVA score")

# Univariate regression


uva_uv_reg_tbl <- severity_data_cc_d28_reg %>%
  select("uva_grouped_fact", "age_cat", "sex", "outcome_d28") %>%
  tbl_uvregression(
    method = glm,
    y = outcome_d28,
    method.args = list(family = binomial),
    exponentiate = TRUE,
    pvalue_fun = ~style_pvalue(.x, digits = 2), 
    hide_n = T
  ) %>%
    modify_cols_merge(
    pattern = "{estimate} ({conf.low} - {conf.high})",
    rows = !is.na(estimate)
  ) %>%
    modify_column_hide(columns = ci) %>%
  add_global_p() %>% # add global p-value 
  modify_header(update = list(
    label ~ "Variable",
    estimate ~ "Unadjusted OR",
    # ci ~ "95% CI",
    p.value ~ "p-value"
  )
  )


#  Multivariate regression

uva_mv_reg <- glm(outcome_d28 ~ uva_grouped_fact + age_cat + sex, family = "binomial", data = severity_data_cc_d28_reg)

# linear and quadratic term


uva_mv_reg_tbl <- tbl_regression(uva_mv_reg, 
               exponentiate = TRUE, 
               pvalue_fun = ~style_pvalue(.x, digits = 2)
  ) %>%
  modify_cols_merge(
    pattern = "{estimate} ({conf.low} - {conf.high})",
    rows = !is.na(estimate)
  ) %>%
    modify_column_hide(columns = ci) %>%
  add_global_p() %>%
    modify_header(update = list(
    label ~ "Variable",
    estimate ~ "Adjusted OR (95% CI)",
    p.value ~ "p-value"
  )
  )

# counts

uva_tbl_count <- severity_data_cc_d28_reg %>% 
  select(uva_grouped_fact, age_cat, sex, outcome_d28) %>% # keep only columns of interest
  tbl_summary(     
    by = outcome_d28,                                               # stratify entire table by outcome
    percent = "row",
    statistic = list(c(age_cat) ~ "{n} ({p}%)"),   # stats and format for categorical columns
    digits = all_continuous() ~ 1,                              # rounding for continuous columns
    type   = all_categorical() ~ "categorical"                 # force all categorical levels to display
) %>%
  add_overall(
          statistic = ~"{n}"
  ) %>%
      modify_header(update = list(
    label ~ "Variable"
  )
  )

uva_tbl_count$table_body$stat_0 <- (str_remove_all(uva_tbl_count$table_body$stat_0, "\\(100%\\)")
)

uva_reg_tbl <- tbl_merge(
  tbls = list(uva_tbl_count, uva_uv_reg_tbl, uva_mv_reg_tbl)
) %>%
  modify_spanning_header(everything() ~ NA_character_) %>%
  as_flex_table() %>%
  flextable::fontsize(part = "header", size = 8) %>%
  flextable::add_footer_lines("OR = Odds ratio, CI = Confidence intervals, Adjusted for age and gender") %>%
  align_text_col(align = "center") %>%
    bold(bold = TRUE, part = "header")



uva_reg_tbl

```

```{r, fig8a, fig.cap = "UVA grouped, adjusted odds ratio plot", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T }

uva_mv_reg_plot <- uva_mv_reg %>%
  broom::tidy(exponentiate = TRUE, conf.int = TRUE) %>%
  mutate(across(where(is.numeric), round, digits = 2)) %>%
  tibble()

## remove the intercept term from your multivariable results

fig8a <- uva_mv_reg_plot %>% 
    filter(str_detect(term, "uva_grouped_fact*")) %>% 

  ggplot(aes(x = estimate, y = term)) +
  geom_point(size = 3, shape=19) + 
  geom_errorbar(aes(xmin = conf.low, xmax = conf.high), width = .3) + 
  geom_vline(xintercept = 1, linetype = "dashed") +
  scale_x_log10(name = "Adjusted odds ratio with 95% CI", 
                breaks = scales::log_breaks(8),
                limits = c(0.9, 800)
                ) +
  scale_y_discrete(name = "UVA score", 
                   limits = c("uva_grouped_fact8", "uva_grouped_fact7", "uva_grouped_fact6", "uva_grouped_fact5", "uva_grouped_fact4", "uva_grouped_fact3", "uva_grouped_fact2", "uva_grouped_fact1"), 
                   labels = c("8-10", "7", "6", "5", "4", "3", "2", "1")
                   ) +
  labs(caption = "") +
  # coord_flip() +
  theme_minimal()

fig8a

# Adjusted for age and gender. A UVA score of 8, also includes outcomes for scores 9 and 10

```



\newpage
## MEWS ROC results

```{r, include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T }


mews_auc <- pROC::roc(severity_data_cc_d28$outcome_d28, 
           severity_data_cc_d28$mews, percent = T)

mews_auc_value <- round(as.numeric(auc(mews_auc)[1])/100, 2)
 
mews_auc_ci <- paste0("95% CI: ", round(pROC::ci.auc(mews_auc)[1]/100, 2), "-", round(ci.auc(mews_auc)[3]/100, 2))

mews_auc_fig_text <- paste0("AUC = ", mews_auc_value, "\n", mews_auc_ci)


mews_hiv_auc <- pROC::roc(severity_data_cc_d28$outcome_d28, 
           severity_data_cc_d28$mews_hiv, percent = T)

mews_hiv_auc_value <- round(as.numeric(auc(mews_hiv_auc)[1])/100, 2)

mews_hiv_value_auc_ci <- paste0("95% CI: ", round(pROC::ci.auc(mews_hiv_auc)[1]/100, 2), "-", round(ci.auc(mews_hiv_auc)[3]/100, 2))

```

MEWS area under the curve = `r mews_auc_value`,  `r mews_auc_ci`   
   
  
  

```{r, tab.cap = "MEWS Specificity (%) and specificity values", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T }

mews_values <- coords(roc = mews_auc, x = "all", transpose = FALSE, ret = "all")

mews_values <- mews_values %>%
  select(threshold, specificity, sensitivity, "1-specificity", accuracy, npv, ppv) %>%
  mutate(threshold_cat = case_when(
    threshold -Inf & threshold <1.5 ~ 0,
    threshold >=1.5 & threshold <2.5 ~ 1,
    threshold >=2.5 & threshold <3.5 ~ 2,
    threshold >=3.5 & threshold <4.5 ~ 3,
    threshold >=4.5 & threshold <5.5 ~ 4,
    threshold >=5.5 & threshold <6.5 ~ 5,
    threshold >=6.5 & threshold <7.5 ~ 6,
    threshold >=7.5 & threshold <8.5 ~ 7,
    threshold >=8.5 & threshold <9.5 ~ 8,
    threshold >=9.5 & threshold <10.5 ~ 9,
    threshold >=10.5 & threshold <11.5 ~ 10,
    threshold >=11.5 ~ 11
  )
  )


# %>%
#   mutate(across(c(sensitivity, specificity, "1-specificity", accuracy, npv, ppv))) 

```


```{r, fig9, fig.cap = "MEWS Receiver operating characteristic curve"}

mews_values_plot <- mews_values %>%
  select(threshold_cat, threshold, specificity, specificity_1 = "1-specificity", sensitivity, specificity) %>%
  # mutate(across(c(threshold, specificity, specificity_1, sensitivity)) %>%
  mutate(label = paste0(threshold_cat, " ", "(", round(specificity, 1), ", ", round(sensitivity, 1), ")"))


mews_values_plot <- mews_values_plot %>%
  ggplot(mapping = aes(x = specificity, y = sensitivity)) +
  geom_line() +
  geom_point() +
  annotate("text", x = 40, y = 50, label = mews_auc_fig_text) +
  geom_label(aes(specificity +.5 , sensitivity + 0.5, label = label), 
             hjust = 0, nudge_x = 3,
             label.size = NA
             ) +
  scale_x_reverse("Specificity (%)", limits = c(100, 0),  breaks = scales::breaks_extended(10)) + 
  scale_y_continuous("Sensitivity (%)", limits = c(0, 100), breaks = scales::breaks_extended(10)) 

mews_values_plot

mews_values <- mews_values %>% 
  select(threshold_cat, specificity, sensitivity, npv, ppv) %>%
  mutate(across(c(specificity, sensitivity, npv, ppv), ~round(.x, 2))
         ) %>%
  flextable()

mews_values

```


\newpage

```{r, tab.cap = "MEWS score by Day 28 outcome", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T }


severity_data_cc_d28 %>%
  select(mews, age_cat, sex, outcome_d28) %>%
  mutate(mews = as.factor(mews)) %>%
  tbl_summary(by = outcome_d28)  %>%
    add_stat_label() %>%
     modify_header(list(label ~"Variable")) %>%
    as_flex_table() %>%
      flextable::fontsize(part = "header", size = 8) %>%
      flextable::autofit()


```


```{r, tab.cap = "Clinical signs for a MEWS score of 0", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = F }


severity_data_cc_d28 %>%
  select(mews, px_eyes, px_eyes_no, temp, bp_systolic, heart_rate, respiratory_rate, modGCS) %>%
  filter(mews < 1) %>%
  tbl_summary(
    statistic = list(c("px_eyes", "px_eyes_no", "temp", "bp_systolic", "heart_rate", "respiratory_rate", "modGCS") ~ "{n}")
  ) %>%
    add_stat_label() %>%
     modify_header(list(label ~"Variable")) %>%
    as_flex_table() %>%
      flextable::fontsize(part = "header", size = 8) %>%
      flextable::autofit()


```

\newpage

```{r, tab.cap = "MEWS unadjusted and adjusted regression results", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T }


severity_data_cc_d28_reg <- severity_data_cc_d28 %>%
  mutate(mews_fact = as.factor(mews),
         hiv_comb = fct_relevel(hiv_comb, c("Negative", "Positive", "Unknown", "Indeterminant"))
         ) %>%
  filter(as.character(hiv_comb) == "Negative" | as.character(hiv_comb) == "Positive" ) %>%
  mutate(qsofa_fact = as.factor(qsofa))


mews_uv_reg_tbl <- severity_data_cc_d28_reg %>%
  select("mews_fact", "age_cat", "sex", "hiv_comb", "outcome_d28") %>%
  tbl_uvregression(
    method = glm,
    y = outcome_d28,
    method.args = list(family = binomial),
    exponentiate = TRUE,
    pvalue_fun = ~style_pvalue(.x, digits = 2), 
    hide_n = T
  ) %>%
  add_global_p() %>% # add global p-value 
  modify_header(update = list(
    label ~ "Variable",
    estimate ~ "Unadjusted OR",
    ci ~ "95% CI",
    p.value ~ "p-value"
  )
  )

#  Multivariate regression

mews_mv_reg <- glm(outcome_d28 ~ mews_fact + age_cat + sex, family = "binomial", data = severity_data_cc_d28_reg)


mews_mv_reg_tbl <- tbl_regression(mews_mv_reg, 
               exponentiate = TRUE, 
               pvalue_fun = ~style_pvalue(.x, digits = 2)
  ) %>%
  add_global_p() %>%
    modify_header(update = list(
    label ~ "Variable",
    estimate ~ "Adjusted OR",
    ci ~ "95% CI",
    p.value ~ "p-value"
  )
  )

# counts

mews_tbl_count <- severity_data_cc_d28_reg %>% 
  select(mews_fact, age_cat, sex, outcome_d28) %>% # keep only columns of interest
  tbl_summary(     
    by = outcome_d28,                                               # stratify entire table by outcome
    percent = "row",
    statistic = list(c(age_cat) ~ "{n} ({p}%)"),   # stats and format for categorical columns
    digits = all_continuous() ~ 1,                              # rounding for continuous columns
    type   = all_categorical() ~ "categorical"                 # force all categorical levels to display
) %>%
  add_overall(
      statistic = ~"{n}"
      ) %>%
      modify_header(update = list(
    label ~ "Variable"
  )
  )


mews_reg_tbl <- tbl_merge(
  tbls = list(mews_tbl_count, mews_uv_reg_tbl, mews_mv_reg_tbl)
) %>%
  # tab_spanner = c("****","**Univariate**", "**Multivariable**")) %>%
  as_tibble() 

col_Names <- names(mews_reg_tbl)

col_Names <- str_remove_all(col_Names, "\\*")

for(i in 1 : length(mews_reg_tbl))
{
  col_Names[i] <- paste0(col_Names[i], paste0(rep("\r", i), collapse = ""), collapse = "")
}

colnames(mews_reg_tbl) <- col_Names

mews_reg_tbl <- mews_reg_tbl %>%
  mutate("Overall, N = 2,779\r\r" = str_remove_all(mews_reg_tbl[[2]], " \\(100%\\)"))

mews_reg_tbl <- mews_reg_tbl %>%
  flextable() %>%
  flextable::fontsize(part = "header", size = 8) %>%
  flextable::add_footer_lines("OR = Odds ratio, CI = Confidence intervals, Adjusted for age and gender, Reference is a MEWS score of 1 (rather than 0)") %>%
  align_text_col(align = "center")


mews_reg_tbl
```

\newpage
```{r eval=, echo=F, message=FALSE, warning=FALSE, fig9_OR, fig.cap="MEWS Adjusted odds ratio plot", include=T, paged.print=FALSE, results='asis', eval = F}

mews_mv_reg_plot <- mews_mv_reg %>%
  broom::tidy(exponentiate = TRUE, conf.int = TRUE) %>%
  mutate(across(where(is.numeric), round, digits = 2)) %>%
  tibble()

## remove the intercept term from your multivariable results

fig9 <- mews_mv_reg_plot %>% 
    filter(str_detect(term, "mews_fact*") & 
           term != "mews_fact11") %>% 
  #set order of levels to appear along y-axis
  mutate(term = fct_relevel(term,
    "mews_fact2", "mews_fact3", "mews_fact4","mews_fact5", "mews_fact6", "mews_fact7", "mews_fact8", "mews_fact9", "mews_fact10")
    ) %>%

  ggplot(aes(x = estimate, y = term)) +
  geom_point(size = 3, shape=19) + 
  geom_errorbar(aes(xmin = conf.low, xmax = conf.high), width = .3) + 
  geom_vline(xintercept = 1, linetype = "dashed") +
  scale_x_log10(breaks = scales::log_breaks(8), 
                limits = c(1, 800)
) +
  scale_y_discrete(labels = c("2", "3", "4", "5", "6", "7", "8", "9", "10")
                   ) +
  labs(caption = "Adjusted for age and gender, reference group is a MEWS score of 1, rather than 0.", 
       x = "Adjusted odds ratio with 95% CI",
       y = "Modified early warning score (MEWS)"
       ) +
  coord_flip() +
  theme_minimal()

fig9

```

\newpage

Due to the small number of outcomes in MEWS scores of 10 and 11, these scores were grouped into a MEWS score of 9.  
### Table S5: 
```{r, tab.cap = "Grouped MEWS unadjusted and adjusted regression results", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T }


severity_data_cc_d28_reg <- severity_data_cc_d28 %>%
  mutate(mews_grouped = case_when(
    mews == 0 ~ 1,
    mews == 10 ~ 9,
    mews == 11 ~ 9,
    TRUE ~ mews
  ) ) %>%
  # filter(uva != 10) %>%
  mutate(mews_grouped_fact = as.factor(mews_grouped),
         hiv_comb = fct_relevel(hiv_comb, c("Negative", "Positive", "Unknown", "Indeterminant"))
         ) %>%
  filter(as.character(hiv_comb) == "Negative" | as.character(hiv_comb) == "Positive" ) %>%
    mutate(qsofa_fact = as.factor(qsofa))



severity_data_cc_d28_reg$hiv_comb <-   droplevels(severity_data_cc_d28_reg$hiv_comb, c("Unknown", "Indeterminant"))


severity_data_cc_d28_reg$hiv_comb <- sjlabelled::set_label(severity_data_cc_d28_reg$hiv_comb, "HIV status")

severity_data_cc_d28_reg$mews_grouped_fact <- sjlabelled::set_label(severity_data_cc_d28_reg$mews_grouped_fact, "MEWS")

severity_data_cc_d28_reg$qsofa_fact <- sjlabelled::set_label(severity_data_cc_d28_reg$qsofa_fact, "qSOFA score")


# Univariate regression


mews_uv_reg_tbl <- severity_data_cc_d28_reg %>%
  select("mews_grouped_fact", "age_cat", "sex", "hiv_comb", "outcome_d28")  %>%
  tbl_uvregression(
    method = glm,
    y = outcome_d28,
    method.args = list(family = binomial),
    exponentiate = TRUE,
    pvalue_fun = ~style_pvalue(.x, digits = 2), 
    hide_n = T
  ) %>%
  modify_cols_merge(
    pattern = "{estimate} ({conf.low} - {conf.high})",
    rows = !is.na(estimate)
  ) %>%
  modify_column_hide(columns = ci) %>%
  add_global_p() %>% # add global p-value 
  modify_header(update = list(
    label ~ "Variable",
    estimate ~ "Unadjusted OR (95% CI)",
    p.value ~ "p-value"
  )
  )


#  Multivariate regression

mews_mv_reg <- glm(outcome_d28 ~ mews_grouped_fact + age_cat + sex + hiv_comb, family = "binomial", data = severity_data_cc_d28_reg)

# linear and quadratic term


mews_mv_reg_tbl <- tbl_regression(mews_mv_reg, 
               exponentiate = TRUE, 
               pvalue_fun = ~style_pvalue(.x, digits = 2)
  ) %>%
   modify_cols_merge(
    pattern = "{estimate} ({conf.low} - {conf.high})",
    rows = !is.na(estimate)
  ) %>%
  modify_column_hide(columns = ci) %>%
  add_global_p() %>%
    modify_header(update = list(
    label ~ "Variable",
    estimate ~ "Adjusted OR  (95% CI)",
    p.value ~ "p-value"
  )
  )

# counts

mews_tbl_count <- severity_data_cc_d28_reg %>% 
  select(mews_grouped_fact, age_cat, sex, hiv_comb, outcome_d28) %>% # keep only columns of interest
  tbl_summary(     
    by = outcome_d28,                                               # stratify entire table by outcome
    percent = "row",
    statistic = list(c(age_cat) ~ "{n} ({p}%)"),   # stats and format for categorical columns
    digits = all_continuous() ~ 1,                              # rounding for continuous columns
    type   = all_categorical() ~ "categorical"                 # force all categorical levels to display
) %>%
  add_overall(statistic = ~"{n}") %>%
      modify_header(update = list(
    label ~ "Variable"
  )
  )

mews_tbl_count$table_body$stat_0 <- (str_remove_all(mews_tbl_count$table_body$stat_0, "\\(100%\\)")
)

mews_reg_tbl <- tbl_merge(
  tbls = list(mews_tbl_count, mews_uv_reg_tbl, mews_mv_reg_tbl)
) %>%
  modify_spanning_header(everything() ~ NA_character_) %>%
  as_flex_table() %>%
flextable::fontsize(part = "header", size = 8) %>%
  flextable::add_footer_lines("OR = Odds ratio, CI = Confidence intervals, Adjusted for age and gender, MEWS scores 10 and 11 are grouped into a MEWS score of 9, reference group is a MEWS score of 1, rather than 0.") %>%
  align_text_col(align = "center") %>%
  bold(bold = TRUE, part = "header")

mews_reg_tbl

```

```{r, fig9b, fig.cap = "MEWS grouped, adjusted odds ratio plot", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T }

mews_mv_reg_plot <- mews_mv_reg %>%
  broom::tidy(exponentiate = TRUE, conf.int = TRUE) %>%
  mutate(across(where(is.numeric), round, digits = 2)) %>%
  tibble()

## remove the intercept term from your multivariable results

fig9b <- mews_mv_reg_plot %>% 
    filter(str_detect(term, "mews_grouped_fact*")) %>% 
     arrange((term)) %>%

  ggplot(aes(x = estimate, y = term)) +
  geom_point(size = 3, shape=19) + 
  geom_errorbar(aes(xmin = conf.low, xmax = conf.high), width = .3) + 
  geom_vline(xintercept = 1, linetype = "dashed") +
  scale_x_log10(name = "", 
                breaks = scales::log_breaks(8), 
                                limits = c(0.9, 800)
) +
  scale_y_discrete(name = "MEWS", 
                   limits = c("mews_grouped_fact9", "mews_grouped_fact8", "mews_grouped_fact7", "mews_grouped_fact6", "mews_grouped_fact5", "mews_grouped_fact4", "mews_grouped_fact3", "mews_grouped_fact2"),
                   # , 
                   labels = c("9-11", "8", "7", "6", "5", "4", "3", "2")
                   
                   ) +
  # scale_y_reverse() +
  labs(caption = "") +
  # scale_x_reverse() +
  # coord_flip() +
  theme_minimal()

fig9b

# Adjusted for age and gender. A MEWS of 9, also includes outcomes for scores 10 and 11, reference group is a MEWS of 1
# 
# h <- ggplot(diamonds, aes(carat)) +
#   geom_histogram()
# h
# h + coord_flip()
# h + coord_flip() + scale_x_reverse()

```

\newpage

## qSOFA tables and figures

```{r, fig10, fig.cap = "qSOFA Receiver operating characteristic curve", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T }


qsofa_auc <- pROC::roc(severity_data_cc_d28$outcome_d28, 
           severity_data_cc_d28$qsofa, percent = T)

qsofa_auc_value <- round(as.numeric(auc(qsofa_auc)[1])/100, 2)
 
qsofa_auc_ci <- paste0("95% CI: ", round(pROC::ci.auc(qsofa_auc)[1]/100, 2), "-", round(ci.auc(qsofa_auc)[3]/100, 2))

qsofa_auc_fig_text <- paste0("AUC = ", qsofa_auc_value, "\n", qsofa_auc_ci)



qsofa_hiv_auc <- pROC::roc(severity_data_cc_d28$outcome_d28, 
           severity_data_cc_d28$qsofa_hiv, percent = T)

qsofa_hiv_auc_value <- round(as.numeric(auc(qsofa_hiv_auc)[1])/100, 2)

qsofa_hiv_value_auc_ci <- paste0("95% CI: ", round(pROC::ci.auc(qsofa_hiv_auc)[1]/100, 2), "-", round(ci.auc(qsofa_hiv_auc)[3]/100, 2))


```

qSOFA area under the curve = `r qsofa_auc_value`,  `r qsofa_auc_ci`


```{r, tab.cap = "qSOFA Specificity (%) and specificity values", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T }

qsofa_values <- coords(roc = qsofa_auc, x = "all", transpose = FALSE, ret = "all")

qsofa_values <- qsofa_values %>%
  select(threshold, specificity, sensitivity, "1-specificity", accuracy, npv, ppv) %>%
  mutate(threshold_cat = case_when(
    threshold -Inf & threshold <0.5 ~ 0,
    threshold >=0.5 & threshold <1.5 ~ 1,
    threshold >=1.5 & threshold <2.5 ~ 2,
    threshold >=2.5 & threshold <3.5 ~ 3,
    threshold >=3.5  ~ 4
    
  )
  )

# %>%
#   mutate(across(c(sensitivity, specificity, "1-specificity", accuracy, npv, ppv))) 

qsofa_values_plot <- qsofa_values %>%
  select(threshold_cat, threshold, specificity, specificity_1 = "1-specificity", sensitivity, specificity) %>%
  # mutate(across(c(threshold, specificity, specificity_1, sensitivity)) %>%
  mutate(label = paste0(threshold_cat, " ", "(", round(specificity, 1), ", ", round(sensitivity, 1), ")"))


qsofa_values_plot <- qsofa_values_plot %>%
  ggplot(mapping = aes(x = specificity, y = sensitivity)) +
  geom_line() +
  geom_point() +
  annotate("text", x = 40, y = 50, label = qsofa_auc_fig_text) +
  geom_label(aes(specificity +.5 , sensitivity + 0.5, label = label), 
             hjust = 0, nudge_x = 3,
             label.size = NA
             ) +
  scale_x_reverse("Specificity (%)", limits = c(100, 0),  breaks = scales::breaks_extended(10)) + 
  scale_y_continuous("Sensitivity (%)", limits = c(0, 100), breaks = scales::breaks_extended(10)) 

qsofa_values_plot 

```
\newpage

```{r, tab.cap = "qSOFA Specificity (%) and specificity values", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T }


qsofa_values <- qsofa_values %>%
  select(threshold_cat, specificity, sensitivity, "1-specificity", accuracy, npv, ppv) 
  

qsofa_values %>%
  select(threshold_cat, specificity, sensitivity, npv, ppv) %>%
  mutate(across(c(specificity, sensitivity, npv, ppv), ~round(.x, 2))
         ) %>%
  flextable()
# %>%
#   mutate(across(c(sensitivity, specificity, "1-specificity", accuracy, npv, ppv))) 

```
### Table S6: 
```{r, tab.cap = "qSOFA unadjusted and adjusted regression results", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T }



qsofa_uv_reg_tbl <- severity_data_cc_d28_reg %>%
  select("qsofa_fact", "age_cat", "sex", "hiv_comb", "outcome_d28") %>%
  tbl_uvregression(
    method = glm,
    y = outcome_d28,
    method.args = list(family = binomial),
    exponentiate = TRUE,
    pvalue_fun = ~style_pvalue(.x, digits = 2), 
    hide_n = T
  ) %>%
  modify_cols_merge(
    pattern = "{estimate} ({conf.low} - {conf.high})",
    rows = !is.na(estimate)
  ) %>%
    modify_column_hide(columns = ci) %>%
  add_global_p() %>% # add global p-value 
  modify_header(update = list(
    label ~ "Variable",
    estimate ~ "Unadjusted OR",
    p.value ~ "p-value"
  )
  )

#  Multivariate regression

qsofa_mv_reg <- glm(outcome_d28 ~ qsofa_fact + age_cat + sex + hiv_comb, family = "binomial", data = severity_data_cc_d28_reg)

qsofa_mv_reg_tbl <- tbl_regression(qsofa_mv_reg, 
               exponentiate = TRUE, 
               pvalue_fun = ~style_pvalue(.x, digits = 2)
  ) %>%
    modify_cols_merge(
    pattern = "{estimate} ({conf.low} - {conf.high})",
    rows = !is.na(estimate)
  ) %>%
  modify_column_hide(columns = ci) %>%
  add_global_p() %>%
    modify_header(update = list(
    label ~ "Variable",
    estimate ~ "Adjusted OR (95% CI)",
    p.value ~ "p-value"
  )
  )

# counts

qsofa_tbl_count <- severity_data_cc_d28_reg %>% 
  select(qsofa_fact, age_cat, sex, hiv_comb, outcome_d28) %>% # keep only columns of interest
  tbl_summary(     
    by = outcome_d28,                                               # stratify entire table by outcome
    percent = "row",
    statistic = list(c(age_cat) ~ "{n} ({p}%)"),   # stats and format for categorical columns
    digits = all_continuous() ~ 1,                              # rounding for continuous columns
    type   = all_categorical() ~ "categorical"                 # force all categorical levels to display
) %>%
  add_overall(statistic = ~"{n}") %>%
      modify_header(update = list(
    label ~ "Variable"
  )
  )
qsofa_tbl_count$table_body$stat_0 <- (str_remove_all(qsofa_tbl_count$table_body$stat_0, "\\(100%\\)")
)

qsofa_reg_tbl <- tbl_merge(
  tbls = list(qsofa_tbl_count, qsofa_uv_reg_tbl, qsofa_mv_reg_tbl)
) %>%
  modify_spanning_header(everything() ~ NA_character_) %>%
  as_flex_table() %>%
  flextable::fontsize(part = "header", size = 8) %>%
  flextable::add_footer_lines("OR = Odds ratio, CI = Confidence intervals, Adjusted for age and gender") %>%
  align_text_col(align = "center") %>%
    bold(bold = TRUE, part = "header")



qsofa_reg_tbl

```

\newpage
```{r, fig10_OR, fig.cap = "qSOFA Adjusted odds ratio plot", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T }

qsofa_mv_reg_plot <- qsofa_mv_reg %>%
  broom::tidy(exponentiate = TRUE, conf.int = TRUE) %>%
  mutate(across(where(is.numeric), round, digits = 2)) %>%
  tibble()

## remove the intercept term from your multivariable results

fig10 <- qsofa_mv_reg_plot %>% 
    filter(str_detect(term, "qsofa_fact*") & 
           term != "qsofa_fact11") %>% 
  #set order of levels to appear along y-axis
  mutate(term = fct_relevel(term,
    "qsofa_fact1", "qsofa_fact2", "qsofa_fact3")
    ) %>%

  ggplot(aes(x = estimate, y = term)) +
  geom_point(size = 3, shape=19) + 
  geom_errorbar(aes(xmin = conf.low, xmax = conf.high), width = .3) + 
  geom_vline(xintercept = 1, linetype = "dashed") +
  scale_x_log10(breaks = scales::log_breaks(8),
                                limits = c(0.9, 800)
) +
  scale_y_discrete(limits = c("qsofa_fact3", "qsofa_fact2","qsofa_fact1"), 
    labels = c("3", "2", "1")
                   ) +
  labs(caption = "", 
       x = "",
       y = "qSOFA score"
       ) +
  # coord_flip() +
  theme_minimal()

fig10

```

\newpage
## All ORs 
### Figure 3: Combined OR plot 

```{r all_ORs_combined, fig.cap = "Odds ratio plots for A) MEWs, B) qSOFA, C) UVA", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T, fig.width = 10, fig.height=7}

combined_ors <- cowplot::plot_grid(fig9b, NULL, fig10, NULL, fig8a,
                                   nrow = 5, 
                                   rel_heights = c(2, -0.3, 2, -0.3, 2),
                               labels = c('A', "", 'B', "" ,"C"), label_size = 10, align = "v", vjust = c(1.5, 0, 0, -1, 1.5)
                               )

combined_ors

```
### Figure 4: Combined ROC
```{r all_rocs, fig.cap = "ROCs for 1) MEWs, 2) qSOFA, 3) UVA", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T, fig.width = 9, fig.height=6}

uva_values_plot$layers[[4]] <- NULL

uva_values_plot <- uva_values_plot +
  ggtitle("UVA") +
  theme_minimal()

mews_values_plot$layers[[4]] <- NULL

mews_values_plot <- mews_values_plot +
  ggtitle("MEWS") +
  theme_minimal()

qsofa_values_plot$layers[[4]] <- NULL

qsofa_values_plot <- qsofa_values_plot +
  ggtitle("qSOFA") +
  theme_minimal()

combined <- cowplot::plot_grid(mews_values_plot, qsofa_values_plot, uva_values_plot, ncol = 3, 
                               labels = c('A', 'B', "C"), label_size = 12, 
                               align = "h")

combined  

```


```{r HIV, include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T }

uva_v_qsofa <- roc.test(uva_auc, qsofa_auc, method = c("delong"))
uva_v_qsofa_z <- as.numeric(uva_v_qsofa$statistic)
uva_v_qsofa_p <- format(uva_v_qsofa$p.value, scientific = F)

uva_v_mews <- roc.test(uva_auc, mews_auc, method = c("delong"))
uva_v_mews_z <- as.numeric(uva_v_mews$statistic)
uva_v_mews_p <- format(uva_v_mews$p.value, scientific = F)

mews_v_qsofa <- roc.test(mews_auc, qsofa_auc, method = c("delong"))
mews_v_qsofa_z <- as.numeric(mews_v_qsofa$statistic)
mews_v_qsofa_p <- format(mews_v_qsofa$p.value, scientific = F)

# MEWS vs MEWS HIV

mews_v_mews_hiv <- roc.test(mews_auc, mews_hiv_auc, method = c("delong"))
mews_v_mews_hiv_z <- as.numeric(mews_v_mews_hiv$statistic)
mews_v_mews_hiv_p <- format(mews_v_mews_hiv$p.value, scientific = F)

# qSOFA vs qSOFA HIV

qsofa_v_qsofa_hiv <- roc.test(qsofa_auc, qsofa_hiv_auc, method = c("delong"))
qsofa_v_qsofa_hiv_z <- as.numeric(qsofa_v_qsofa_hiv$statistic)
qsofa_v_qsofa_hiv_p <- format(qsofa_v_qsofa_hiv$p.value, scientific = F)

# MEWS HIV vs UVA

uva_v_mews_hiv <- roc.test(mews_hiv_auc, uva_auc, method = c("delong"))
uva_v_mews_hiv_z <- as.numeric(uva_v_mews_hiv$statistic)
uva_v_mews_hiv_p <- format(uva_v_mews_hiv$p.value, scientific = F)


# qSOFA HIV vs UVA

uva_v_qsofa_hiv <- roc.test(qsofa_hiv_auc, uva_auc, method = c("delong"))
uva_v_qsofa_hiv_z <- as.numeric(uva_v_qsofa_hiv$statistic)
uva_v_qsofa_hiv_p <- format(uva_v_qsofa_hiv$p.value, scientific = F)

# qSOFA HIV vs MEWS HIV


qsofa_hiv_v_mews_hiv <- roc.test(qsofa_hiv_auc, mews_hiv_auc, method = c("delong"))
qsofa_hiv_v_mews_hiv_z <- as.numeric(qsofa_hiv_v_mews_hiv$statistic)
qsofa_hiv_v_mews_hiv_p <- format(qsofa_hiv_v_mews_hiv$p.value, scientific = F)

# UVA vs UVA age

uva_v_uva_age <- roc.test(uva_auc, uva_age_auc, method = c("delong"))

# AUC of UVA AUC v UVA_Age 82.0, 84.3 p = 0.005

```

DeLong test, a pairwise comparison test for the equality of the area under the curve between;   
  1) UVA (AUC:`r uva_auc_value`, `r uva_auc_ci`) vs. MEWS (AUC:`r mews_auc_value`,  `r mews_auc_ci`), Z = `r uva_v_qsofa_z`, *p-value* = `r uva_v_qsofa_p`.   
  2) UVA (AUC:`r uva_auc_value`, `r uva_auc_ci`) vs. qSOFA (AUC:`r qsofa_auc_value`,  `r qsofa_auc_ci`), Z = `r uva_v_mews_z`, *p-value* = `r uva_v_mews_p`.   
  3) MEWs (AUC:`r mews_auc_value`,  `r mews_auc_ci`) vs. qSOFA (AUC:`r qsofa_auc_value`,  `r qsofa_auc_ci`), Z = `r mews_v_qsofa_z`, *p-value* = `r mews_v_qsofa_p`.   


\newpage

## ROCs for those aged 55 years and above


```{r all_rocs_old, fig.cap = "ROCs 55yrs and above for 1) MEWs, 2) qSOFA 3) UVA", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T, fig.width = 9, fig.height=6}

severity_data_cc_d28_old <- severity_data_cc_d28 %>%
  filter(age >54)

# UVA 55+

uva_auc_old <- pROC::roc(severity_data_cc_d28_old$outcome_d28, 
           severity_data_cc_d28_old$uva, percent = T)

uva_auc_value_old <- round(as.numeric(auc(uva_auc_old)[1])/100, 2)

uva_values_old <- coords(roc = uva_auc_old, x = "all", transpose = FALSE, ret = "all")

uva_auc_ci_old <- paste0("95% CI: ", round(pROC::ci.auc(uva_auc_old)[1]/100, 2), "-", round(ci.auc(uva_auc_old)[3]/100, 2))

uva_auc_fig_text_old <- paste0("AUC = ", uva_auc_value_old, "\n", uva_auc_ci_old)

uva_values_plot_old <- uva_values_old %>%
  ggplot(mapping = aes(x = specificity, y = sensitivity)) +
  geom_line() +
  geom_point() +
  annotate("text", x = 50, y = 50, label = uva_auc_fig_text_old) +
  # geom_label(aes(specificity +.5 , sensitivity + 0.5, label = label), 
  #            hjust = 0, nudge_x = 3,
  #            label.size = NA
  #            ) +
  scale_x_reverse("Specificity (%)", limits = c(100, 0),  breaks = scales::breaks_extended(10)) + 
  scale_y_continuous("Sensitivity (%)", limits = c(0, 100), breaks = scales::breaks_extended(10)) 

# MEWS 55+

mews_auc_old <- pROC::roc(severity_data_cc_d28_old$outcome_d28, 
           severity_data_cc_d28_old$mews, percent = T)

mews_auc_value_old <- round(as.numeric(auc(mews_auc_old)[1])/100, 2)

mews_values_old <- coords(roc = mews_auc_old, x = "all", transpose = FALSE, ret = "all")

mews_auc_ci_old <- paste0("95% CI: ", round(pROC::ci.auc(mews_auc_old)[1]/100, 2), "-", round(ci.auc(mews_auc_old)[3]/100, 2))

mews_auc_fig_text_old <- paste0("AUC = ", mews_auc_value_old, "\n", mews_auc_ci_old)

mews_values_plot_old <- mews_values_old %>%
  ggplot(mapping = aes(x = specificity, y = sensitivity)) +
  geom_line() +
  geom_point() +
  annotate("text", x = 50, y = 50, label = mews_auc_fig_text_old) +
  # geom_label(aes(specificity +.5 , sensitivity + 0.5, label = label), 
  #            hjust = 0, nudge_x = 3,
  #            label.size = NA
  #            ) +
  scale_x_reverse("Specificity (%)", limits = c(100, 0),  breaks = scales::breaks_extended(10)) + 
  scale_y_continuous("Sensitivity (%)", limits = c(0, 100), breaks = scales::breaks_extended(10)) 

# qSOFA 55+

qsofa_auc_old <- pROC::roc(severity_data_cc_d28_old$outcome_d28, 
           severity_data_cc_d28_old$qsofa, percent = T)

qsofa_auc_value_old <- round(as.numeric(auc(qsofa_auc_old)[1])/100, 2)

qsofa_values_old <- coords(roc = qsofa_auc_old, x = "all", transpose = FALSE, ret = "all")

qsofa_auc_ci_old <- paste0("95% CI: ", round(pROC::ci.auc(qsofa_auc_old)[1]/100, 2), "-", round(ci.auc(qsofa_auc_old)[3]/100, 2))

qsofa_auc_fig_text_old <- paste0("AUC = ", qsofa_auc_value_old, "\n", qsofa_auc_ci_old)

qsofa_values_plot_old <- qsofa_values_old %>%
  ggplot(mapping = aes(x = specificity, y = sensitivity)) +
  geom_line() +
  geom_point() +
  annotate("text", x = 50, y = 50, label = qsofa_auc_fig_text_old) +
  # geom_label(aes(specificity +.5 , sensitivity + 0.5, label = label), 
  #            hjust = 0, nudge_x = 3,
  #            label.size = NA
  #            ) +
  scale_x_reverse("Specificity (%)", limits = c(100, 0),  breaks = scales::breaks_extended(10)) + 
  scale_y_continuous("Sensitivity (%)", limits = c(0, 100), breaks = scales::breaks_extended(10)) 


uva_values_plot_old$layers[[4]] <- NULL

uva_values_plot_old <- uva_values_plot_old +
  ggtitle("UVA ROC (55yrs+)")

mews_values_plot_old$layers[[4]] <- NULL

mews_values_plot_old <- mews_values_plot_old +
  ggtitle("MEWS ROC (55yrs+)")

qsofa_values_plot_old$layers[[4]] <- NULL

qsofa_values_plot_old <- qsofa_values_plot_old +
  ggtitle("qSOFA ROC (55yrs+)")

combined_old <- cowplot::plot_grid(mews_values_plot_old, qsofa_values_plot_old, uva_values_plot_old, ncol = 3, 
                               labels = c('1', '2', "3"), label_size = 12, 
                               align = "h")

combined_old

```

```{r, include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T}

uva_v_qsofa_old <- roc.test(uva_auc_old, qsofa_auc_old, method = c("delong"))
uva_v_qsofa_old_z <- as.numeric(uva_v_qsofa_old$statistic)
uva_v_qsofa_old_p <- format(uva_v_qsofa_old$p.value, scientific = F)

uva_v_mews_old <- roc.test(uva_auc_old, mews_auc_old, method = c("delong"))
uva_v_mews_old_z <- as.numeric(uva_v_mews_old$statistic)
uva_v_mews_old_p <- format(uva_v_mews_old$p.value, scientific = F)

mews_v_qsofa_old <- roc.test(mews_auc_old, qsofa_auc_old, method = c("delong"))
mews_v_qsofa_old_z <- as.numeric(mews_v_qsofa_old$statistic)
mews_v_qsofa_old_p <- format(mews_v_qsofa_old$p.value, scientific = F)

```

DeLong test, a pairwise comparison test for the equality of the area under the curve between for 55yrs and above.   
  1) UVA (AUC: `r uva_auc_value_old`, `r uva_auc_ci_old`) vs. MEWS (AUC: `r mews_auc_value_old`,  `r mews_auc_ci_old`), Z = `r uva_v_qsofa_old_z`, *p-value* = `r uva_v_qsofa_old_p`.   
  2) UVA (AUC: `r uva_auc_value_old`, `r uva_auc_ci_old`) vs. qSOFA (AUC: `r qsofa_auc_value_old`,  `r qsofa_auc_ci_old`), Z = `r uva_v_mews_old_z`, *p-value* = `r uva_v_mews_old_p`.   
  3) MEWs (AUC: `r mews_auc_value_old`,  `r mews_auc_ci_old`) vs. qSOFA (AUC: `r qsofa_auc_value_old`,  `r qsofa_auc_ci_old`), Z = `r mews_v_qsofa_old_z`, *p-value* = `r mews_v_qsofa_old_p`.
  
\newpage

## Specificity (%) analysis   

In this analysis we assume all those who are missing a day 28 outcome are either a) alive or b) dead and calculate AUC for each severity score.   

### a) Assume all missing are alive. 

```{r, sensitivity data}

sensitivity_analysis_alive <- severity_data %>%
  filter(if_all(c(modGCS, qsofa, uva, mews), ~ !is.na(.))) %>%
  mutate(outcome_d28 = fct_explicit_na(outcome_d28, na_level = "Alive")) %>%
  mutate(outcome_d28 = fct_recode(outcome_d28, "lost to fup" = "Alive"))

  
  
sensitivity_analysis_dead <- severity_data %>%
  filter(if_all(c(modGCS, qsofa, uva, mews), ~ !is.na(.))) %>%
  mutate(outcome_d28 = fct_explicit_na(outcome_d28, na_level = "Dead")) %>%
  mutate(outcome_d28 = fct_recode(outcome_d28, "lost to fup" = "Dead"))





  # uva_auc <- pROC::roc(severity_data_cc_d28$outcome_d28, 
  #          severity_data_cc_d28$uva, percent = T)

```

```{r, tab.cap = "Specificity (%) analysis (a) assuming all missing a day 28 outcome are alive", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T }


tbl_count_alive <- sensitivity_analysis_alive %>% 
  select(uva, mews, qsofa, age_cat, sex, outcome_d28) %>%
    mutate(across(c(uva, mews, qsofa), ~as.factor(.))) %>%
  tbl_summary(     
    by = outcome_d28,
    percent = "row",
    digits = all_continuous() ~ 1,
    type   = all_categorical() ~ "categorical"
) %>%
  add_overall(last = T, statistic = ~ "{n}") %>%
      modify_header(update = list(
    label ~ "Variable"
  )
  ) 

tbl_count_alive$table_body$stat_0 <- str_replace_all(tbl_count_alive$table_body$stat_0, " \\(100%\\)", "")

tbl_count_alive <- tbl_count_alive %>%
  as_flex_table() %>%
      flextable::fontsize(part = "header", size = 8) %>%
      flextable::autofit()


tbl_count_alive

# Calculate AUC for each score ----

#> UVA ----

uva_auc_s <- pROC::roc(sensitivity_analysis_alive$outcome_d28,
           sensitivity_analysis_alive$uva, percent = T)

uva_auc_value_s <- round(as.numeric(auc(uva_auc_s)[1])/100, 2)
 
uva_auc_ci_s <- paste0("95% CI: ", round(pROC::ci.auc(uva_auc_s)[1]/100, 2), "-", round(ci.auc(uva_auc_s)[3]/100, 2))

uva_auc_fig_text_s <- paste0("AUC = ", uva_auc_value_s, "\n", uva_auc_ci_s)


#> MEWS ----

mews_auc_s <- pROC::roc(sensitivity_analysis_alive$outcome_d28,
           sensitivity_analysis_alive$mews, percent = T)

mews_auc_value_s <- round(as.numeric(auc(mews_auc_s)[1])/100, 2)
 
mews_auc_ci_s <- paste0("95% CI: ", round(pROC::ci.auc(mews_auc_s)[1]/100, 2), "-", round(ci.auc(mews_auc_s)[3]/100, 2))

mews_auc_fig_text_s <- paste0("AUC = ", mews_auc_value_s, "\n", mews_auc_ci_s)


#> qSOFA ----

qsofa_auc_s <- pROC::roc(sensitivity_analysis_alive$outcome_d28,
           sensitivity_analysis_alive$qsofa, percent = T)

qsofa_auc_value_s <- round(as.numeric(auc(qsofa_auc_s)[1])/100, 2)
 
qsofa_auc_ci_s <- paste0("95% CI: ", round(pROC::ci.auc(qsofa_auc_s)[1]/100, 2), "-", round(ci.auc(qsofa_auc_s)[3]/100, 2))

qsofa_auc_fig_text_s <- paste0("AUC = ", qsofa_auc_value_s, "\n", qsofa_auc_ci_s)


# Create plot for each score ----

#> UVA ----

uva_values_s <- coords(roc = uva_auc_s, x = "all", transpose = FALSE, ret = "all")

uva_values_plot_s <- uva_values_s %>%
  ggplot(mapping = aes(x = specificity, y = sensitivity)) +
  geom_line() +
  geom_point() +
  annotate("text", x = 40, y = 50, label = uva_auc_fig_text_s)  +
  scale_x_reverse("Specificity (%)", limits = c(100, 0),  breaks = scales::breaks_extended(10)) + 
  scale_y_continuous("Sensitivity (%)", limits = c(0, 100), breaks = scales::breaks_extended(10)) + 
  ggtitle("UVA")



#> MEWS ----

mews_values_s <- coords(roc = mews_auc_s, x = "all", transpose = FALSE, ret = "all")

mews_values_plot_s <- mews_values_s %>%
  ggplot(mapping = aes(x = specificity, y = sensitivity)) +
  geom_line() +
  geom_point() +
  annotate("text", x = 40, y = 50, label = mews_auc_fig_text_s)  +
  scale_x_reverse("Specificity (%)", limits = c(100, 0),  breaks = scales::breaks_extended(10)) + 
  scale_y_continuous("Sensitivity (%)", limits = c(0, 100), breaks = scales::breaks_extended(10)) +
  ggtitle("MEWS")



#> qSOFA ----

qsofa_values_s <- coords(roc = qsofa_auc_s, x = "all", transpose = FALSE, ret = "all")

qsofa_values_plot_s <- qsofa_values_s %>%
  ggplot(mapping = aes(x = specificity, y = sensitivity)) +
  geom_line() +
  geom_point() +
  annotate("text", x = 40, y = 50, label = qsofa_auc_fig_text_s)  +
  scale_x_reverse("Specificity (%)", limits = c(100, 0),  breaks = scales::breaks_extended(10)) + 
  scale_y_continuous("Sensitivity (%)", limits = c(0, 100), breaks = scales::breaks_extended(10)) +
  ggtitle("qSOFA")


```

```{r, include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T}

uva_v_qsofa_alive <- roc.test(uva_auc_s, qsofa_auc_s, method = c("delong"))
uva_v_qsofa_alive_z <- as.numeric(uva_v_qsofa_alive$statistic)
uva_v_qsofa_alive_p <- format(uva_v_qsofa_alive$p.value, scientific = F)

uva_v_mews_alive <- roc.test(uva_auc_s, mews_auc_s, method = c("delong"))
uva_v_mews_alive_z <- as.numeric(uva_v_mews_alive$statistic)
uva_v_mews_alive_p <- format(uva_v_mews_alive$p.value, scientific = F)

mews_v_qsofa_alive <- roc.test(mews_auc_s, qsofa_auc_s, method = c("delong"))
mews_v_qsofa_alive_z <- as.numeric(mews_v_qsofa_alive$statistic)
mews_v_qsofa_alive_p <- format(mews_v_qsofa_alive$p.value, scientific = F)

```

DeLong test, a pairwise comparison test for the equality of the area under the curve between for those with day 28 outcome missing set to alive.   
  1) UVA (AUC: `r uva_auc_value_s`, `r uva_auc_ci_s`) vs. MEWS (AUC: `r mews_auc_value_s`,  `r mews_auc_ci_s`), Z = `r uva_v_mews_alive_z`, *p-value* = `r uva_v_mews_alive_p`.   
  2) UVA (AUC: `r uva_auc_value_s`, `r uva_auc_ci_s`) vs. qSOFA (AUC: `r qsofa_auc_value_s`,  `r qsofa_auc_ci_s`), Z = `r uva_v_qsofa_alive_z`, *p-value* = `r uva_v_qsofa_alive_p`.   
  3) MEWs (AUC: `r mews_auc_value_s`,  `r mews_auc_ci_s`) vs. qSOFA (AUC: `r qsofa_auc_value_s`,  `r qsofa_auc_ci_s`), Z = `r mews_v_qsofa_alive_z`, *p-value* = `r mews_v_qsofa_alive_p`.

\newpage
### Figure S2
```{r all_rocs_sensitivity_alive, fig.cap = "Specificity (%) analysis (a), assuming all missing are alive, ROCs for S1) UVA, S2) MEWS, S3) qSOFA ", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T, fig.width = 9, fig.height=6}




combined_alive_s <- cowplot::plot_grid(mews_values_plot_s, qsofa_values_plot_s, uva_values_plot_s, ncol = 3, 
                               labels = c('A', 'B', "C"), label_size = 12, 
                               align = "h")

combined_alive_s  

```

\newpage

### b) Assume all missing died.

```{r, tab.cap = "Specificity (%) analysis (a) assuming all missing a day 28 outcome died", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T }

tbl_count_dead <- sensitivity_analysis_dead %>% 
  select(uva, mews, qsofa, age_cat, sex, outcome_d28) %>%
    mutate(across(c(uva, mews, qsofa), ~as.factor(.))) %>%
  tbl_summary(     
    by = outcome_d28,
    digits = all_continuous() ~ 1,
    type   = all_categorical() ~ "categorical",
      percent = "row"

) %>%
  add_overall(last = T, statistic = ~ "{n}") %>%
      modify_header(update = list(
    label ~ "Variable"
  )
  ) %>%
    as_flex_table() %>%
      flextable::fontsize(part = "header", size = 8) %>%
      flextable::autofit()


tbl_count_dead

# Calculate AUC for each score ----

#> UVA ----

uva_auc_s_d <- pROC::roc(sensitivity_analysis_dead$outcome_d28,
           sensitivity_analysis_dead$uva, percent = T)

uva_auc_value_s_d <- round(as.numeric(auc(uva_auc_s_d)[1])/100, 2)
 
uva_auc_ci_s_d <- paste0("95% CI: ", round(pROC::ci.auc(uva_auc_s_d)[1]/100, 2), "-", round(ci.auc(uva_auc_s_d)[3]/100, 2))

uva_auc_fig_text_s_d <- paste0("AUC = ", uva_auc_value_s_d, "\n", uva_auc_ci_s_d)


#> MEWS ----

mews_auc_s_d <- pROC::roc(sensitivity_analysis_dead$outcome_d28,
           sensitivity_analysis_dead$mews, percent = T)

mews_auc_value_s_d <- round(as.numeric(auc(mews_auc_s_d)[1])/100, 2)
 
mews_auc_ci_s_d <- paste0("95% CI: ", round(pROC::ci.auc(mews_auc_s_d)[1]/100, 2), "-", round(ci.auc(mews_auc_s_d)[3]/100, 2))

mews_auc_fig_text_s_d <- paste0("AUC = ", mews_auc_value_s_d, "\n", mews_auc_ci_s_d)


#> qSOFA ----

qsofa_auc_s_d <- pROC::roc(sensitivity_analysis_dead$outcome_d28,
           sensitivity_analysis_dead$qsofa, percent = T)

qsofa_auc_value_s_d <- round(as.numeric(auc(qsofa_auc_s_d)[1])/100, 2)
 
qsofa_auc_ci_s_d <- paste0("95% CI: ", round(pROC::ci.auc(qsofa_auc_s_d)[1]/100, 2), "-", round(ci.auc(qsofa_auc_s_d)[3]/100, 2))

qsofa_auc_fig_text_s_d <- paste0("AUC = ", qsofa_auc_value_s_d, "\n", qsofa_auc_ci_s_d)


# Create plot for each score ----

#> UVA ----

uva_values_s_d <- coords(roc = uva_auc_s_d, x = "all", transpose = FALSE, ret = "all")

uva_values_plot_s_d <- uva_values_s_d %>%
  ggplot(mapping = aes(x = specificity, y = sensitivity)) +
  geom_line() +
  geom_point() +
  annotate("text", x = 40, y = 50, label = uva_auc_fig_text_s_d)  +
  scale_x_reverse("Specificity (%)", limits = c(100, 0),  breaks = scales::breaks_extended(10)) + 
  scale_y_continuous("Sensitivity (%)", limits = c(0, 100), breaks = scales::breaks_extended(10)) + 
  ggtitle("UVA")



#> MEWS ----

mews_values_s_d <- coords(roc = mews_auc_s_d, x = "all", transpose = FALSE, ret = "all")

mews_values_plot_s_d <- mews_values_s_d %>%
  ggplot(mapping = aes(x = specificity, y = sensitivity)) +
  geom_line() +
  geom_point() +
  annotate("text", x = 35, y = 50, label = mews_auc_fig_text_s_d)  +
  scale_x_reverse("Specificity (%)", limits = c(100, 0),  breaks = scales::breaks_extended(10)) + 
  scale_y_continuous("Sensitivity (%)", limits = c(0, 100), breaks = scales::breaks_extended(10)) +
  ggtitle("MEWS")


#> qSOFA ----

qsofa_values_s_d <- coords(roc = qsofa_auc_s_d, x = "all", transpose = FALSE, ret = "all")

qsofa_values_plot_s_d <- qsofa_values_s_d %>%
  ggplot(mapping = aes(x = specificity, y = sensitivity)) +
  geom_line() +
  geom_point() +
  annotate("text", x = 30, y = 50, label = qsofa_auc_fig_text_s_d)  +
  scale_x_reverse("Specificity (%)", limits = c(100, 0),  breaks = scales::breaks_extended(10)) + 
  scale_y_continuous("Sensitivity (%)", limits = c(0, 100), breaks = scales::breaks_extended(10)) +
  ggtitle("qSOFA")

```

\newpage
### Figure S3
```{r all_rocs_sensitivity_dead, fig.cap = "Specificity (%) analysis (b), assuming all missing day 28 outcomes died, ROCs for S4) UVA, S5) MEWS, S6) qSOFA ", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T, fig.width = 9, fig.height=6}




combined_alive_s_d <- cowplot::plot_grid(mews_values_plot_s_d, qsofa_values_plot_s_d, uva_values_plot_s_d, ncol = 3, 
                               labels = c('A', 'B', "C"), label_size = 12, 
                               align = "h") 


combined_alive_s_d  

```

```{r, include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T}

uva_v_qsofa_dead <- roc.test(uva_auc_s_d, qsofa_auc_s_d, method = c("delong"))
uva_v_qsofa_dead_z <- as.numeric(uva_v_qsofa_dead$statistic)
uva_v_qsofa_dead_p <- format(uva_v_qsofa_dead$p.value, scientific = F)

uva_v_mews_dead <- roc.test(uva_auc_s_d, mews_auc_s_d, method = c("delong"))
uva_v_mews_dead_z <- as.numeric(uva_v_mews_dead$statistic)
uva_v_mews_dead_p <- format(uva_v_mews_dead$p.value, scientific = F)

mews_v_qsofa_dead <- roc.test(mews_auc_s_d, qsofa_auc_s_d, method = c("delong"))
mews_v_qsofa_dead_z <- as.numeric(mews_v_qsofa_dead$statistic)
mews_v_qsofa_dead_p <- format(mews_v_qsofa_dead$p.value, scientific = F)

```

DeLong test, a pairwise comparison test for the equality of the area under the curve between for those with day 28 outcome missing set to dead.   
  1) UVA (AUC: `r uva_auc_value_s_d`, `r uva_auc_ci_s_d`) vs. MEWS (AUC: `r mews_auc_value_s_d`,  `r mews_auc_ci_s_d`), Z = `r uva_v_mews_dead_z`, *p-value* = `r uva_v_mews_dead_p`.   
  2) UVA (AUC: `r uva_auc_value_s_d`, `r uva_auc_ci_s_d`) vs. qSOFA (AUC: `r qsofa_auc_value_s_d`,  `r qsofa_auc_ci_s_d`), Z = `r uva_v_qsofa_dead_z`, *p-value* = `r uva_v_qsofa_dead_p`.   
  3) MEWs (AUC: `r mews_auc_value_s_d`,  `r mews_auc_ci_s_d`) vs. qSOFA (AUC: `r qsofa_auc_value_s_d`,  `r qsofa_auc_ci_s_d`), Z = `r mews_v_qsofa_dead_z`, *p-value* = `r mews_v_qsofa_dead_p`.

\newpage

<!---BLOCK_LANDSCAPE_STOP--->



\newpage

```{r fig_los, fig.cap = "Length of inpatient stay (days) overall and by site", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T}


time2discharge_px_tbl <- severity_data_cc_d28 %>%
  # filter(!is.na(time_to_discharge)) %>%
  select(country, time_to_discharge) %>%
  tbl_summary(
    by = country,
    type = list(all_dichotomous() ~ "categorical",
                all_continuous() ~ "continuous2"
                ),
    statistic = list(c("time_to_discharge") ~ 
                     c("{median} ({p25}, {p75})", "{min}, {max}")
    ),
    digits = all_continuous() ~ 1,
    missing_text = "(Missing)"
      ) %>%
    add_overall() %>%
    add_stat_label() %>%
     modify_header(list(label ~"Variable")) %>%
    as_flex_table() %>%
      flextable::fontsize(part = "header", size = 8) %>%
      flextable::autofit()


time2discharge_px_tbl

```

\newpage
<!---BLOCK_LANDSCAPE_START--->
## Figure S4: ROCs for inpatients
```{r all_rocs_inpatient, fig.cap = "ROCs among inpatients 1) MEWs, 2) qSOFA 3) UVA", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T, fig.width = 9, fig.height=6}

severity_data_cc_d28_inp <- severity_data_cc_d28 %>%
  mutate(px_group = as.character(px_group)) %>%
  filter(str_detect(px_group, "Inpatient"))

# UVA Inpatients

uva_auc_inp <- pROC::roc(severity_data_cc_d28_inp$outcome_d28, severity_data_cc_d28_inp$uva, percent = T)

uva_auc_value_inp <- round(as.numeric(auc(uva_auc_inp)[1])/100, 2)

uva_values_inp <- coords(roc = uva_auc_inp, x = "all", transpose = FALSE, ret = "all")

uva_auc_ci_inp <- paste0("95% CI: ", round(pROC::ci.auc(uva_auc_inp)[1]/100, 2), "-", round(ci.auc(uva_auc_inp)[3]/100, 2))

uva_auc_fig_text_inp <- paste0("AUC = ", uva_auc_value_inp, "\n", uva_auc_ci_inp)

uva_values_plot_inp <- uva_values_inp %>%
  ggplot(mapping = aes(x = specificity, y = sensitivity)) +
  geom_line() +
  geom_point() +
  annotate("text", x = 50, y = 50, label = uva_auc_fig_text_inp) +
  # geom_label(aes(specificity +.5 , sensitivity + 0.5, label = label), 
  #            hjust = 0, nudge_x = 3,
  #            label.size = NA
  #            ) +
  scale_x_reverse("Specificity (%)", limits = c(100, 0),  breaks = scales::breaks_extended(10)) + 
  scale_y_continuous("Sensitivity (%)", limits = c(0, 100), breaks = scales::breaks_extended(10)) 

# MEWS Inpatients

mews_auc_inp <- pROC::roc(severity_data_cc_d28_inp$outcome_d28, 
           severity_data_cc_d28_inp$mews, percent = T)

mews_auc_value_inp <- round(as.numeric(auc(mews_auc_inp)[1])/100, 2)

mews_values_inp <- coords(roc = mews_auc_inp, x = "all", transpose = FALSE, ret = "all")

mews_auc_ci_inp <- paste0("95% CI: ", round(pROC::ci.auc(mews_auc_inp)[1]/100, 2), "-", round(ci.auc(mews_auc_inp)[3]/100, 2))

mews_auc_fig_text_inp <- paste0("AUC = ", mews_auc_value_inp, "\n", mews_auc_ci_inp)

mews_values_plot_inp <- mews_values_inp %>%
  ggplot(mapping = aes(x = specificity, y = sensitivity)) +
  geom_line() +
  geom_point() +
  annotate("text", x = 40, y = 50, label = mews_auc_fig_text_inp) +
  # geom_label(aes(specificity +.5 , sensitivity + 0.5, label = label), 
  #            hjust = 0, nudge_x = 3,
  #            label.size = NA
  #            ) +
  scale_x_reverse("Specificity (%)", limits = c(100, 0),  breaks = scales::breaks_extended(10)) + 
  scale_y_continuous("Sensitivity (%)", limits = c(0, 100), breaks = scales::breaks_extended(10)) 

# qSOFA Inpatients

qsofa_auc_inp <- pROC::roc(severity_data_cc_d28_inp$outcome_d28, 
           severity_data_cc_d28_inp$qsofa, percent = T)

qsofa_auc_value_inp <- round(as.numeric(auc(qsofa_auc_inp)[1])/100, 2)

qsofa_values_inp <- coords(roc = qsofa_auc_inp, x = "all", transpose = FALSE, ret = "all")

qsofa_auc_ci_inp <- paste0("95% CI: ", round(pROC::ci.auc(qsofa_auc_inp)[1]/100, 2), "-", round(ci.auc(qsofa_auc_inp)[3]/100, 2))

qsofa_auc_fig_text_inp <- paste0("AUC = ", qsofa_auc_value_inp, "\n", qsofa_auc_ci_inp)

qsofa_values_plot_inp <- qsofa_values_inp %>%
  ggplot(mapping = aes(x = specificity, y = sensitivity)) +
  geom_line() +
  geom_point() +
  annotate("text", x = 40, y = 50, label = qsofa_auc_fig_text_inp) +
  # geom_label(aes(specificity +.5 , sensitivity + 0.5, label = label), 
  #            hjust = 0, nudge_x = 3,
  #            label.size = NA
  #            ) +
  scale_x_reverse("Specificity (%)", limits = c(100, 0),  breaks = scales::breaks_extended(10)) + 
  scale_y_continuous("Sensitivity (%)", limits = c(0, 100), breaks = scales::breaks_extended(10)) 


uva_values_plot_inp$layers[[4]] <- NULL

uva_values_plot_inp <- uva_values_plot_inp +
  ggtitle("UVA")

mews_values_plot_inp$layers[[4]] <- NULL

mews_values_plot_inp <- mews_values_plot_inp +
  ggtitle("MEWS")

qsofa_values_plot_inp$layers[[4]] <- NULL

qsofa_values_plot_inp <- qsofa_values_plot_inp +
  ggtitle("qSOFA")

combined_inp <- cowplot::plot_grid(mews_values_plot_inp, qsofa_values_plot_inp, uva_values_plot_inp, ncol = 3, 
                               labels = c('A', 'B', "C"), label_size = 12, 
                               align = "h")

combined_inp

```

```{r, include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T}

uva_v_qsofa_inp <- roc.test(uva_auc_inp, qsofa_auc_inp, method = c("delong"))
uva_v_qsofa_inp_z <- as.numeric(uva_v_qsofa_inp$statistic)
uva_v_qsofa_inp_p <- format(uva_v_qsofa_inp$p.value, scientific = F)

uva_v_mews_inp <- roc.test(uva_auc_inp, mews_auc_inp, method = c("delong"))
uva_v_mews_inp_z <- as.numeric(uva_v_mews_inp$statistic)
uva_v_mews_inp_p <- format(uva_v_mews_inp$p.value, scientific = F)

mews_v_qsofa_inp <- roc.test(mews_auc_inp, qsofa_auc_inp, method = c("delong"))
mews_v_qsofa_inp_z <- as.numeric(mews_v_qsofa_inp$statistic)
mews_v_qsofa_inp_p <- format(mews_v_qsofa_inp$p.value, scientific = F)

```

DeLong test, for inpatients.   
  1) UVA (AUC: `r uva_auc_value_inp`, `r uva_auc_ci_inp`) vs. MEWS (AUC: `r mews_auc_value_inp`,  `r mews_auc_ci_inp`), Z = `r uva_v_mews_inp_z`, *p-value* = `r uva_v_mews_inp_p`.   
  2) UVA (AUC: `r uva_auc_value_inp`, `r uva_auc_ci_inp`) vs. qSOFA (AUC: `r qsofa_auc_value_inp`,  `r qsofa_auc_ci_inp`), Z = `r uva_v_qsofa_inp_z`, *p-value* = `r uva_v_qsofa_inp_p`.   
  3) MEWs (AUC: `r mews_auc_value_inp`,  `r mews_auc_ci_inp`) vs. qSOFA (AUC: `r qsofa_auc_value_inp`,  `r qsofa_auc_ci_inp`), Z = `r mews_v_qsofa_inp_z`, *p-value* = `r mews_v_qsofa_inp_p`.


  
\newpage
## Figure S5: ROCs for outpatients

```{r all_rocs_outpatients, fig.cap = "ROCs among outpatients for 1) MEWs, 2) qSOFA 3) UVA", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T, fig.width = 9, fig.height=6}

severity_data_cc_d28_op <- severity_data_cc_d28 %>%
  mutate(px_group = as.character(px_group)) %>%
  filter(str_detect(px_group, "Outpatient"))

# UVA outpatients

uva_auc_op <- pROC::roc(severity_data_cc_d28_op$outcome_d28, severity_data_cc_d28_op$uva, percent = T)

uva_auc_value_op <- round(as.numeric(auc(uva_auc_op)[1])/100, 2)

uva_values_op <- coords(roc = uva_auc_op, x = "all", transpose = FALSE, ret = "all")

uva_auc_ci_op <- paste0("95% CI: ", round(pROC::ci.auc(uva_auc_op)[1]/100, 2), "-", round(ci.auc(uva_auc_op)[3]/100, 2))

uva_auc_fig_text_op <- paste0("AUC = ", uva_auc_value_op, "\n", uva_auc_ci_op)

uva_values_plot_op <- uva_values_op %>%
  ggplot(mapping = aes(x = specificity, y = sensitivity)) +
  geom_line() +
  geom_point() +
  annotate("text", x = 40, y = 50, label = uva_auc_fig_text_op) +
  # geom_label(aes(specificity +.5 , sensitivity + 0.5, label = label), 
  #            hjust = 0, nudge_x = 3,
  #            label.size = NA
  #            ) +
  scale_x_reverse("Specificity (%)", limits = c(100, 0),  breaks = scales::breaks_extended(10)) + 
  scale_y_continuous("Sensitivity (%)", limits = c(0, 100), breaks = scales::breaks_extended(10)) 

# MEWS outpatients

mews_auc_op <- pROC::roc(severity_data_cc_d28_op$outcome_d28, 
           severity_data_cc_d28_op$mews, percent = T)

mews_auc_value_op <- round(as.numeric(auc(mews_auc_op)[1])/100, 2)

mews_values_op <- coords(roc = mews_auc_op, x = "all", transpose = FALSE, ret = "all")

mews_auc_ci_op <- paste0("95% CI: ", round(pROC::ci.auc(mews_auc_op)[1]/100, 2), "-", round(ci.auc(mews_auc_op)[3]/100, 2))

mews_auc_fig_text_op <- paste0("AUC = ", mews_auc_value_op, "\n", mews_auc_ci_op)

mews_values_plot_op <- mews_values_op %>%
  ggplot(mapping = aes(x = specificity, y = sensitivity)) +
  geom_line() +
  geom_point() +
  annotate("text", x = 29, y = 50, label = mews_auc_fig_text_op) +
  # geom_label(aes(specificity +.5 , sensitivity + 0.5, label = label), 
  #            hjust = 0, nudge_x = 3,
  #            label.size = NA
  #            ) +
  scale_x_reverse("Specificity (%)", limits = c(100, 0),  breaks = scales::breaks_extended(10)) + 
  scale_y_continuous("Sensitivity (%)", limits = c(0, 100), breaks = scales::breaks_extended(10)) 

# qSOFA outpatients

qsofa_auc_op <- pROC::roc(severity_data_cc_d28_op$outcome_d28, 
           severity_data_cc_d28_op$qsofa, percent = T)

qsofa_auc_value_op <- round(as.numeric(auc(qsofa_auc_op)[1])/100, 2)

qsofa_values_op <- coords(roc = qsofa_auc_op, x = "all", transpose = FALSE, ret = "all")

qsofa_auc_ci_op <- paste0("95% CI: ", round(pROC::ci.auc(qsofa_auc_op)[1]/100, 2), "-", round(ci.auc(qsofa_auc_op)[3]/100, 2))

qsofa_auc_fig_text_op <- paste0("AUC = ", qsofa_auc_value_op, "\n", qsofa_auc_ci_op)

qsofa_values_plot_op <- qsofa_values_op %>%
  ggplot(mapping = aes(x = specificity, y = sensitivity)) +
  geom_line() +
  geom_point() +
  annotate("text", x = 29, y = 50, label = qsofa_auc_fig_text_op) +
  # geom_label(aes(specificity +.5 , sensitivity + 0.5, label = label), 
  #            hjust = 0, nudge_x = 3,
  #            label.size = NA
  #            ) +
  scale_x_reverse("Specificity (%)", limits = c(100, 0),  breaks = scales::breaks_extended(10)) + 
  scale_y_continuous("Sensitivity (%)", limits = c(0, 100), breaks = scales::breaks_extended(10)) 


uva_values_plot_op$layers[[4]] <- NULL

uva_values_plot_op <- uva_values_plot_op +
  ggtitle("UVA")

mews_values_plot_op$layers[[4]] <- NULL

mews_values_plot_op <- mews_values_plot_op +
  ggtitle("MEWS")

qsofa_values_plot_op$layers[[4]] <- NULL

qsofa_values_plot_op <- qsofa_values_plot_op +
  ggtitle("qSOFA")

combined_op <- cowplot::plot_grid(mews_values_plot_op, qsofa_values_plot_op, uva_values_plot_op, ncol = 3, 
                               labels = c('A', 'B', "C"), label_size = 12, 
                               align = "h")

combined_op

```


```{r, include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T }

uva_v_uva_inp <- roc.test(uva_auc_op, mews_auc_op, method = c("delong"))
uva_v_uva_inp_z <- as.numeric(uva_v_uva_inp$statistic)
uva_v_uva_inp_p <- format(uva_v_uva_inp$p.value, scientific = F)

mews_v_mews_inp <- roc.test(uva_auc_op, qsofa_auc_op, method = c("delong"))
mews_v_mews_inp_z <- as.numeric(mews_v_mews_inp$statistic)
mews_v_mews_inp_p <- format(mews_v_mews_inp$p.value, scientific = F)

qsofa_v_qsofa_inp <- roc.test(mews_auc_op, qsofa_auc_op, method = c("delong"))
qsofa_v_qsofa_inp_z <- as.numeric(qsofa_v_qsofa_inp$statistic)
qsofa_v_qsofa_inp_p <- format(qsofa_v_qsofa_inp$p.value, scientific = F)

```


DeLong test, for outpatients.      
     1) UVA (AUC: `r uva_auc_value_op`, `r uva_auc_ci_op`) vs. MEWS (AUC: `r mews_auc_value_op`,  `r mews_auc_ci_op`), Z = `r uva_v_uva_inp_z`, *p-value* = `r uva_v_uva_inp_p`.   
  2) UVA (AUC: `r uva_auc_value_op`, `r uva_auc_ci_op`) vs. qSOFA (AUC: `r qsofa_auc_value_op`,  `r qsofa_auc_ci_op`), Z = `r mews_v_mews_inp_z`, *p-value* = `r mews_v_mews_inp_p`.   
  3) MEWs (AUC: `r mews_auc_value_op`,  `r mews_auc_ci_op`) vs. qSOFA (AUC: `r qsofa_auc_value_op`,  `r qsofa_auc_ci_op`), Z = `r qsofa_v_qsofa_inp_z`, *p-value* = `r qsofa_v_qsofa_inp_p`.




\newpage




## Figure S6: Laos ROCs
```{r la_rocs, fig.cap = "Laos ROCs for 1) MEWs, 2) qSOFA, 3) UVA", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T, fig.width = 8, fig.height=5}

severity_data_cc_d28_la <- severity_data_cc_d28 %>%
  filter(as.character(country) == "Laos")

str(severity_data_cc_d28$outcome_d28)
unique(severity_data_cc_d28$outcome_d28)


# UVA Laos

# roc(severity_data_cc_d28_la$outcome_d28, severity_data_cc_d28_la$uva, percent = T, levels = c("Alive", "Dead"))

uva_auc_la <- pROC::roc(severity_data_cc_d28_la$outcome_d28, severity_data_cc_d28_la$uva, percent = T)

uva_auc_value_la <- round(as.numeric(auc(uva_auc_la)[1])/100, 2)

uva_values_la <- coords(roc = uva_auc_la, x = "all", transpose = FALSE, ret = "all")

uva_auc_ci_la <- paste0("95% CI: ", round(pROC::ci.auc(uva_auc_la)[1]/100, 2), "-", round(ci.auc(uva_auc_la)[3]/100, 2))

uva_auc_fig_text_la <- paste0("AUC = ", uva_auc_value_la, "\n", uva_auc_ci_la)

uva_values_plot_la <- uva_values_la %>%
  ggplot(mapping = aes(x = specificity, y = sensitivity)) +
  geom_line() +
  geom_point() +
  annotate("text", x = 50, y = 50, label = uva_auc_fig_text_la) +
  # geom_label(aes(specificity +.5 , sensitivity + 0.5, label = label), 
  #            hjust = 0, nudge_x = 3,
  #            label.size = NA
  #            ) +
  scale_x_reverse("Specificity (%)", limits = c(100, 0),  breaks = scales::breaks_extended(10)) + 
  scale_y_continuous("Sensitivity (%)", limits = c(0, 100), breaks = scales::breaks_extended(10)) 

# MEWS Laos

mews_auc_la <- pROC::roc(severity_data_cc_d28_la$outcome_d28, 
           severity_data_cc_d28_la$mews, percent = T)

mews_auc_value_la <- round(as.numeric(auc(mews_auc_la)[1])/100, 2)

mews_values_la <- coords(roc = mews_auc_la, x = "all", transpose = FALSE, ret = "all")

mews_auc_ci_la <- paste0("95% CI: ", round(pROC::ci.auc(mews_auc_la)[1]/100, 2), "-", round(ci.auc(mews_auc_la)[3]/100, 2))

mews_auc_fig_text_la <- paste0("AUC = ", mews_auc_value_la, "\n", mews_auc_ci_la)

mews_values_plot_la <- mews_values_la %>%
  ggplot(mapping = aes(x = specificity, y = sensitivity)) +
  geom_line() +
  geom_point() +
  annotate("text", x = 30, y = 50, label = mews_auc_fig_text_la) +
  # geom_label(aes(specificity +.5 , sensitivity + 0.5, label = label), 
  #            hjust = 0, nudge_x = 3,
  #            label.size = NA
  #            ) +
  scale_x_reverse("Specificity (%)", limits = c(100, 0),  breaks = scales::breaks_extended(10)) + 
  scale_y_continuous("Sensitivity (%)", limits = c(0, 100), breaks = scales::breaks_extended(10)) 

# qSOFA Laos

qsofa_auc_la <- pROC::roc(severity_data_cc_d28_la$outcome_d28, 
           severity_data_cc_d28_la$qsofa, percent = T)

qsofa_auc_value_la <- round(as.numeric(auc(qsofa_auc_la)[1])/100, 2)

qsofa_values_la <- coords(roc = qsofa_auc_la, x = "all", transpose = FALSE, ret = "all")

qsofa_auc_ci_la <- paste0("95% CI: ", round(pROC::ci.auc(qsofa_auc_la)[1]/100, 2), "-", round(ci.auc(qsofa_auc_la)[3]/100, 2))

qsofa_auc_fig_text_la <- paste0("AUC = ", qsofa_auc_value_la, "\n", qsofa_auc_ci_la)

qsofa_values_plot_la <- qsofa_values_la %>%
  ggplot(mapping = aes(x = specificity, y = sensitivity)) +
  geom_line() +
  geom_point() +
  annotate("text", x = 50, y = 50, label = qsofa_auc_fig_text_la) +
  # geom_label(aes(specificity +.5 , sensitivity + 0.5, label = label), 
  #            hjust = 0, nudge_x = 3,
  #            label.size = NA
  #            ) +
  scale_x_reverse("Specificity (%)", limits = c(100, 0),  breaks = scales::breaks_extended(10)) + 
  scale_y_continuous("Sensitivity (%)", limits = c(0, 100), breaks = scales::breaks_extended(10)) 


uva_values_plot_la$layers[[4]] <- NULL

uva_values_plot_la <- uva_values_plot_la +
  ggtitle("UVA")

mews_values_plot_la$layers[[4]] <- NULL

mews_values_plot_la <- mews_values_plot_la +
  ggtitle("MEWS")

qsofa_values_plot_la$layers[[4]] <- NULL

qsofa_values_plot_la <- qsofa_values_plot_la +
  ggtitle("qSOFA")

combined_la <- cowplot::plot_grid(mews_values_plot_la, qsofa_values_plot_la, uva_values_plot_la, ncol = 3, 
                               labels = c('A', 'B', "C"), label_size = 12, 
                               align = "h")

combined_la


```


```{r, include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T }

la_uva_v_qsofa <- roc.test(uva_auc_la, qsofa_auc_la, method = c("delong"))
la_uva_v_qsofa_z <- as.numeric(la_uva_v_qsofa$statistic)
la_uva_v_qsofa_p <- format(la_uva_v_qsofa$p.value, scientific = F)

la_uva_v_mews <- roc.test(uva_auc_la, mews_auc_la, method = c("delong"))
la_uva_v_mews_z <- as.numeric(la_uva_v_mews$statistic)
la_uva_v_mews_p <- format(la_uva_v_mews$p.value, scientific = F)

la_mews_v_qsofa <- roc.test(mews_auc_la, qsofa_auc_la, method = c("delong"))
la_mews_v_qsofa_z <- as.numeric(la_mews_v_qsofa$statistic)
la_mews_v_qsofa_p <- format(la_mews_v_qsofa$p.value, scientific = F)

```

DeLong test for Laos site;   
  1) UVA (AUC:`r uva_auc_value_la`, `r uva_auc_ci_la`) vs. MEWS (AUC:`r mews_auc_value_la`,  `r mews_auc_ci_la`), Z = `r la_uva_v_qsofa_z`, *p-value* = `r la_uva_v_mews_p`.   
  2) UVA (AUC:`r uva_auc_value_la`, `r uva_auc_ci_la`) vs. qSOFA (AUC:`r qsofa_auc_value_la`,  `r qsofa_auc_ci_la`), Z = `r la_uva_v_qsofa_z`, *p-value* = `r la_uva_v_qsofa_p`.   
  3) MEWs (AUC:`r mews_auc_value_la`,  `r mews_auc_ci_la`) vs. qSOFA (AUC:`r qsofa_auc_value_la`,  `r qsofa_auc_ci_la`), Z = `r la_mews_v_qsofa_z`, *p-value* = `r la_mews_v_qsofa_p`.   
  
\newpage

## Figure S7: Malawi ROCs

```{r mw_rocs, fig.cap = "Malawi ROCs for 1) MEWs, 2) qSOFA, 3) UVA", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T, fig.width = 8, fig.height=5}

severity_data_cc_d28_mw <- severity_data_cc_d28 %>%
  filter(as.character(country) == "Malawi")

# UVA Malawi

uva_auc_mw <- pROC::roc(severity_data_cc_d28_mw$outcome_d28, severity_data_cc_d28_mw$uva, percent = T)

uva_auc_value_mw <- round(as.numeric(auc(uva_auc_mw)[1])/100, 2)

uva_values_mw <- coords(roc = uva_auc_mw, x = "all", transpose = FALSE, ret = "all")

uva_auc_ci_mw <- paste0("95% CI: ", round(pROC::ci.auc(uva_auc_mw)[1]/100, 2), "-", round(ci.auc(uva_auc_mw)[3]/100, 2))

uva_auc_fig_text_mw <- paste0("AUC = ", uva_auc_value_mw, "\n", uva_auc_ci_mw)

uva_values_plot_mw <- uva_values_mw %>%
  ggplot(mapping = aes(x = specificity, y = sensitivity)) +
  geom_line() +
  geom_point() +
  annotate("text", x = 50, y = 50, label = uva_auc_fig_text_mw) +
  scale_x_reverse("Specificity (%)", limits = c(100, 0),  breaks = scales::breaks_extended(10)) + 
  scale_y_continuous("Sensitivity (%)", limits = c(0, 100), breaks = scales::breaks_extended(10)) 

# MEWS Malawi

mews_auc_mw <- pROC::roc(severity_data_cc_d28_mw$outcome_d28, 
           severity_data_cc_d28_mw$mews, percent = T)

mews_auc_value_mw <- round(as.numeric(auc(mews_auc_mw)[1])/100, 2)

mews_values_mw <- coords(roc = mews_auc_mw, x = "all", transpose = FALSE, ret = "all")

mews_auc_ci_mw <- paste0("95% CI: ", round(pROC::ci.auc(mews_auc_mw)[1]/100, 2), "-", round(ci.auc(mews_auc_mw)[3]/100, 2))

mews_auc_fig_text_mw <- paste0("AUC = ", mews_auc_value_mw, "\n", mews_auc_ci_mw)

mews_values_plot_mw <- mews_values_mw %>%
  ggplot(mapping = aes(x = specificity, y = sensitivity)) +
  geom_line() +
  geom_point() +
  annotate("text", x = 50, y = 50, label = mews_auc_fig_text_mw) +
  # geom_mwbel(aes(specificity +.5 , sensitivity + 0.5, label = label), 
  #            hjust = 0, nudge_x = 3,
  #            label.size = NA
  #            ) +
  scale_x_reverse("Specificity (%)", limits = c(100, 0),  breaks = scales::breaks_extended(10)) + 
  scale_y_continuous("Sensitivity (%)", limits = c(0, 100), breaks = scales::breaks_extended(10)) 

# qSOFA Malawi

qsofa_auc_mw <- pROC::roc(severity_data_cc_d28_mw$outcome_d28, 
           severity_data_cc_d28_mw$qsofa, percent = T)

qsofa_auc_value_mw <- round(as.numeric(auc(qsofa_auc_mw)[1])/100, 2)

qsofa_values_mw <- coords(roc = qsofa_auc_mw, x = "all", transpose = FALSE, ret = "all")

qsofa_auc_ci_mw <- paste0("95% CI: ", round(pROC::ci.auc(qsofa_auc_mw)[1]/100, 2), "-", round(ci.auc(qsofa_auc_mw)[3]/100, 2))

qsofa_auc_fig_text_mw <- paste0("AUC = ", qsofa_auc_value_mw, "\n", qsofa_auc_ci_mw)

qsofa_values_plot_mw <- qsofa_values_mw %>%
  ggplot(mapping = aes(x = specificity, y = sensitivity)) +
  geom_line() +
  geom_point() +
  annotate("text", x = 50, y = 50, label = qsofa_auc_fig_text_mw) +
  # geom_mwbel(aes(specificity +.5 , sensitivity + 0.5, label = label), 
  #            hjust = 0, nudge_x = 3,
  #            label.size = NA
  #            ) +
  scale_x_reverse("Specificity (%)", limits = c(100, 0),  breaks = scales::breaks_extended(10)) + 
  scale_y_continuous("Sensitivity (%)", limits = c(0, 100), breaks = scales::breaks_extended(10)) 


uva_values_plot_mw$layers[[4]] <- NULL

uva_values_plot_mw <- uva_values_plot_mw +
  ggtitle("UVA")

mews_values_plot_mw$layers[[4]] <- NULL

mews_values_plot_mw <- mews_values_plot_mw +
  ggtitle("MEWS")

qsofa_values_plot_mw$layers[[4]] <- NULL

qsofa_values_plot_mw <- qsofa_values_plot_mw +
  ggtitle("qSOFA")

combined_mw <- cowplot::plot_grid(mews_values_plot_mw, qsofa_values_plot_mw, uva_values_plot_mw, ncol = 3, 
                               labels = c('A', 'B', "C"), label_size = 12, 
                               align = "h")

combined_mw


```


```{r, include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T }

mw_uva_v_qsofa <- roc.test(uva_auc_mw, qsofa_auc_mw, method = c("delong"))
mw_uva_v_qsofa_z <- as.numeric(mw_uva_v_qsofa$statistic)
mw_uva_v_qsofa_p <- format(mw_uva_v_qsofa$p.value, scientific = F)

mw_uva_v_mews <- roc.test(uva_auc_mw, mews_auc_mw, method = c("delong"))
mw_uva_v_mews_z <- as.numeric(mw_uva_v_mews$statistic)
mw_uva_v_mews_p <- format(mw_uva_v_mews$p.value, scientific = F)

mw_mews_v_qsofa <- roc.test(mews_auc_mw, qsofa_auc_mw, method = c("delong"))
mw_mews_v_qsofa_z <- as.numeric(mw_mews_v_qsofa$statistic)
mw_mews_v_qsofa_p <- format(mw_mews_v_qsofa$p.value, scientific = F)

```

DeLong test for Malawi site;   
  1) UVA (AUC:`r uva_auc_value_mw`, `r uva_auc_ci_mw`) vs. MEWS (AUC:`r mews_auc_value_mw`,  `r mews_auc_ci_mw`), Z = `r mw_uva_v_mews_z`, *p-value* = `r mw_uva_v_mews_p`.   
  2) UVA (AUC:`r uva_auc_value_mw`, `r uva_auc_ci_mw`) vs. qSOFA (AUC:`r qsofa_auc_value_mw`,  `r qsofa_auc_ci_mw`), Z = `r mw_uva_v_qsofa_z`, *p-value* = `r mw_uva_v_qsofa_p`.   
  3) MEWs (AUC:`r mews_auc_value_mw`,  `r mews_auc_ci_mw`) vs. qSOFA (AUC:`r qsofa_auc_value_mw`,  `r qsofa_auc_ci_mw`), Z = `r mw_mews_v_qsofa_z`, *p-value* = `r mw_mews_v_qsofa_p`.   
  
\newpage

## Figure S8: Mozambique ROCs

```{r mz_rocs, fig.cap = "Mozambique ROCs for 1) MEWs, 2) qSOFA, 3) UVA", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T, fig.width = 8, fig.height=5}

severity_data_cc_d28_mz <- severity_data_cc_d28 %>%
  filter(as.character(country) == "Mozambique")

# UVA Mozambique

uva_auc_mz <- pROC::roc(severity_data_cc_d28_mz$outcome_d28, severity_data_cc_d28_mz$uva, percent = T)

uva_auc_value_mz <- round(as.numeric(auc(uva_auc_mz)[1])/100, 2)

uva_values_mz <- coords(roc = uva_auc_mz, x = "all", transpose = FALSE, ret = "all")

uva_auc_ci_mz <- paste0("95% CI: ", round(pROC::ci.auc(uva_auc_mz)[1]/100, 2), "-", round(ci.auc(uva_auc_mz)[3]/100, 2))

uva_auc_fig_text_mz <- paste0("AUC = ", uva_auc_value_mz, "\n", uva_auc_ci_mz)

uva_values_plot_mz <- uva_values_mz %>%
  ggplot(mapping = aes(x = specificity, y = sensitivity)) +
  geom_line() +
  geom_point() +
  annotate("text", x = 50, y = 50, label = uva_auc_fig_text_mz) +
  # geom_mzbel(aes(specificity +.5 , sensitivity + 0.5, label = label), 
  #            hjust = 0, nudge_x = 3,
  #            label.size = NA
  #            ) +
  scale_x_reverse("Specificity (%)", limits = c(100, 0),  breaks = scales::breaks_extended(10)) + 
  scale_y_continuous("Sensitivity (%)", limits = c(0, 100), breaks = scales::breaks_extended(10)) 

# MEWS Mozambique

mews_auc_mz <- pROC::roc(severity_data_cc_d28_mz$outcome_d28, 
           severity_data_cc_d28_mz$mews, percent = T)

mews_auc_value_mz <- round(as.numeric(auc(mews_auc_mz)[1])/100, 2)

mews_values_mz <- coords(roc = mews_auc_mz, x = "all", transpose = FALSE, ret = "all")

mews_auc_ci_mz <- paste0("95% CI: ", round(pROC::ci.auc(mews_auc_mz)[1]/100, 2), "-", round(ci.auc(mews_auc_mz)[3]/100, 2))

mews_auc_fig_text_mz <- paste0("AUC = ", mews_auc_value_mz, "\n", mews_auc_ci_mz)

mews_values_plot_mz <- mews_values_mz %>%
  ggplot(mapping = aes(x = specificity, y = sensitivity)) +
  geom_line() +
  geom_point() +
  annotate("text", x = 50, y = 50, label = mews_auc_fig_text_mz) +
  # geom_mzbel(aes(specificity +.5 , sensitivity + 0.5, label = label), 
  #            hjust = 0, nudge_x = 3,
  #            label.size = NA
  #            ) +
  scale_x_reverse("Specificity (%)", limits = c(100, 0),  breaks = scales::breaks_extended(10)) + 
  scale_y_continuous("Sensitivity (%)", limits = c(0, 100), breaks = scales::breaks_extended(10)) 

# qSOFA Mozambique

qsofa_auc_mz <- pROC::roc(severity_data_cc_d28_mz$outcome_d28, 
           severity_data_cc_d28_mz$qsofa, percent = T)

qsofa_auc_value_mz <- round(as.numeric(auc(qsofa_auc_mz)[1])/100, 2)

qsofa_values_mz <- coords(roc = qsofa_auc_mz, x = "all", transpose = FALSE, ret = "all")

qsofa_auc_ci_mz <- paste0("95% CI: ", round(pROC::ci.auc(qsofa_auc_mz)[1]/100, 2), "-", round(ci.auc(qsofa_auc_mz)[3]/100, 2))

qsofa_auc_fig_text_mz <- paste0("AUC = ", qsofa_auc_value_mz, "\n", qsofa_auc_ci_mz)

qsofa_values_plot_mz <- qsofa_values_mz %>%
  ggplot(mapping = aes(x = specificity, y = sensitivity)) +
  geom_line() +
  geom_point() +
  annotate("text", x = 50, y = 50, label = qsofa_auc_fig_text_mz) +
  # geom_mzbel(aes(specificity +.5 , sensitivity + 0.5, label = label), 
  #            hjust = 0, nudge_x = 3,
  #            label.size = NA
  #            ) +
  scale_x_reverse("Specificity (%)", limits = c(100, 0),  breaks = scales::breaks_extended(10)) + 
  scale_y_continuous("Sensitivity (%)", limits = c(0, 100), breaks = scales::breaks_extended(10)) 


uva_values_plot_mz$layers[[4]] <- NULL

uva_values_plot_mz <- uva_values_plot_mz +
  ggtitle("UVA")

mews_values_plot_mz$layers[[4]] <- NULL

mews_values_plot_mz <- mews_values_plot_mz +
  ggtitle("MEWS")

qsofa_values_plot_mz$layers[[4]] <- NULL

qsofa_values_plot_mz <- qsofa_values_plot_mz +
  ggtitle("qSOFA")

combined_mz <- cowplot::plot_grid(mews_values_plot_mz, qsofa_values_plot_mz, uva_values_plot_mz, ncol = 3, 
                               labels = c('A', 'B', "C"), label_size = 12, 
                               align = "h")

combined_mz


```


```{r, include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T }

mz_uva_v_qsofa <- roc.test(uva_auc_mz, qsofa_auc_mz, method = c("delong"))
mz_uva_v_qsofa_z <- as.numeric(mz_uva_v_qsofa$statistic)
mz_uva_v_qsofa_p <- format(mz_uva_v_qsofa$p.value, scientific = F)

mz_uva_v_mews <- roc.test(uva_auc_mz, mews_auc_mz, method = c("delong"))
mz_uva_v_mews_z <- as.numeric(mz_uva_v_mews$statistic)
mz_uva_v_mews_p <- format(mz_uva_v_mews$p.value, scientific = F)

mz_mews_v_qsofa <- roc.test(mews_auc_mz, qsofa_auc_mz, method = c("delong"))
mz_mews_v_qsofa_z <- as.numeric(mz_mews_v_qsofa$statistic)
mz_mews_v_qsofa_p <- format(mz_mews_v_qsofa$p.value, scientific = F)

```

DeLong test for Mozambique site;   
  1) UVA (AUC:`r uva_auc_value_mz`, `r uva_auc_ci_mz`) vs. MEWS (AUC:`r mews_auc_value_mz`,  `r mews_auc_ci_mz`), Z = `r mz_uva_v_mews_z`, *p-value* = `r mz_uva_v_mews_p`.   
  2) UVA (AUC:`r uva_auc_value_mz`, `r uva_auc_ci_mz`) vs. qSOFA (AUC:`r qsofa_auc_value_mz`,  `r qsofa_auc_ci_mz`), Z = `r mz_uva_v_qsofa_z`, *p-value* = `r mz_uva_v_qsofa_p`.   
  3) MEWs (AUC:`r mews_auc_value_mz`,  `r mews_auc_ci_mz`) vs. qSOFA (AUC:`r qsofa_auc_value_mz`,  `r qsofa_auc_ci_mz`), Z = `r mz_mews_v_qsofa_z`, *p-value* = `r mz_mews_v_qsofa_p`.   

\newpage
## Figure S9: Zimbabwe ROCs

```{r zw_rocs, fig.cap = "Zimbabwe ROCs for 1) MEWs, 2) qSOFA, 3) UVA", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T, fig.width = 8, fig.height=5}

severity_data_cc_d28_zw <- severity_data_cc_d28 %>%
  filter(as.character(country) == "Zimbabwe")

# UVA Zimbabwe

uva_auc_zw <- pROC::roc(severity_data_cc_d28_zw$outcome_d28, severity_data_cc_d28_zw$uva, percent = T)

uva_auc_value_zw <- round(as.numeric(auc(uva_auc_zw)[1])/100, 2)

uva_values_zw <- coords(roc = uva_auc_zw, x = "all", transpose = FALSE, ret = "all")

uva_auc_ci_zw <- paste0("95% CI: ", round(pROC::ci.auc(uva_auc_zw)[1]/100, 2), "-", round(ci.auc(uva_auc_zw)[3]/100, 2))

uva_auc_fig_text_zw <- paste0("AUC = ", uva_auc_value_zw, "\n", uva_auc_ci_zw)

uva_values_plot_zw <- uva_values_zw %>%
  ggplot(mapping = aes(x = specificity, y = sensitivity)) +
  geom_line() +
  geom_point() +
  annotate("text", x = 50, y = 50, label = uva_auc_fig_text_zw) +
  # geom_zwbel(aes(specificity +.5 , sensitivity + 0.5, label = label), 
  #            hjust = 0, nudge_x = 3,
  #            label.size = NA
  #            ) +
  scale_x_reverse("Specificity (%)", limits = c(100, 0),  breaks = scales::breaks_extended(10)) + 
  scale_y_continuous("Sensitivity (%)", limits = c(0, 100), breaks = scales::breaks_extended(10)) 

# MEWS Zimbabwe

mews_auc_zw <- pROC::roc(severity_data_cc_d28_zw$outcome_d28, 
           severity_data_cc_d28_zw$mews, percent = T)

mews_auc_value_zw <- round(as.numeric(auc(mews_auc_zw)[1])/100, 2)

mews_values_zw <- coords(roc = mews_auc_zw, x = "all", transpose = FALSE, ret = "all")

mews_auc_ci_zw <- paste0("95% CI: ", round(pROC::ci.auc(mews_auc_zw)[1]/100, 2), "-", round(ci.auc(mews_auc_zw)[3]/100, 2))

mews_auc_fig_text_zw <- paste0("AUC = ", mews_auc_value_zw, "\n", mews_auc_ci_zw)

mews_values_plot_zw <- mews_values_zw %>%
  ggplot(mapping = aes(x = specificity, y = sensitivity)) +
  geom_line() +
  geom_point() +
  annotate("text", x = 35, y = 50, label = mews_auc_fig_text_zw) +
  # geom_zwbel(aes(specificity +.5 , sensitivity + 0.5, label = label), 
  #            hjust = 0, nudge_x = 3,
  #            label.size = NA
  #            ) +
  scale_x_reverse("Specificity (%)", limits = c(100, 0),  breaks = scales::breaks_extended(10)) + 
  scale_y_continuous("Sensitivity (%)", limits = c(0, 100), breaks = scales::breaks_extended(10)) 

# qSOFA Zimbabwe

qsofa_auc_zw <- pROC::roc(severity_data_cc_d28_zw$outcome_d28, 
           severity_data_cc_d28_zw$qsofa, percent = T)

qsofa_auc_value_zw <- round(as.numeric(auc(qsofa_auc_zw)[1])/100, 2)

qsofa_values_zw <- coords(roc = qsofa_auc_zw, x = "all", transpose = FALSE, ret = "all")

qsofa_auc_ci_zw <- paste0("95% CI: ", round(pROC::ci.auc(qsofa_auc_zw)[1]/100, 2), "-", round(ci.auc(qsofa_auc_zw)[3]/100, 2))

qsofa_auc_fig_text_zw <- paste0("AUC = ", qsofa_auc_value_zw, "\n", qsofa_auc_ci_zw)

qsofa_values_plot_zw <- qsofa_values_zw %>%
  ggplot(mapping = aes(x = specificity, y = sensitivity)) +
  geom_line() +
  geom_point() +
  annotate("text", x = 35, y = 50, label = qsofa_auc_fig_text_zw) +
  # geom_zwbel(aes(specificity +.5 , sensitivity + 0.5, label = label), 
  #            hjust = 0, nudge_x = 3,
  #            label.size = NA
  #            ) +
  scale_x_reverse("Specificity (%)", limits = c(100, 0),  breaks = scales::breaks_extended(10)) + 
  scale_y_continuous("Sensitivity (%)", limits = c(0, 100), breaks = scales::breaks_extended(10)) 


uva_values_plot_zw$layers[[4]] <- NULL

uva_values_plot_zw <- uva_values_plot_zw +
  ggtitle("UVA")

mews_values_plot_zw$layers[[4]] <- NULL

mews_values_plot_zw <- mews_values_plot_zw +
  ggtitle("MEWS")

qsofa_values_plot_zw$layers[[4]] <- NULL

qsofa_values_plot_zw <- qsofa_values_plot_zw +
  ggtitle("qSOFA")

combined_zw <- cowplot::plot_grid(mews_values_plot_zw, qsofa_values_plot_zw, uva_values_plot_zw, ncol = 3, 
                               labels = c('A', 'B', "C"), label_size = 12, 
                               align = "h")

combined_zw


```


```{r, include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T }

mz_uva_v_qsofa <- roc.test(uva_auc_zw, qsofa_auc_zw, method = c("delong"))
mz_uva_v_qsofa_z <- as.numeric(mz_uva_v_qsofa$statistic)
mz_uva_v_qsofa_p <- format(mz_uva_v_qsofa$p.value, scientific = F)

mz_uva_v_mews <- roc.test(uva_auc_zw, mews_auc_zw, method = c("delong"))
mz_uva_v_mews_z <- as.numeric(mz_uva_v_mews$statistic)
mz_uva_v_mews_p <- format(mz_uva_v_mews$p.value, scientific = F)

mz_mews_v_qsofa <- roc.test(mews_auc_zw, qsofa_auc_zw, method = c("delong"))
mz_mews_v_qsofa_z <- as.numeric(mz_mews_v_qsofa$statistic)
mz_mews_v_qsofa_p <- format(mz_mews_v_qsofa$p.value, scientific = F)

```

DeLong test for Zimbabwe site;   
  1) UVA (AUC:`r uva_auc_value_zw`, `r uva_auc_ci_zw`) vs. MEWS (AUC:`r mews_auc_value_zw`,  `r mews_auc_ci_zw`), Z = `r mz_uva_v_mews_z`, *p-value* = `r mz_uva_v_mews_p`.   
  2) UVA (AUC:`r uva_auc_value_zw`, `r uva_auc_ci_zw`) vs. qSOFA (AUC:`r qsofa_auc_value_zw`,  `r qsofa_auc_ci_zw`), Z = `r mz_uva_v_qsofa_z`, *p-value* = `r mz_uva_v_qsofa_p`.   
  3) MEWs (AUC:`r mews_auc_value_zw`,  `r mews_auc_ci_zw`) vs. qSOFA (AUC:`r qsofa_auc_value_zw`,  `r qsofa_auc_ci_zw`), Z = `r mz_mews_v_qsofa_z`, *p-value* = `r mz_mews_v_qsofa_p`.   


\newpage
## Figure S10: MEWS + HIV ROCs

```{r mews_hiv_roc, fig.cap = "MEWS ROCs for A) MEWS, B) MEWS HIV, C) UVA", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T, fig.width = 8, fig.height=5}

# MEWS HIV ROC

mews_hiv_auc_value <- round(as.numeric(auc(mews_hiv_auc)[1])/100, 2)

mews_hiv_values <- coords(roc = mews_hiv_auc, x = "all", transpose = FALSE, ret = "all")

mews_hiv_auc_ci <- paste0("95% CI: ", round(pROC::ci.auc(mews_hiv_auc)[1]/100, 2), "-", round(ci.auc(mews_hiv_auc)[3]/100, 2))

mews_hiv_auc_fig_text <- paste0("AUC = ", mews_hiv_auc_value, "\n", mews_hiv_auc_ci)

mews_hiv_values_plot <- mews_hiv_values %>%
  ggplot(mapping = aes(x = specificity, y = sensitivity)) +
  geom_line() +
  geom_point() +
  annotate("text", x = 50, y = 50, label = mews_hiv_auc_fig_text) +
  # geom_zwbel(aes(specificity +.5 , sensitivity + 0.5, label = label), 
  #            hjust = 0, nudge_x = 3,
  #            label.size = NA
  #            ) +
  scale_x_reverse("Specificity (%)", limits = c(100, 0),  breaks = scales::breaks_extended(10)) + 
  scale_y_continuous("Sensitivity (%)", limits = c(0, 100), breaks = scales::breaks_extended(10)) +
  ggtitle("MEWS + HIV") +
  theme_minimal()


combined_mews_hiv <- cowplot::plot_grid(mews_values_plot, mews_hiv_values_plot, uva_values_plot, ncol = 3, 
                               labels = c('A', 'B', "C"), label_size = 12, 
                               align = "h")

combined_mews_hiv

# MEWS vs MEWS+HIV 
mews_v_mews_hiv <- roc.test(mews_hiv_auc, mews_auc, method = c("delong"))
mews_v_mews_hiv_z <- as.numeric(mews_v_mews_hiv$statistic)
mews_v_mews_hiv_p <- format(mews_v_mews_hiv$p.value, scientific = F)

# UVA vs MEWS+HIV
uva_v_mews_hiv <- roc.test(mews_hiv_auc, uva_auc, method = c("delong"))
uva_v_mews_hiv_p <- format(uva_v_mews_hiv$p.value, scientific = F)


```


\newpage

## Figure S11: qSOFA + HIV ROCs

```{r qsofa_hiv_roc, fig.cap = "qSOFA ROCs for A) qSOFA, B) qSOFA HIV, C) UVA", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T, fig.width = 8, fig.height=5}

# qsofa HIV ROC

qsofa_hiv_auc_value <- round(as.numeric(auc(qsofa_hiv_auc)[1])/100, 2)

qsofa_hiv_values <- coords(roc = qsofa_hiv_auc, x = "all", transpose = FALSE, ret = "all")

qsofa_hiv_auc_ci <- paste0("95% CI: ", round(pROC::ci.auc(qsofa_hiv_auc)[1]/100, 2), "-", round(ci.auc(qsofa_hiv_auc)[3]/100, 2))

qsofa_hiv_auc_fig_text <- paste0("AUC = ", qsofa_hiv_auc_value, "\n", qsofa_hiv_auc_ci)

qsofa_hiv_values_plot <- qsofa_hiv_values %>%
  ggplot(mapping = aes(x = specificity, y = sensitivity)) +
  geom_line() +
  geom_point() +
  annotate("text", x = 50, y = 50, label = qsofa_hiv_auc_fig_text) +
  # geom_zwbel(aes(specificity +.5 , sensitivity + 0.5, label = label), 
  #            hjust = 0, nudge_x = 3,
  #            label.size = NA
  #            ) +
  scale_x_reverse("Specificity (%)", limits = c(100, 0),  breaks = scales::breaks_extended(10)) + 
  scale_y_continuous("Sensitivity (%)", limits = c(0, 100), breaks = scales::breaks_extended(10)) +
  ggtitle("qSOFA + HIV") +
  theme_minimal()


combined_qsofa_hiv <- cowplot::plot_grid(qsofa_values_plot, qsofa_hiv_values_plot, uva_values_plot, ncol = 3, 
                               labels = c('A', 'B', "C"), label_size = 12, 
                               align = "h")

combined_qsofa_hiv



```

## Table S4: All AUC results

```{r, tab.cap = "ROC AUC values for complete data and senstivity analyses", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T }


score <- c('MEWS','qSOFA','UVA')

complete_data <- c(paste0(mews_auc_value, "\n", mews_auc_ci), 
                   paste0(qsofa_auc_value, "\n", qsofa_auc_ci),
                   paste0(uva_auc_value,"\n", uva_auc_ci)
)

sensitivity_alive <- c(paste0(mews_auc_value_s, "\n", mews_auc_ci_s), 
                   paste0(qsofa_auc_value_s, "\n", qsofa_auc_ci_s),
                   paste0(uva_auc_value_s,"\n", uva_auc_ci_s)
)


sensitivity_dead <- c(paste0(mews_auc_value_s_d, "\n", mews_auc_ci_s_d), 
                   paste0(qsofa_auc_value_s_d, "\n", qsofa_auc_ci_s_d),
                   paste0(uva_auc_value_s_d,"\n", uva_auc_ci_s_d)
)


inpatients <- c(paste0(mews_auc_value_inp, "\n", mews_auc_ci_inp),
                paste0(qsofa_auc_value_inp, "\n", qsofa_auc_ci_inp),                
                paste0(uva_auc_value_inp,"\n", uva_auc_ci_inp)
)

outpatients <- c(paste0(mews_auc_value_op, "\n", mews_auc_ci_op),
                paste0(qsofa_auc_value_op, "\n", qsofa_auc_ci_op),                
                paste0(uva_auc_value_op,"\n", uva_auc_ci_op)
)

laos <- c(paste0(mews_auc_value_la, "\n", mews_auc_ci_la),
                paste0(qsofa_auc_value_la, "\n", qsofa_auc_ci_la),                
                paste0(uva_auc_value_la,"\n", uva_auc_ci_la)
)

malawi <- c(paste0(mews_auc_value_mw, "\n", mews_auc_ci_mw),
                paste0(qsofa_auc_value_mw, "\n", qsofa_auc_ci_mw),                
                paste0(uva_auc_value_mw,"\n", uva_auc_ci_mw)
)

moz <- c(paste0(mews_auc_value_mz, "\n", mews_auc_ci_mz),
                paste0(qsofa_auc_value_mz, "\n", qsofa_auc_ci_mz),                
                paste0(uva_auc_value_mz,"\n", uva_auc_ci_mz)
)

zw <- c(paste0(mews_auc_value_zw, "\n", mews_auc_ci_zw),
                paste0(qsofa_auc_value_zw, "\n", qsofa_auc_ci_zw),                
                paste0(uva_auc_value_zw,"\n", uva_auc_ci_zw)
)

auc_all <- data.frame(score, complete_data, sensitivity_alive, sensitivity_dead, inpatients, outpatients, laos, malawi, moz, zw)
  
auc_all <- auc_all %>%
  pivot_longer(
    cols = complete_data:zw,
    names_to = "analysis",
    values_to = "auc"
  )

auc_all <- auc_all %>%
  mutate(auc = str_replace(auc, "95% CI: ", " ("),
         auc = str_replace(auc, "\\d{2}$", paste0(str_extract(auc, "\\d{2}$"), ")"))
         )

auc_all <- auc_all %>%
  pivot_wider(
    # cols = score,
    names_from = "score",
    values_from = "auc"
  ) %>%
  select(analysis, MEWS, qSOFA, UVA)



# auc_all$complete_data <- sjlabelled::set_label(auc_all$complete_data, "Complete cases")
# auc_all$sensitivity_alive <- sjlabelled::set_label(auc_all$sensitivity_alive, "Missing set to alive (a)")
# auc_all$sensitivity_dead <- sjlabelled::set_label(auc_all$sensitivity_dead, "Missing set to dead (b)")

# tbl_summary(auc_all)


auc_all_tbl <- auc_all %>%
  flextable()  %>%
  # align()
  add_header(MEWS = "Severity score AUC (95% CI)", qSOFA = "Severity score AUC (95% CI)", UVA = "Severity score AUC (95% CI)") %>%
   merge_h(part = "header")

auc_all_tbl


# kableExtra::
# auc_all

```

\newpage

```{r flow_chart, fig.cap = "Flowchart of adult fever patients analysed", echo = F, message = F, fig.width=9, fig.height = 6}


sample <- nrow(severity_data_cc_d28)


miss1 <- severity_data %>%
  mutate(outcome_d28 = as.character(outcome_d28)) %>%
  filter(!is.na(outcome_d28)) %>%
  filter(outcome_d28 == "lost to fup") %>%
  nrow()


miss_scores <- severity_data_v1 %>%
    filter(if_any(c(modGCS, qsofa, uva, mews), ~ is.na(.))) %>%
    nrow()


miss3 <- miss_scores

total_missing <- (miss1 + miss3)

total_missing_p <- paste0(round(total_missing/total_px*100, 0), "%")

     
grid.newpage()


org_cohort <- boxGrob(glue("Total number of participants (≥15 years) with fever enrolled",
                           "n = {pop}",
                           pop = txtInt(total_px),
                           .sep = "\n"),
                      x = .5,
                      y = .8
                      )

included <- boxGrob(glue("Participants (≥15 years) with complete data analysed",
                         "n = {sample}",
                         sample = txtInt(sample),
                         .sep = "\n"),
                    x = .5, 
                    y = .2
                    )

excluded <- boxGrob(glue("Excluded n = {total_missing} ({total_missing_p}):",
                         " - Loss to follow-up: {miss1}",
                         " - Missing severity score \n (UVA, MEWS, qSOFA): {miss3}",
                         .sep = "\n"),
                    just = "left"
                    )



# vert <- spreadVertical(org_cohort,
#                        included
#                        )

excluded <- moveBox(excluded,
                    x = 0.8,
                    y = .5
                    )


org_cohort
included
excluded
connectGrob(org_cohort, excluded, "L")
connectGrob(org_cohort, included, type = "vert") 



```

## Figures - physiological signs 

```{r fig_temperature, fig.cap = "Distribution of temperature", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T}



temp <- severity_data_cc_d28 %>%
  ggplot(mapping = aes(x = temp)) +
  geom_histogram( color='#e9ecef', alpha=0.8, bins = 50, boundary = 0, closed = "left") +
  scale_x_continuous(breaks=seq(37.5, 43, 2)) +
  facet_wrap(vars(country)) 

temp


```


```{r fig_heart_rate, fig.cap = "Distribution of heart rate", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T}


# summary(severity_data_cc_d28$heart_rate)

heart_rate <- severity_data_cc_d28 %>%
  ggplot(mapping = aes(x = heart_rate)) +
  geom_histogram( color='#e9ecef', alpha=0.8, bins = 50, boundary = 0, closed = "left") +
  scale_x_continuous(breaks=seq(44, 192, 20)) +
  facet_wrap(vars(country)) 

heart_rate


```

```{r fig_respiratory_rate, fig.cap = "Distribution of respiratory rate", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T}


# summary(severity_data_cc_d28$respiratory_rate)

respiratory_rate <- severity_data_cc_d28 %>%
  ggplot(mapping = aes(x = respiratory_rate)) +
  geom_histogram( color='#e9ecef', alpha=0.8, bins = 15, boundary = 0, closed = "left") +
  scale_x_continuous(breaks=seq(10, 100, 5)) +
  facet_wrap(vars(country)) 

respiratory_rate


```

```{r fig_o2_sats, fig.cap = "Distribution of o2_sats", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = T}


# summary(severity_data_cc_d28$o2_sats)

o2_sats <- severity_data_cc_d28 %>%
  ggplot(mapping = aes(x = o2_sats)) +
  geom_histogram( color='#e9ecef', alpha=0.8, bins = 30, boundary = 0, closed = "left") +
  scale_x_continuous(breaks=seq(20, 100, 20)) +
  facet_wrap(vars(country)) 

o2_sats


```

```{r age_death, fig.cap = "Distribution of patient outcomes by age", include = T, results = 'asis', echo = F, message = F, warning = F, paged.print = F, eval = F}

fig3 <- ggplot(
  data = severity_data_cc_d28, 
  aes(y = age,           # numeric variable
      x = outcome_d28)) +    # group variable
  geom_violin(
    aes(fill = outcome_d28), # fill (color of violin background)
    color = "white",     # white outline
    alpha = 0.2)+        # transparency
  geom_sina(
    size=1,                # Change the size of the jitter
    aes(color = outcome_d28))+ # color (color of dots)
  scale_fill_manual(       # Define fill for violin background by death/recover
    values = c("Death" = "#bf5300", 
              "Recover" = "#11118c")) + 
  scale_color_manual(      # Define colours for points by death/recover
    values = c("Death" = "#bf5300", 
              "Recover" = "#11118c")) + 
  theme_minimal() +                                # Remove the gray background
  theme(legend.position = "none") +                # Remove unnecessary legend
  labs(title = "") 

fig3


```

<!---BLOCK_LANDSCAPE_STOP--->



